---
title: "Appendix D. Provisional EVMP and East Troublesome Fire analysis"
author: ''
date: ''
output:
  html_document:
    toc: yes
    fig_caption: yes
    theme: paper
  word_document:
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.align = 'center', cache = FALSE)
opts_knit$set(root.dir=normalizePath('../')) # this is required if Rmd is nested below the project directory
opts_chunk$set(fig.path = "../output/figures/") # specify a subdirectory for figs

```

**Updated:** 11/9/2020
```{r, echo=FALSE}
library(ggspatial)
library(tidyverse)
library(sf)
library(mapview)
library(gganimate)
library(DT)
library(janitor)
library(gt)
library(patchwork)
library(gtsummary)
```

## Data and methods
This appendix provides a provisional assessment of EVMP plots in relation to East Troublesome fire perimeter data. Fire perimeter data approved from the 2020 National Incident Feature Service (NIFS) were downloaded from:
https://data-nifc.opendata.arcgis.com/datasets/archived-wildfire-perimeters-2

EVMP plots were encoded as “burned” for 2020 if they fell within the 10/27/2020 fire perimeter. Attribution _does not represent field-verified fire effects on the plots_. Data were processed using functions in the "sf" package for R.

```{r, message=FALSE}
## read in EVMP sites and set same crs
evmp <- st_read("data/EVMP_derived/site_info_clean.shp", quiet = TRUE)

evmp.4326 <- st_transform(evmp,4326)


# et.1027 <- st_transform(et.1027,2957)

```

```{r}
# EPSG:2957
# NAD83(CSRS) / UTM zone 13N (Google it)
# 
# WGS84 Bounds: -108.0000, 49.0000, -102.0000, 81.0000
# Projected Bounds: 280586.2987, 5431792.8644, 719413.7013, 8994558.8477
# Scope: Large and medium scale topographic mapping and engineering survey.
# Last Revised: May 29, 2007
# Area: Canada - 108°W to 102°W
```


```{r, message=FALSE}
# C:\Users\Ed\Git_repos\EVMP\data\fire_2020
peri.arch <- st_read("data/fire_2020/Public_NIFS_Archive_Perimeters.shp", quiet = TRUE)

```

## Tables and figures

### East Troublesome Fire perimeter

```{r}
#
et <- peri.arch %>%
  filter(IncidentNa == "East Troublesome")

et <- et %>% 
  filter(GDB_FROM_D != "2020-11-02" & GDB_FROM_D != "2020-10-25")



```

```{r, eval=FALSE}
st_crs(et)
## calc area
et$area <- st_transform(et, 26913) %>% 
  st_area(et)

et <- et %>%
  mutate(area = as.numeric(area)) %>% 
  mutate(area_km2 = area/10000) %>% 
  mutate(area_acre = area*0.000247105)

# scrap
# et %>% 
#   mutate(date = anytime::anydate(GDB_FROM_D)) %>%
#   filter(date < "2020-10-25") %>% 
#   group_by(GDB_FROM_D) %>%
#   summarise(area_acre = sum(GISAcres)) %>%
#   mutate(area_acre = round(area_acre,2)) %>% 
#   ggplot(aes(GDB_FROM_D,area_acre)) +
#   geom_label(aes(label = area_acre)) +
#   geom_col() +
#   theme_minimal()
```



```{r}
pl.growth <- et %>%
  ggplot() +
  # annotation_map_tile(type = "http://c.tiles.wmflabs.org/hillshading/${z}/${x}/${y}.png",
  #                     zoom = 14) +
  geom_sf() +
  annotation_scale(location = "br") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~GDB_FROM_D, ncol = 3) +
  theme_void() +
  # theme_minimal() +
  labs(title = "East Troublesome perimeters",caption = "National Interagency Fire Center data")

pl.growth

```

```{r}
et.1027 <- et %>%
  filter(GDB_FROM_D == "2020-10-27")
# mapview(et.1027)
```

```{r, eval=FALSE}
et.ani <- et %>%
  filter(!is.na(GDB_FROM_D)) %>% 
  ggplot() +
  geom_sf() +
  transition_states(GDB_FROM_D,
                    transition_length = 2,
                    state_length = 1)

et.ani
```

### EVMP plots and 10/27/2020 fire perimeter

```{r}
# mapview(et.1027)
# mapview(evmp.4326)

map01 <- et.1027 %>%
  ggplot() +
  annotation_map_tile(type = "http://c.tiles.wmflabs.org/hillshading/${z}/${x}/${y}.png",
                      zoom = 11) +
  annotation_spatial(et.1027, color= "red", alpha = .31) +
  layer_spatial(evmp, aes(col = pType)) +
  viridis::scale_color_viridis(discrete = TRUE) +
  theme_minimal() +
  labs(title = "10/27 fire perimeter and EVMP Plots", color = "")
map01

# ggsave("output/figures_exported/fire2020_points_map.png", width = 7, height = 6, dpi=300)

```

```{r}
## write to disk
# et.1027 %>% 
  # st_write("output/exported_data/spatial/east_trbl_1027.shp")
```


```{r, eval=FALSE}
ggplot() +
  annotation_spatial(et.1027, fill = "grey") +
  layer_spatial(evmp, aes(col = pType)) +
  annotation_scale(location = "tl") +
  annotation_north_arrow(location = "br", which_north = "true") +
  theme_minimal() +
  labs(color = "")
# ggsave("output/figures_exported/east_troublesome1.png", width = 7, height = 7, dpi = 230)
```


```{r}
# join 
plots_in_burn <- st_join(evmp.4326, et.1027, join = st_within)

## clean
plots_in_burn <- plots_in_burn %>% 
  mutate(burned2020 = case_when(is.na(IncidentNa) ~ "Unburned",
                                TRUE ~ "Burned"))%>%
  select(1:13,burned2020)

plots_in_burn <- plots_in_burn %>% 
  mutate(BURNED = case_when(BURNED == "Unburned" ~ "Unburned",
                            BURNED == "N Y" ~ "Unburned",
                            BURNED == "Completely" ~ "Burned",
                            BURNED == "Moderately" ~ "Burned",
                            BURNED == "Moderately to Completely" ~ "Burned",
                            BURNED == "Y" ~ "Burned",
                            BURNED == "N" ~ "Unburned",
                            BURNED == "Not burned" ~ "Unburned",
                            is.na(BURNED) ~ "Unburned",
                            TRUE ~ BURNED)
         ) %>% 
  mutate(FENCED = case_when(FENCED == "Y_but_fence_down_since_2013" ~ "Unfenced",
                             FENCED == "Y" ~ "Fenced",
                             FENCED == "N" ~ "Unfenced"))

```

```{r, eval=FALSE}
## write shapefile to disk
plots_in_burn %>%
  st_write("output/exported_data/data_deliverables/evmp_sites.shp", append = FALSE)
```

```{r, eval=FALSE}
plots_in_burn %>% 
  mapview(zcol = "burned2020")

```


```{r}
# clean
plots_in_burn <- plots_in_burn %>%
  as_tibble() %>%
  rename(burned2012 = BURNED) %>% 
  mutate(burned2012 = case_when(burned2012 == "N" ~ "Unburned",
                                burned2012 == "Y" ~ "Burned",
                                TRUE ~ burned2012)) %>% 
  select(-c(GDB_ARCHIV, REMOVED, OBJECTID, WILDERN, IncidentNa, EastWst, VALLEY)) %>%  
  relocate(burned2020, .after = burned2012) %>% 
  relocate(pType, .after = SITE_ID) 

# write for csv table

# plots_in_burn %>%
#   gt() %>%
#   tab_header(title = "EVMP plots: 2012 and 2020 burn status") %>%
#   gt::gtsave("output/exported_data/evmp_burned_2012_2020.rtf")
# %>%
# write_csv("output/exported_data/evmp_burned_2012_2020.csv")

```

### Count of plots in 10/27/2020 burn perimeter by valley

Note: Plots were encoded as "burned" for 2020 if they fell within the 10/27/2020 fire perimeter; fire effects have not been assessed in the field.

```{r}

plots_in_burn %>%
  as_tibble() %>% 
  tabyl(burned2020) %>% 
  arrange(burned2020) %>%
  rename(Burned = burned2020) %>%
  mutate(percent = round(100*percent,1)) %>% 
  gt() %>% 
  tab_header(title = "In 2020 burn perimeter") 


```

### Count of plots in 10/27/2020 burn perimeter by plot type

Note: Plots were encoded as "burned" for 2020 if they fell within the 10/27/2020 fire perimeter; fire effects have not been assessed in the field.

```{r}
plots_in_burn %>% 
  as_tibble() %>% 
  tabyl(pType, burned2020) %>% 
  arrange(-(Burned)) %>% 
  gt() %>% 
  tab_header(title = "In 2020 burn perimeter")

```

### Count of plots in 10/27/2020 burn perimeter by valley

Note: Plots were encoded as "burned" for 2020 if they fell within the 10/27/2020 fire perimeter; fire effects have not been assessed in the field.

```{r}


plots_in_burn %>% 
  as_tibble() %>% 
  tabyl(vlly_fl,burned2020) %>% 
  rename(Valley = vlly_fl) %>% 
  arrange(-(Burned)) %>% 
  gt()%>% 
  tab_header(title = "In 2020 burn perimeter")

```

### Count of fenced plots in 10/27/2020 burn perimeter

Note: Plots were encoded as "burned" for 2020 if they fell within the 10/27/2020 fire perimeter; fire effects have not been assessed in the field.

```{r}
plots_in_burn %>%
  as_tibble() %>% 
  tabyl(FENCED, burned2020) %>% 
  rename(Fenced = FENCED) %>% 
  gt()

```


