---
title: "Appendix D. EVMP Upland Statistical Analyses"
author: ""
date: ""
output: 
  html_document: 
    theme: journal
    toc: yes
    toc_depth: 2
    code_folding: hide
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

_Code author:_ 
E. Gage  
mycologica@gmail.com  
https://github.com/egage/EVMP  
 

```{r setup, include=FALSE, echo=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = 'center')
opts_knit$set(root.dir=normalizePath('../')) # this is required if Rmd is nested below the project directory
opts_chunk$set(fig.path = "../output/figures/")
set.seed(1234)
```

```{r,pkg_load, echo=FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(fs))
suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(janitor))
suppressPackageStartupMessages(library(readxl))
library(glue)
suppressPackageStartupMessages(library(mapview))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(ggstance))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(lubridate))
# suppressPackageStartupMessages(library(skimr)) 
suppressPackageStartupMessages(library(dataMaid))
library(lubridate)
library(emmeans)
library(rstanarm)
options(mc.cores = parallel::detectCores())
library(bayestestR)
library(gt)
library(patchwork)
library(ggtext)

```

```{r, palletes}
# custom gradients and palettes
dbhclass.gradient5 <- c("#26185F", "#0095AF", "#9ADCBB", "#FCFFDD", "grey50")

colfunc1<-colorRampPalette(c("#26185F","#0095AF","#9ADCBB", "#FCFFDD"))
colfunc2<-colorRampPalette(c("#0095AF","#9ADCBB", "#FCFFDD"))
colfunc3<-colorRampPalette(c("#0095AF","#9ADCBB"))
colfunc3_RdBu <- colorRampPalette(c("#EF8A62", "#FDDBC7", "#D1E5F0", "#67A9CF"))

# plot(rep(1,50),col=(colfunc1(50)), pch=19,cex=2)
# plot(rep(1,5),col=(colfunc1(5)), pch=19,cex=2)
pal.5a <- colfunc1(5)
pal.7a <- colfunc1(7)

# palette3 <- RColorBrewer::brewer.pal(6,"RdBu")
# RColorBrewer::display.brewer.pal(6,"RdBu")
```


```{r, funcs, echo=FALSE, eval = TRUE}

## Functions
### Table functions

# SEXIT summary
fun.sexit.gt <- function(model){
  bayestestR::sexit(model, ci=0.9) %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric), round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Effect Existence and Significance Testing") 
  }

## model summary to DT
fun.model.summary <- function(model){
  summary(model,
          probs = c(0.1, 0.5, 0.9),
        digits = 1) %>%
  as_tibble(rownames = "Parameter") %>%
  mutate(across(where(is.numeric),round,2)) %>%
  datatable(caption = "Model summary")
}

## function to describe model out
fun.desc.post1 <- function(model){
  describe_posterior(
  model,
  effects = "fixed",
  component = "all",
  ci_method = "hdi",
  ci=0.9,
  rope_ci=0.9,
  test = c("p_direction", "p_significance"),
  centrality = "all"
) %>% 
    as_tibble() %>% 
    DT::datatable(caption = "Posterior description")
}


## gt
fun.desc.post1.gt <- function(model){
  describe_posterior(
  model,
  effects = "fixed",
  component = "all",
  ci_method = "hdi",
  test = c("rope"),
  #rope_range(c(-0.1,0.1)),
  rope_range("default"),
  centrality = "median"
) %>% 
    as_tibble() %>% 
    gt() %>% 
    tab_header(title = "Posterior description") %>% 
    fmt_number(
    columns = 2:12,
    decimals = 1,
    suffixing = TRUE
  )
}

## adds 
fun.desc.post1.gt.rope <- function(model){
  describe_posterior(
  model,
  effects = "fixed",
  component = "all",
  ci_method = "hdi",
  test = c("p_direction", "p_significance","rope"),
  rope_range("default"),
  centrality = "median"
) %>% 
    as_tibble() %>% 
    gt() %>% 
    tab_header(title = "Posterior description") %>% 
    fmt_number(
    columns = 2:11,
    decimals = 1,
    suffixing = TRUE
  )
}

```
# below check contrast plot spec to linerange color  
```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=FALSE, eval=TRUE}
#### Plotting functions

fun.contrast.plot <- function(df){
  df %>% 
  ggplot(aes(x = contrast)) +
  geom_linerange(aes(ymin = lower.HPD, ymax = upper.HPD), color = "#0095AF", size = 2) +
  geom_point(aes(y = estimate), size = 2.5) +
  theme_minimal() +
  labs(caption = "Point estimate displayed: median 
HPD interval probability", x = "Estimate", y = "Estimate") +
  coord_flip()
}

```


```{r, eval=FALSE}
# purrr plots
tmp1 <- function(df){
  dcanopy.all.willow.sum  %>% 
  group_nest(SITE_TYPE) %>% 
  mutate(plots = map2(.y = SITE_TYPE, .x = data, ~{ggplot(data = .x) +
                              geom_boxplot(aes(y = cover.allwillow, x = timeClass), fill="Gray80") +
                              geom_hline(aes(xintercept=0.31)) +
                              # ggtitle(paste0("Willow cover, site: ", .y)) +
                              labs(title = paste0("Site type: ", .y), x = "", y = "Willow cover (%)") +
                              theme_minimal()}))

tmp1 %>% 
  pull(plots)

```


```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=FALSE, eval=TRUE}
## ppcheck.plot
ppc.plot <- function(model){
  pp_check(model) +
  labs(x = "Value", y = "Density") +
  theme_minimal()
}


```

# Introduction

This document contains code for running statistical analyses of of Elk
Vegetation Management Plan (EVMP) data collected through the 2018
sampling season. Analyses are limited to data from plots established in
upland communities in areas of the elk winter range (core and noncore
area, see Zeigenfuss et al. 2011 for details) and is provided as a
supplement to the NPS NRR Report "Monitoring of Vegetation Response to
Elk Population and Habitat Management in Rocky Mountain National Park -
Analysis of Elk Vegetation Management Plan monitoring Data: 2008--2018".
For information on sampling and the broader analysis and interpretation
of the results, refer to this report, past analyses [@zeigenfuss2015],
and the original EVMP monitoring plan [@zeigenfuss2011].

# Methods

The code and results presented here start from and build upon derived
data produced in a separate code document focused on ingesting,
compiling, and cleaning raw data provided by RMNP staff. Raw files were combined and pre-processed in a separate code file (Appendix B for the main project report).
Packages included in the "Tidyverse"
ecosystem of R packages were used for data transformation and
visualization [@wickham2019], although specialized packages particular
to specific tasks (e.g., the "bayesplot" plot for visualization of
Bayesian posterior distributions) were also used [@gabry2019].

Bayesian repeated measures analyses were fit separately for combined
core and non-core winter range, core winter range (WC) and non-core
winter range plots (WNC) with plots treated as random factors using the
'stan_glmer' function in the rstanrm package
[@goodrich2020][@brilleman2018]. Bayesian estimation was performed via
MCMC adding independent weakly informed priors specific to the data type
being modeled for regression coefficients for difference response
variable types [@lemoine2019]. Zero-bounded continuous response variables (e.g.,
shrub height) were fit with a Gamma distribution, count data were modeled as poisson processes, while proportion data (e.g.,
cover) were modeled using a beta distribution (Ferrari and Cribari-Neto
2004) that constrains values from 0 to 1.

```{r}
### Data import and munging
## read in derived data
upl.allshrubs <- read_csv("./output/exported_data/upland_LI_allShrubsPooled_20200814.csv")
upl.byspp <- read_csv("./output/exported_data/upland_LI_byShrubSpp_20200814.csv")

### upland pooled species
cov.allshrubs <- upl.allshrubs %>% 
  filter(!is.na(perc_cover)) %>%  
  mutate(time_class = case_when(yr == 2007 ~ "BL",
                              TRUE ~ as.character(yr))) %>% 
  mutate(timeClass = as.factor(time_class)) %>%
  mutate(site_type = as.factor(site_type)) %>%
  mutate(time_class = fct_relevel(timeClass, "2018", "2013", "BL")) 

## reorder the time class
cov.allshrubs <- cov.allshrubs %>% 
  mutate(time_class = fct_rev(time_class))

### upland by species
cov.byspp <- upl.byspp %>% 
  filter(!is.na(perc_cover)) %>%  
  mutate(time_class = case_when(yr == 2007 ~ "BL",
                              TRUE ~ as.character(yr))) %>% 
  mutate(time_class = as.factor(time_class)) %>%
  mutate(site_type = as.factor(site_type)) %>%
  mutate(time_class = fct_relevel(time_class, "2018", "2013", "BL")) 

## reorder the time class
cov.byspp <- cov.byspp %>% 
  mutate(time_class = fct_rev(time_class))
# %>% 
#   mutate(timeClass = case_when(yr == 2008 ~ "BL",
#                               yr == 2009 ~ "BL", 
#                               TRUE ~ as.character(yr))) %>% 
#   mutate(timeClass = as.factor(timeClass)) %>% 
#   mutate(timeClass = fct_relevel(timeClass, "2018", "2013", "BL"))
```

```{r, echoFALSE, eval=FALSE}
## create seperate subsets of ac and anc
cov.allshrubs %>% 
  glimpse()

cov.allshrubs %>% 
  names()

# cov.allshrubs %>% 
#   tabyl()

```

```{r}
# join in site info
## join in the site info

site.info.clean <- read_csv("./data/EVMP_derived/site_info_clean.csv")
site.info.cleannames <- site.info.clean %>% 
  clean_names() %>% 
  select(-starts_with('utm')) %>% 
  select(-c(wilderness,east_west)) %>% 
  distinct() %>% 
  mutate(burned = case_when(burned == "Not burned" ~ "Unburned",
                            is.na(burned) ~ "Unburned",
                            TRUE ~ burned)) 

cov.allshrubs <- left_join(cov.allshrubs, site.info.cleannames, by = "site_id")
cov.byspp <- left_join(cov.byspp, site.info.cleannames, by = "site_id")
```

# Results
```{r}
# purrr plots
list.pl.allshrub.wcwnc <- cov.allshrubs %>% 
  group_nest(site_type) %>% 
  mutate(plots = map2(.y = site_type, .x = data, ~{ggplot(data = .x) +
                              geom_boxplot(aes(y = perc_cover, x = time_class), fill="Gray80") +
                              # geom_hline(aes(xintercept=0.31)) +
                              # ggtitle(paste0("Willow cover, site: ", .y)) +
                              labs(title = paste0("Site type: ", .y), x = "", y = "All shrub cover (%)") +
                              theme_minimal()}))

list.pl.allshrub.wcwnc %>% 
  pull(plots)

```

```{r, eval=FALSE}
# purrr plots
list.pl.allshrub.wc.f <- cov.allshrubs %>%
  filter(site_type == "core range") %>% 
  group_nest(fenced) %>% 
  mutate(plots = map2(.y = fenced, .x = data, ~{ggplot(data = .x) +
                              geom_boxplot(aes(y = perc_cover, x = time_class), fill="Gray80") +
                              # geom_hline(aes(xintercept=0.31)) +
                              # ggtitle(paste0("Willow cover, site: ", .y)) +
                              labs(title = paste0("Site type: ", .y), x = "", y = "All shrub cover (%)") +
                              theme_minimal()}))

list.pl.allshrub.wc.f %>% 
  pull(plots)

```


```{r}

```


## Upland Shrub Cover, Pooled Shrub Species - Combined Winter Range
```{r}
#### summary tables
cov.allshrubs %>% 
  group_by(time_class, site_type) %>% 
  summarytools::descr(perc_cover, stats = "common") %>% 
  summarytools::tb() %>%
  mutate_if(.predicate = is.numeric,.funs = round, digits = 2) %>% 
  gt() %>% 
  tab_header(title = "Upland cover - site type + time_class")
  
```

```{r, fig.width=7.5, fig.height=7.5}
cov.allshrubs %>% 
  filter(!is.na(range_type)) %>%
  ggplot(aes(time_class, perc_cover)) +
  geom_boxplot(fill="grey80") +
  geom_point(color="#0095AF") +
  theme_minimal() +
  labs(x="Time class", y="% cover") +
  facet_grid(burned~range_type)
  

# save plots
ggsave("./output/figures_202107_stats_upland/cover_range_type_boxplot.png", dpi = 300, width = 5.75, height = 3.75)
```


```{r, fig.width=7.5, fig.height=7.5}
cov.allshrubs %>% 
  filter(!is.na(valley_full)) %>% 
  ggplot(aes(time_class, perc_cover)) +
  geom_boxplot(fill="grey80") +
  geom_point(color="#0095AF") +
  theme_minimal() +
  labs(x="Time class", y="% cover") +
  facet_wrap(~valley_full, ncol=3)

# save plots
ggsave("./output/figures_202107_stats_upland/cover_valley_boxplot.png", dpi = 300, width = 6.75, height = 6.25)
```



### Modeling and Posterior Description
```{r, cache=TRUE}
## add cover in 0-1 range
cov.allshrubs <- cov.allshrubs %>%
  mutate(cover = perc_cover/100)
  
stmod_cov.allshr <- stan_glmer(cover ~ time_class + (1 | site_id), 
                      data = cov.allshrubs,
                      family = mgcv::betar,
                      prior_aux = exponential(2),
                      iter = 8000,
                      seed = 1234)

```

```{r}
# extract posteriors
posteriors.upl.allshrubcov <- insight::get_parameters(stmod_cov.allshr)

# tidy
posteriors.upl.allshrubcov.tidy <- posteriors.upl.allshrubcov %>%
  pivot_longer(cols = starts_with("time"), names_to = "parameter", values_to = "values")

```

**Prior summary**: cover ~ time_class + (1 | site_id)    
```{r}
prior_summary(stmod_cov.allshr)
```

```{r}
fun.model.summary(stmod_cov.allshr)
```

```{r}
posteriors.upl.allshrubcov.tidy %>%
  ggplot() +
  geom_density(aes(fill = parameter, x = values),color="grey", alpha = .25) +
  # scale_fill_viridis(discrete = TRUE) +
  scale_fill_manual(values = colfunc2(ncol(posteriors.upl.allshrubcov.tidy))) +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland, all shrub cover - Combined Winter Range")
```

```{r}
#### PP check plot
pp_check(stmod_cov.allshr) +
  labs(x = "Value", y = "Density", caption = "upland cover ppc plot. all shrubs pooled") +
  theme_minimal()

# save plots
ggsave("./output/figures_202107_stats_upland/ppc_stemcnt_upl_allshrubs_ucunc.png", dpi = 300, width = 4.75, height = 3.75)
# ggsave("./output/figures_202107_stats_upland/ppc_stemcnt_upl_allshrubs.pdf", width = 4.75, height = 3.75)
```

```{r}
fun.sexit.gt(stmod_cov.allshr$stanfit)

fun.sexit.gt(stmod_cov.allshr$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/xPUTR_allshrubs_stmod_cover1_posteriors.rtf")
```


```{r, echo=FALSE, eval=FALSE}
fun.desc.post1.gt.rope(stmod_cov.allshr$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range")

```

```{r, eval = FALSE}
## save
fun.desc.post1.gt(stmod_cov.allshr$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/PUTR_allshrubs_stmod_cover1_posteriors.rtf")

```
### Effect Description  
#### Probability of Direction
```{r}
# Visualize the pd
#### PD - AC
p_direction(stmod_cov.allshr) %>% plot() +
  theme_minimal() +
  scale_fill_manual(values = colfunc3(2)) +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland Shrub Cover Combined Winter Range") +
  scale_fill_manual(values = colfunc3(2))

# save plots
ggsave("./output/figures_202107_stats_upland/pd_uplcov_ucunc.png", dpi = 300, width = 4.75, height = 3.75)

```

#### Region of Practical Equivalence (ROPE)
```{r}
rope.uplcov.wcwnc.gt <- rope(stmod_cov.allshr$stanfit, ci=0.9) %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 2,
    suffixing = TRUE
  ) 

rope.uplcov.wcwnc.gt
# rope.uplcov.wcwnc.gt %>%
#   gt::gtsave(filename = "./output/tables/upland_cover/rope_uplcov_wcwnc.rtf")
```

### Contrasts

EMM summaries with type = "response", the tests and confidence intervals
are done before back-transforming. The ratios estimated are ratios of
geometric means. A model with a log response is in fact a model for
relative effects of any of its linear predictors, and this
back-transformation to ratios goes hand-in-hand with that.

```{r, include=FALSE, echo=FALSE}
# library(emmeans)
# Note: EMMs for later factor levels are subtracted from those for earlier levels; if you want the comparisons to go in the other direction, use pairs(pigs.emm.s, reverse = TRUE). Also, in multi-factor situations, you may specify by factor(s) to perform the comparisons separately at the levels of those factors.

# Results are given on the log (not the response) scale.
emmeans(stmod_cov.allshr, ~ time_class) %>% 
  pairs() %>% 
  plot(type="response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Combined Winter Range, All Shrub Cover")

# ggsave("./output/figures_202107_stats_upland/emm_shrubcov_wcwnc.pdf", width = 4.5, height = 3.55) # save plot
ggsave("./output/figures_202107_stats_upland/emm_shrubcov_wcwnc_unc.png", width = 4.5, height = 3.55) # save plot

# table of contrasts
emmeans(stmod_cov.allshr, ~ time_class) %>% 
  pairs() %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>%
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log scale")

# table of contrasts. Results are given on the response scale.
emmeans(stmod_cov.allshr, ~ time_class) %>% 
  pairs(reverse = FALSE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale.")

```

## Upland Shrub Cover, Pooled Shrub Species - Core Winter Range
```{r}
#### summary tables
cov.allshrubs %>% 
  group_by(time_class, site_type) %>% 
  summarytools::descr(perc_cover, stats = "common") %>% 
  summarytools::tb() %>%
  mutate_if(.predicate = is.numeric,.funs = round, digits = 2) %>% 
  gt() %>% 
  tab_header(title = "Upland cover - site type + time_class")

```

### Modeling and Posterior Description
```{r, cache=TRUE}
cov.allshrubs.uc <- cov.allshrubs %>% 
  filter(site_type == "core range") # add 

## add cover in 0-1 range
cov.allshrubs.uc <- cov.allshrubs.uc %>%
  mutate(cover = perc_cover/100)
  
stmod_cov.allshr.uc <- stan_glmer(cover ~ time_class + (1 | site_id), 
                      data = cov.allshrubs,
                      family = mgcv::betar,
                      prior_aux = exponential(2),
                      iter = 8000,
                      seed = 1234)

```

```{r}
# extract posteriors
posteriors.upl.allshrubcov.uc <- insight::get_parameters(stmod_cov.allshr.uc)

# tidy
posteriors.upl.allshrubcov.tidy <- posteriors.upl.allshrubcov %>%
  pivot_longer(cols = starts_with("time"), names_to = "parameter", values_to = "values")

```

**Prior summary**: cover ~ time_class + (1 | site_id)    
```{r}
prior_summary(stmod_cov.allshr.uc)
```

```{r}
fun.model.summary(stmod_cov.allshr.uc)
```

```{r}
posteriors.upl.allshrubcov.tidy %>%
  ggplot() +
  geom_density(aes(fill = parameter, x = values),color="grey", alpha = .25) +
  # scale_fill_viridis(discrete = TRUE) +
  scale_fill_manual(values = colfunc2(ncol(posteriors.upl.allshrubcov.tidy))) +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland, all shrub cover - Combined Winter Range")
```

```{r}
#### PP check plot
pp_check(stmod_cov.allshr.uc) +
  labs(x = "Value", y = "Density", caption = "upland cover ppc plot. all shrubs pooled") +
  theme_minimal()

# save plots
ggsave("./output/figures_202107_stats_upland/ppc_stemcnt_upl_allshrubs_uc.png", dpi = 300, width = 4.75, height = 3.75)
# ggsave("./output/figures_202107_stats_upland/ppc_stemcnt_upl_allshrubs.pdf", width = 4.75, height = 3.75)
```

```{r}
fun.sexit.gt(stmod_cov.allshr.uc$stanfit)

fun.sexit.gt(stmod_cov.allshr.uc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/xPUTR_allshrubs_stmod_cover1_posteriors.rtf")
```


```{r, echo=FALSE, eval=FALSE}
fun.desc.post1.gt.rope(stmod_cov.allshr.uc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range")

```

```{r, eval = FALSE}
## save
fun.desc.post1.gt(stmod_cov.allshr.uc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/PUTR_allshrubs_stmod_cover1_posteriors.rtf")

```
### Effect Description  
#### Probability of Direction
```{r}
# Visualize the pd
#### PD - AC
p_direction(stmod_cov.allshr.uc) %>% plot() +
  theme_minimal() +
  scale_fill_manual(values = colfunc3(2)) +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland Shrub Cover Combined Winter Range") +
  scale_fill_manual(values = colfunc3(2))

# save plots
ggsave("./output/figures_202107_stats_upland/pd_upl_uc.png", dpi = 300, width = 4.75, height = 3.75)

```


#### Region of Practical Equivalence (ROPE)
```{r}
rope.uplcov.wcwnc.gt <- rope(stmod_cov.allshr.uc$stanfit, ci=0.9) %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 2,
    suffixing = TRUE
  ) 

rope.uplcov.wcwnc.gt
# rope.uplcov.wcwnc.gt %>%
#   gt::gtsave(filename = "./output/tables/upland_cover/rope_uplcov_wcwnc.rtf")
```

### Contrasts

EMM summaries with type = "response", the tests and confidence intervals
are done before back-transforming. The ratios estimated are ratios of
geometric means. A model with a log response is in fact a model for
relative effects of any of its linear predictors, and this
back-transformation to ratios goes hand-in-hand with that.

```{r, include=FALSE, echo=FALSE}
# library(emmeans)
# Note: EMMs for later factor levels are subtracted from those for earlier levels; if you want the comparisons to go in the other direction, use pairs(pigs.emm.s, reverse = TRUE). Also, in multi-factor situations, you may specify by factor(s) to perform the comparisons separately at the levels of those factors.

# Results are given on the log (not the response) scale.
emmeans(stmod_cov.allshr.uc, ~ time_class) %>% 
  pairs() %>% 
  plot(type="response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Combined Winter Range, All Shrub Cover")

# ggsave("./output/figures_202107_stats_upland/emm_shrubcov_wcwnc.pdf", width = 4.5, height = 3.55) # save plot
ggsave("./output/figures_202107_stats_upland/emm_shrubcov_uc.png", width = 4.5, height = 3.55) # save plot

# table of contrasts
emmeans(stmod_cov.allshr.uc, ~ time_class) %>% 
  pairs() %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>%
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log scale")

# table of contrasts. Results are given on the response scale.
emmeans(stmod_cov.allshr.uc, ~ time_class) %>% 
  pairs(reverse = FALSE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale.")

```
## Upland Shrub Cover, Pooled Shrub Species - Noncore Winter Range
### Modeling and Posterior Description
```{r, cache=TRUE}
cov.allshrubs.unc <- cov.allshrubs %>% 
  filter(site_type == "non-core range") # add 

## add cover in 0-1 range
cov.allshrubs.unc <- cov.allshrubs.unc %>%
  mutate(cover = perc_cover/100)
  
stmod_cov.allshr.unc <- stan_glmer(cover ~ time_class + (1 | site_id), 
                      data = cov.allshrubs,
                      family = mgcv::betar,
                      prior_aux = exponential(2),
                      iter = 8000,
                      seed = 1234)

```

```{r}
# extract posteriors
posteriors.upl.allshrubcov.unc <- insight::get_parameters(stmod_cov.allshr.unc)

# tidy
posteriors.upl.allshrubcov.tidy <- posteriors.upl.allshrubcov %>%
  pivot_longer(cols = starts_with("time"), names_to = "parameter", values_to = "values")

```

**Prior summary**: cover ~ time_class + (1 | site_id)    
```{r}
prior_summary(stmod_cov.allshr.unc)
```

```{r}
fun.model.summary(stmod_cov.allshr.unc)
```

```{r}
posteriors.upl.allshrubcov.tidy %>%
  ggplot() +
  geom_density(aes(fill = parameter, x = values),color="grey", alpha = .25) +
  # scale_fill_viridis(discrete = TRUE) +
  scale_fill_manual(values = colfunc2(ncol(posteriors.upl.allshrubcov.tidy))) +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland, all shrub cover - Combined Winter Range")
```

```{r}
#### PP check plot
pp_check(stmod_cov.allshr.unc) +
  labs(x = "Value", y = "Density", caption = "upland cover ppc plot. all shrubs pooled") +
  theme_minimal()

# save plots
ggsave("./output/figures_202107_stats_upland/ppc_stemcnt_upl_allshrubs_unc.png", dpi = 300, width = 4.75, height = 3.75)
# ggsave("./output/figures_202107_stats_upland/ppc_stemcnt_upl_allshrubs.pdf", width = 4.75, height = 3.75)
```

```{r}
fun.sexit.gt(stmod_cov.allshr.unc$stanfit)

fun.sexit.gt(stmod_cov.allshr.unc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/xPUTR_allshrubs_stmod_cover1_posteriors.rtf")
```


```{r, echo=FALSE, eval=FALSE}
fun.desc.post1.gt.rope(stmod_cov.allshr.unc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range")

```

```{r, eval = FALSE}
## save
fun.desc.post1.gt(stmod_cov.allshr.unc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/PUTR_allshrubs_stmod_cover1_posteriors.rtf")

```
### Effect Description  
#### Probability of Direction
```{r}
# Visualize the pd
#### PD - AC
p_direction(stmod_cov.allshr.unc) %>% plot() +
  theme_minimal() +
  scale_fill_manual(values = colfunc3(2)) +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland Shrub Cover Combined Winter Range") +
  scale_fill_manual(values = colfunc3(2))

# save plots
ggsave("./output/figures_202107_stats_upland/pd_uplcov_unc.png", dpi = 300, width = 4.75, height = 3.75)

```


#### Region of Practical Equivalence (ROPE)
```{r}
rope.uplcov.wcwnc.gt <- rope(stmod_cov.allshr.unc$stanfit, ci=0.9) %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 2,
    suffixing = TRUE
  ) 

rope.uplcov.wcwnc.gt
# rope.uplcov.wcwnc.gt %>%
#   gt::gtsave(filename = "./output/tables/upland_cover/rope_uplcov_wcwnc.rtf")
```

### Contrasts

EMM summaries with type = "response", the tests and confidence intervals
are done before back-transforming. The ratios estimated are ratios of
geometric means. A model with a log response is in fact a model for
relative effects of any of its linear predictors, and this
back-transformation to ratios goes hand-in-hand with that.

```{r, include=FALSE, echo=FALSE}
# library(emmeans)
# Note: EMMs for later factor levels are subtracted from those for earlier levels; if you want the comparisons to go in the other direction, use pairs(pigs.emm.s, reverse = TRUE). Also, in multi-factor situations, you may specify by factor(s) to perform the comparisons separately at the levels of those factors.

# Results are given on the log (not the response) scale.
emmeans(stmod_cov.allshr.unc, ~ time_class) %>% 
  pairs() %>% 
  plot(type="response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Combined Winter Range, All Shrub Cover")

# ggsave("./output/figures_202107_stats_upland/emm_shrubcov_unc.pdf", width = 4.5, height = 3.55) # save plot
ggsave("./output/figures_202107_stats_upland/emm_shrubcov_unc.png", width = 4.5, height = 3.55) # save plot

# plot
emmeans(stmod_cov.allshr.unc, ~ time_class) %>% 
  pairs() %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>%
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log scale")

# table of contrasts. Results are given on the response scale.
emmeans(stmod_cov.allshr.unc, ~ time_class) %>% 
  pairs(reverse = FALSE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale.")

```



```{r}
# Point estimate displayed: median 
# HPD interval probability: 0.95

pl.emm.upl.allshr.ucunc <- emmeans(stmod_cov.allshr, ~ time_class) %>% 
  pairs() %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Combined Range", x = "Estimate", y = "Contrast")

pl.emm.upl.allshr.uc <- emmeans(stmod_cov.allshr.uc, ~ time_class) %>% 
  pairs() %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Core Range", x = "Estimate", y = "Contrast")

pl.emm.upl.allshr.unc <- emmeans(stmod_cov.allshr.unc, ~ time_class) %>% 
  pairs() %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Noncore Range", x = "Estimate", y = "Contrast")


pl.emm.upl.allshr.ucunc + pl.emm.upl.allshr.uc + pl.emm.upl.allshr.unc + 
  patchwork::plot_layout(ncol=1) +
  plot_annotation('Upland Shrub Cover', caption = 'All Shrub Species Pooled')

```



# Individual species

```{r}
cov.byspp %>% 
  tabyl(shrub_species) %>% 
  arrange(-n) %>% 
  gt()

```


## Upland Shrub Cover, _Purshia tridentata_ - Combined Winter Range
```{r}
# PUTR
cov.putr <- cov.byspp %>%
  filter(shrub_species == "PUTR")

#### summary tables
cov.putr %>% 
  group_by(time_class, site_type) %>% 
  summarytools::descr(perc_cover, stats = "common") %>% 
  summarytools::tb() %>%
  mutate_if(.predicate = is.numeric,.funs = round, digits = 2) %>% 
  gt() %>% 
  tab_header(title = "PUTR cover ~ site type + time_class")

```

### Modeling and Posterior Description
```{r, cache=TRUE}
## add cover in 0-1 range
cov.putr <- cov.putr %>%
  mutate(cover = perc_cover/100)
  
stmod_cov.putr <- stan_glmer(cover ~ time_class + (1 | site_id), 
                      data = cov.putr,
                      family = mgcv::betar,
                      prior_aux = exponential(2),
                      iter = 8000,
                      seed = 1234)

```

```{r}
# extract posteriors
posteriors.cov.putrubcov <- insight::get_parameters(stmod_cov.putr)

# tidy
posteriors.cov.putrubcov.tidy <- posteriors.cov.putrubcov %>%
  pivot_longer(cols = starts_with("time"), names_to = "parameter", values_to = "values")

```

**Prior summary**: cover ~ time_class + (1 | site_id)    
```{r}
prior_summary(stmod_cov.putr)
```

```{r}
fun.model.summary(stmod_cov.putr)
```

```{r}
posteriors.cov.putrubcov.tidy %>%
  ggplot() +
  geom_density(aes(fill = parameter, x = values),color="grey", alpha = .25) +
  # scale_fill_viridis(discrete = TRUE) +
  scale_fill_manual(values = colfunc2(ncol(posteriors.cov.putrubcov.tidy))) +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland, PUTR cover - Combined Winter Range")
```

```{r}
#### PP check plot
pp_check(stmod_cov.putr) +
  labs(x = "Value", y = "Density", caption = "upland cover ppc plot. PUTR") +
  theme_minimal()

# save plots
ggsave("./output/figures_202107_stats_upland/ppc_PUTR_putrubs_unc.png", dpi = 300, width = 4.75, height = 3.75)
# ggsave("./output/figures_202107_stats_upland/ppc_PUTR_putrubs.pdf", width = 4.75, height = 3.75)
```

```{r}
fun.sexit.gt(stmod_cov.putr$stanfit)

fun.sexit.gt(stmod_cov.putr$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/xPUTR_putr_stmod_cover1_posteriors.rtf")
```


```{r, echo=FALSE, eval=FALSE}
fun.desc.post1.gt.rope(stmod_cov.putr$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description - PUTR", subtitle = "Combined Core & Noncore Winter Range")

```

```{r, eval = FALSE}
## save
fun.desc.post1.gt(stmod_cov.putr$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description - PUTR", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/PUTR_stmod_cover1_posteriors.rtf")

```
### Effect Description  
#### Probability of Direction
```{r}
# Visualize the pd
#### PD - AC
p_direction(stmod_cov.putr) %>% plot() +
  theme_minimal() +
  scale_fill_manual(values = colfunc3(2)) +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland Shrub Cover Combined Winter Range") +
  scale_fill_manual(values = colfunc3(2))

# save plots
ggsave("./output/figures_202107_stats_upland/pd_putr_cov_unc.png", dpi = 300, width = 4.75, height = 3.75)

```


#### Region of Practical Equivalence (ROPE)
```{r}
rope.PUTRcov.wcwnc.gt <- rope(stmod_cov.putr$stanfit, ci=0.9) %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 2,
    suffixing = TRUE
  ) 

rope.PUTRcov.wcwnc.gt
# rope.PUTRcov.wcwnc.gt %>%
#   gt::gtsave(filename = "./output/tables/upland_cover/rope_putr_cov_wcwnc.rtf")
```

### Contrasts

EMM summaries with type = "response", the tests and confidence intervals
are done before back-transforming. The ratios estimated are ratios of
geometric means. A model with a log response is in fact a model for
relative effects of any of its linear predictors, and this
back-transformation to ratios goes hand-in-hand with that.

```{r, include=FALSE, echo=FALSE}
# library(emmeans)
# Note: EMMs for later factor levels are subtracted from those for earlier levels; if you want the comparisons to go in the other direction, use pairs(pigs.emm.s, reverse = TRUE). Also, in multi-factor situations, you may specify by factor(s) to perform the comparisons separately at the levels of those factors.

# Results are given on the log (not the response) scale.
emmeans(stmod_cov.putr, ~ time_class) %>% 
  pairs() %>% 
  plot(type="response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Combined Winter Range, All Shrub Cover")

ggsave("./output/figures_202107_stats_upland/emm_putr_cov_ucunc.png", width = 4.5, height = 3.55) # save plot

# table of contrasts
emmeans(stmod_cov.putr, ~ time_class) %>% 
  pairs() %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>%
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log scale")

# table of contrasts. Results are given on the response scale.
emmeans(stmod_cov.putr, ~ time_class) %>% 
  pairs(reverse = FALSE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale.")

```

## Upland Shrub Cover, _Purshia tridentata_ - Core Winter Range
```{r}
#### summary tables
cov.putr %>% 
  group_by(time_class, site_type) %>% 
  summarytools::descr(perc_cover, stats = "common") %>% 
  summarytools::tb() %>%
  mutate_if(.predicate = is.numeric,.funs = round, digits = 2) %>% 
  gt() %>% 
  tab_header(title = "Upland cover - site type + time_class")

```

### Modeling and Posterior Description
```{r, cache=TRUE}
cov.putr.uc <- cov.putr %>% 
  filter(site_type == "core range") # add 

## add cover in 0-1 range
cov.putr.uc <- cov.putr.uc %>%
  mutate(cover = perc_cover/100)
  
stmod_cov.putr.uc <- stan_glmer(cover ~ time_class + (1 | site_id), 
                      data = cov.putr,
                      family = mgcv::betar,
                      prior_aux = exponential(2),
                      iter = 8000,
                      seed = 1234)

```

```{r}
# extract posteriors
posteriors.cov.putrubcov.uc <- insight::get_parameters(stmod_cov.putr.uc)

# tidy
posteriors.cov.putrubcov.tidy <- posteriors.cov.putrubcov %>%
  pivot_longer(cols = starts_with("time"), names_to = "parameter", values_to = "values")

```

**Prior summary**: cover ~ time_class + (1 | site_id)    
```{r}
prior_summary(stmod_cov.putr.uc)
```

```{r}
fun.model.summary(stmod_cov.putr.uc)
```

```{r}
posteriors.cov.putrubcov.tidy %>%
  ggplot() +
  geom_density(aes(fill = parameter, x = values),color="grey", alpha = .25) +
  # scale_fill_viridis(discrete = TRUE) +
  scale_fill_manual(values = colfunc2(ncol(posteriors.cov.putrubcov.tidy))) +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland,PUTR - Combined Winter Range")
```

```{r}
#### PP check plot
pp_check(stmod_cov.putr.uc) +
  labs(x = "Value", y = "Density", caption = "upland cover ppc plot. PUTR") +
  theme_minimal()

# save plots
ggsave("./output/figures_202107_stats_upland/ppc_PUTR_putrubs_uc.png", dpi = 300, width = 4.75, height = 3.75)
# ggsave("./output/figures_202107_stats_upland/ppc_PUTR_putrubs.pdf", width = 4.75, height = 3.75)
```

```{r}
fun.sexit.gt(stmod_cov.putr.uc$stanfit)

fun.sexit.gt(stmod_cov.putr.uc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/xPUTR_stmod_cover1_posteriors.rtf")
```


```{r, echo=FALSE, eval=FALSE}
fun.desc.post1.gt.rope(stmod_cov.putr.uc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range")

```

```{r, eval = FALSE}
## save
fun.desc.post1.gt(stmod_cov.putr.uc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/RICE_stmod_cover1_posteriors.rtf")

```
### Effect Description  
#### Probability of Direction
```{r}
# Visualize the pd
#### PD - AC
p_direction(stmod_cov.putr.uc) %>% plot() +
  theme_minimal() +
  scale_fill_manual(values = colfunc3(2)) +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland Shrub Cover Combined Winter Range") +
  scale_fill_manual(values = colfunc3(2))

# save plots
ggsave("./output/figures_202107_stats_upland/pd_upl_uc.png", dpi = 300, width = 4.75, height = 3.75)

```


#### Region of Practical Equivalence (ROPE)
```{r}
rope.PUTRcov.wcwnc.gt <- rope(stmod_cov.putr.uc$stanfit, ci=0.9) %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 2,
    suffixing = TRUE
  ) 

rope.PUTRcov.wcwnc.gt
# rope.PUTRcov.wcwnc.gt %>%
#   gt::gtsave(filename = "./output/tables/upland_cover/rope_PUTRcov_wcwnc.rtf")
```

### Contrasts

EMM summaries with type = "response", the tests and confidence intervals
are done before back-transforming. The ratios estimated are ratios of
geometric means. A model with a log response is in fact a model for
relative effects of any of its linear predictors, and this
back-transformation to ratios goes hand-in-hand with that.

```{r, include=FALSE, echo=FALSE}
# library(emmeans)
# Note: EMMs for later factor levels are subtracted from those for earlier levels; if you want the comparisons to go in the other direction, use pairs(pigs.emm.s, reverse = TRUE). Also, in multi-factor situations, you may specify by factor(s) to perform the comparisons separately at the levels of those factors.

# Results are given on the log (not the response) scale.
emmeans(stmod_cov.putr.uc, ~ time_class) %>% 
  pairs() %>% 
  plot(type="response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Combined Winter Range,PUTR")

# ggsave("./output/figures_202107_stats_upland/emm_PUTRcov_wcwnc.pdf", width = 4.5, height = 3.55) # save plot
ggsave("./output/figures_202107_stats_upland/emm_PUTRcov_uc.png", width = 4.5, height = 3.55) # save plot

# table of contrasts
emmeans(stmod_cov.putr.uc, ~ time_class) %>% 
  pairs() %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>%
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log scale")

# table of contrasts. Results are given on the response scale.
emmeans(stmod_cov.putr.uc, ~ time_class) %>% 
  pairs(reverse = FALSE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale.")

```
## Upland Shrub Cover, _Purshia tridentata_ - Noncore Winter Range
### Modeling and Posterior Description
```{r, cache=TRUE}
cov.putr.unc <- cov.putr %>% 
  filter(site_type == "non-core range") # add 

## add cover in 0-1 range
cov.putr.unc <- cov.putr.unc %>%
  mutate(cover = perc_cover/100)
  
stmod_cov.putr.unc <- stan_glmer(cover ~ time_class + (1 | site_id), 
                      data = cov.putr,
                      family = mgcv::betar,
                      prior_aux = exponential(2),
                      iter = 8000,
                      seed = 1234)

```

```{r}
# extract posteriors
posteriors.cov.putrubcov.unc <- insight::get_parameters(stmod_cov.putr.unc)

# tidy
posteriors.cov.putrubcov.tidy <- posteriors.cov.putrubcov %>%
  pivot_longer(cols = starts_with("time"), names_to = "parameter", values_to = "values")

```

**Prior summary**: cover ~ time_class + (1 | site_id)    
```{r}
prior_summary(stmod_cov.putr.unc)
```

```{r}
fun.model.summary(stmod_cov.putr.unc)
```

```{r}
posteriors.cov.putrubcov.tidy %>%
  ggplot() +
  geom_density(aes(fill = parameter, x = values),color="grey", alpha = .25) +
  # scale_fill_viridis(discrete = TRUE) +
  scale_fill_manual(values = colfunc2(ncol(posteriors.cov.putrubcov.tidy))) +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland,PUTR - Combined Winter Range")
```

```{r}
#### PP check plot
pp_check(stmod_cov.putr.unc) +
  labs(x = "Value", y = "Density", caption = "upland cover ppc plot. PUTR") +
  theme_minimal()

# save plots
ggsave("./output/figures_202107_stats_upland/ppc_PUTR_putrubs_unc.png", dpi = 300, width = 4.75, height = 3.75)
# ggsave("./output/figures_202107_stats_upland/ppc_PUTR_putrubs.pdf", width = 4.75, height = 3.75)
```

```{r}
fun.sexit.gt(stmod_cov.putr.unc$stanfit)

fun.sexit.gt(stmod_cov.putr.unc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/xRICE_stmod_cover1_posteriors.rtf")
```


```{r, echo=FALSE, eval=FALSE}
fun.desc.post1.gt.rope(stmod_cov.putr.unc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range")

```

```{r, eval = FALSE}
## save
fun.desc.post1.gt(stmod_cov.putr.unc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/RICE_stmod_cover1_posteriors.rtf")

```
### Effect Description  
#### Probability of Direction
```{r}
# Visualize the pd
#### PD - AC
p_direction(stmod_cov.putr.unc) %>% plot() +
  theme_minimal() +
  scale_fill_manual(values = colfunc3(2)) +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland Shrub Cover Combined Winter Range") +
  scale_fill_manual(values = colfunc3(2))

# save plots
ggsave("./output/figures_202107_stats_upland/pd_PUTRcov_unc.png", dpi = 300, width = 4.75, height = 3.75)

```


#### Region of Practical Equivalence (ROPE)
```{r}
rope.PUTRcov.wcwnc.gt <- rope(stmod_cov.putr.unc$stanfit, ci=0.9) %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 2,
    suffixing = TRUE
  ) 

rope.PUTRcov.wcwnc.gt
# rope.PUTRcov.wcwnc.gt %>%
#   gt::gtsave(filename = "./output/tables/upland_cover/rope_RICEcov_wcwnc.rtf")
```

### Contrasts

EMM summaries with type = "response", the tests and confidence intervals
are done before back-transforming. The ratios estimated are ratios of
geometric means. A model with a log response is in fact a model for
relative effects of any of its linear predictors, and this
back-transformation to ratios goes hand-in-hand with that.

```{r, include=FALSE, echo=FALSE}
# library(emmeans)
# Note: EMMs for later factor levels are subtracted from those for earlier levels; if you want the comparisons to go in the other direction, use pairs(pigs.emm.s, reverse = TRUE). Also, in multi-factor situations, you may specify by factor(s) to perform the comparisons separately at the levels of those factors.

# Results are given on the log (not the response) scale.
emmeans(stmod_cov.putr.unc, ~ time_class) %>% 
  pairs() %>% 
  plot(type="response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Combined Winter Range,PUTR")

# ggsave("./output/figures_202107_stats_upland/emm_PUTRcov_unc.pdf", width = 4.5, height = 3.55) # save plot
ggsave("./output/figures_202107_stats_upland/emm_PUTRcov_unc.png", width = 4.5, height = 3.55) # save plot

# table of contrasts
emmeans(stmod_cov.putr.unc, ~ time_class) %>% 
  pairs() %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>%
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log scale")

# table of contrasts. Results are given on the response scale.
emmeans(stmod_cov.putr.unc, ~ time_class) %>% 
  pairs(reverse = FALSE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale.")

```

## Upland Shrub Cover, _Ribes cereum_ - Combined Winter Range
```{r}
# PUTR
cov.rice <- cov.byspp %>%
  filter(shrub_species == "RICE")

#### summary tables
cov.rice %>% 
  group_by(time_class, site_type) %>% 
  summarytools::descr(perc_cover, stats = "common") %>% 
  summarytools::tb() %>%
  mutate_if(.predicate = is.numeric,.funs = round, digits = 2) %>% 
  gt() %>% 
  tab_header(title = "Upland cover ~ site type + time_class")

```

### Modeling and Posterior Description
```{r, cache=TRUE}
## add cover in 0-1 range
cov.rice <- cov.rice %>%
  mutate(cover = perc_cover/100)
  
stmod_cov.rice <- stan_glmer(cover ~ time_class + (1 | site_id), 
                      data = cov.rice,
                      family = mgcv::betar,
                      prior_aux = exponential(2),
                      iter = 8000,
                      seed = 1234)

```

```{r}
# extract posteriors
posteriors.cov.riceubcov <- insight::get_parameters(stmod_cov.rice)

# tidy
posteriors.cov.riceubcov.tidy <- posteriors.cov.riceubcov %>%
  pivot_longer(cols = starts_with("time"), names_to = "parameter", values_to = "values")

```

**Prior summary**: cover ~ time_class + (1 | site_id)    
```{r}
prior_summary(stmod_cov.rice)
```

```{r}
fun.model.summary(stmod_cov.rice)
```

```{r}
posteriors.cov.riceubcov.tidy %>%
  ggplot() +
  geom_density(aes(fill = parameter, x = values),color="grey", alpha = .25) +
  # scale_fill_viridis(discrete = TRUE) +
  scale_fill_manual(values = colfunc2(ncol(posteriors.cov.riceubcov.tidy))) +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland,RICE - Combined Winter Range")
```

```{r}
#### PP check plot
pp_check(stmod_cov.rice) +
  labs(x = "Value", y = "Density", caption = "upland cover ppc plot. RICE") +
  theme_minimal()

# save plots
ggsave("./output/figures_202107_stats_upland/ppc_PUTR_riceubs_unc.png", dpi = 300, width = 4.75, height = 3.75)
# ggsave("./output/figures_202107_stats_upland/ppc_PUTR_riceubs.pdf", width = 4.75, height = 3.75)
```

```{r}
fun.sexit.gt(stmod_cov.rice$stanfit)

fun.sexit.gt(stmod_cov.rice$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/xPUTR_riceubs_stmod_cover1_posteriors.rtf")
```


```{r, echo=FALSE, eval=FALSE}
fun.desc.post1.gt.rope(stmod_cov.rice$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range")

```

```{r, eval = FALSE}
## save
fun.desc.post1.gt(stmod_cov.rice$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/rice_cov_stmod_cover1_posteriors.rtf")

```
### Effect Description  
#### Probability of Direction
```{r}
# Visualize the pd
#### PD - AC
p_direction(stmod_cov.rice) %>% plot() +
  theme_minimal() +
  scale_fill_manual(values = colfunc3(2)) +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland Shrub Cover Combined Winter Range") +
  scale_fill_manual(values = colfunc3(2))

# save plots
ggsave("./output/figures_202107_stats_upland/pd_RICEcov_ucunc.png", dpi = 300, width = 4.75, height = 3.75)

```


#### Region of Practical Equivalence (ROPE)
```{r}
rope.RICEcov.wcwnc.gt <- rope(stmod_cov.rice$stanfit, ci=0.9) %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 2,
    suffixing = TRUE
  ) 

rope.RICEcov.wcwnc.gt
# rope.RICEcov.wcwnc.gt %>%
#   gt::gtsave(filename = "./output/tables/upland_cover/rope_RICEcov_wcwnc.rtf")
```

### Contrasts

EMM summaries with type = "response", the tests and confidence intervals
are done before back-transforming. The ratios estimated are ratios of
geometric means. A model with a log response is in fact a model for
relative effects of any of its linear predictors, and this
back-transformation to ratios goes hand-in-hand with that.

```{r, include=FALSE, echo=FALSE}
# library(emmeans)
# Note: EMMs for later factor levels are subtracted from those for earlier levels; if you want the comparisons to go in the other direction, use pairs(pigs.emm.s, reverse = TRUE). Also, in multi-factor situations, you may specify by factor(s) to perform the comparisons separately at the levels of those factors.

# Results are given on the log (not the response) scale.
emmeans(stmod_cov.rice, ~ time_class) %>% 
  pairs() %>% 
  plot(type="response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Combined Winter Range,RICE")

# ggsave("./output/figures_202107_stats_upland/emm_RICEcov_wcwnc.pdf", width = 4.5, height = 3.55) # save plot
ggsave("./output/figures_202107_stats_upland/emm_RICEcov_wcwnc_unc.png", width = 4.5, height = 3.55) # save plot

# table of contrasts
emmeans(stmod_cov.rice, ~ time_class) %>% 
  pairs() %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>%
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log scale")

# table of contrasts. Results are given on the response scale.
emmeans(stmod_cov.rice, ~ time_class) %>% 
  pairs(reverse = FALSE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale.")

```

## Upland Shrub Cover, _Ribes cereum_ - Core Winter Range
```{r}
#### summary tables
cov.rice %>% 
  group_by(time_class, site_type) %>% 
  summarytools::descr(perc_cover, stats = "common") %>% 
  summarytools::tb() %>%
  mutate_if(.predicate = is.numeric,.funs = round, digits = 2) %>% 
  gt() %>% 
  tab_header(title = "Upland cover - site type + time_class")

```

### Modeling and Posterior Description
```{r, cache=TRUE}
cov.rice.uc <- cov.rice %>% 
  filter(site_type == "core range") # add 

## add cover in 0-1 range
cov.rice.uc <- cov.rice.uc %>%
  mutate(cover = perc_cover/100)
  
stmod_cov.rice.uc <- stan_glmer(cover ~ time_class + (1 | site_id), 
                      data = cov.rice,
                      family = mgcv::betar,
                      prior_aux = exponential(2),
                      iter = 8000,
                      seed = 1234)

```

```{r}
# extract posteriors
posteriors.cov.riceubcov.uc <- insight::get_parameters(stmod_cov.rice.uc)

# tidy
posteriors.cov.riceubcov.tidy <- posteriors.cov.riceubcov %>%
  pivot_longer(cols = starts_with("time"), names_to = "parameter", values_to = "values")

```

**Prior summary**: cover ~ time_class + (1 | site_id)    
```{r}
prior_summary(stmod_cov.rice.uc)
```

```{r}
fun.model.summary(stmod_cov.rice.uc)
```

```{r}
posteriors.cov.riceubcov.tidy %>%
  ggplot() +
  geom_density(aes(fill = parameter, x = values),color="grey", alpha = .25) +
  # scale_fill_viridis(discrete = TRUE) +
  scale_fill_manual(values = colfunc2(ncol(posteriors.cov.riceubcov.tidy))) +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland,RICE - Combined Winter Range")
```

```{r}
#### PP check plot
pp_check(stmod_cov.rice.uc) +
  labs(x = "Value", y = "Density", caption = "upland cover ppc plot. RICE") +
  theme_minimal()

# save plots
ggsave("./output/figures_202107_stats_upland/ppc_RICE_riceubs_uc.png", dpi = 300, width = 4.75, height = 3.75)
# ggsave("./output/figures_202107_stats_upland/ppc_RICE_riceubs.pdf", width = 4.75, height = 3.75)
```

```{r}
fun.sexit.gt(stmod_cov.rice.uc$stanfit)

fun.sexit.gt(stmod_cov.rice.uc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/RICE_riceubs_stmod_cover1_posteriors.rtf")
```


```{r, echo=FALSE, eval=FALSE}
fun.desc.post1.gt.rope(stmod_cov.rice.uc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range")

```

```{r, eval = FALSE}
## save
fun.desc.post1.gt(stmod_cov.rice.uc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/RICE_riceubs_stmod_cover1_posteriors.rtf")

```
### Effect Description  
#### Probability of Direction
```{r}
# Visualize the pd
#### PD - AC
p_direction(stmod_cov.rice.uc) %>% plot() +
  theme_minimal() +
  scale_fill_manual(values = colfunc3(2)) +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland Shrub Cover Combined Winter Range") +
  scale_fill_manual(values = colfunc3(2))

# save plots
ggsave("./output/figures_202107_stats_upland/pd_upl_uc.png", dpi = 300, width = 4.75, height = 3.75)

```


#### Region of Practical Equivalence (ROPE)
```{r}
rope.RICEcov.wcwnc.gt <- rope(stmod_cov.rice.uc$stanfit, ci=0.9) %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 2,
    suffixing = TRUE
  ) 

rope.RICEcov.wcwnc.gt
# rope.RICEcov.wcwnc.gt %>%
#   gt::gtsave(filename = "./output/tables/upland_cover/rope_RICEcov_wcwnc.rtf")
```

### Contrasts

EMM summaries with type = "response", the tests and confidence intervals
are done before back-transforming. The ratios estimated are ratios of
geometric means. A model with a log response is in fact a model for
relative effects of any of its linear predictors, and this
back-transformation to ratios goes hand-in-hand with that.

```{r, include=FALSE, echo=FALSE}
# library(emmeans)
# Note: EMMs for later factor levels are subtracted from those for earlier levels; if you want the comparisons to go in the other direction, use pairs(pigs.emm.s, reverse = TRUE). Also, in multi-factor situations, you may specify by factor(s) to perform the comparisons separately at the levels of those factors.

# Results are given on the log (not the response) scale.
emmeans(stmod_cov.rice.uc, ~ time_class) %>% 
  pairs() %>% 
  plot(type="response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Combined Winter Range,RICE")

# ggsave("./output/figures_202107_stats_upland/emm_RICEcov_wcwnc.pdf", width = 4.5, height = 3.55) # save plot
ggsave("./output/figures_202107_stats_upland/emm_RICEcov_uc.png", width = 4.5, height = 3.55) # save plot

# table of contrasts
emmeans(stmod_cov.rice.uc, ~ time_class) %>% 
  pairs() %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>%
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log scale")

# table of contrasts. Results are given on the response scale.
emmeans(stmod_cov.rice.uc, ~ time_class) %>% 
  pairs(reverse = FALSE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale.")

```
## Upland Shrub Cover, _Ribes cereum_ - Noncore Winter Range
### Modeling and Posterior Description
```{r, cache=TRUE}
cov.rice.unc <- cov.rice %>% 
  filter(site_type == "non-core range") # add 

## add cover in 0-1 range
cov.rice.unc <- cov.rice.unc %>%
  mutate(cover = perc_cover/100)
  
stmod_cov.rice.unc <- stan_glmer(cover ~ time_class + (1 | site_id), 
                      data = cov.rice.unc,
                      family = mgcv::betar,
                      prior_aux = exponential(2),
                      iter = 8000,
                      seed = 1234)

```

```{r}
# extract posteriors
posteriors.cov.riceubcov.unc <- insight::get_parameters(stmod_cov.rice.unc)

# tidy
posteriors.cov.riceubcov.tidy <- posteriors.cov.riceubcov %>%
  pivot_longer(cols = starts_with("time"), names_to = "parameter", values_to = "values")

```

**Prior summary**: cover ~ time_class + (1 | site_id)    
```{r}
prior_summary(stmod_cov.rice.unc)
```

```{r}
fun.model.summary(stmod_cov.rice.unc)
```

```{r}
posteriors.cov.riceubcov.tidy %>%
  ggplot() +
  geom_density(aes(fill = parameter, x = values),color="grey", alpha = .25) +
  # scale_fill_viridis(discrete = TRUE) +
  scale_fill_manual(values = colfunc2(ncol(posteriors.cov.riceubcov.tidy))) +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland,RICE - Combined Winter Range")
```

```{r}
#### PP check plot
pp_check(stmod_cov.rice.unc) +
  labs(x = "Value", y = "Density", caption = "upland cover ppc plot. RICE") +
  theme_minimal()

# save plots
ggsave("./output/figures_202107_stats_upland/ppc_RICE_riceubs_unc.png", dpi = 300, width = 4.75, height = 3.75)
# ggsave("./output/figures_202107_stats_upland/ppc_RICE_riceubs.pdf", width = 4.75, height = 3.75)
```

```{r}
fun.sexit.gt(stmod_cov.rice.unc$stanfit)

fun.sexit.gt(stmod_cov.rice.unc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/xRICE_riceubs_stmod_cover1_posteriors.rtf")
```


```{r, echo=FALSE, eval=FALSE}
fun.desc.post1.gt.rope(stmod_cov.rice.unc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range")

```

```{r, eval = FALSE}
## save
fun.desc.post1.gt(stmod_cov.rice.unc$stanfit) %>%
  tab_header(title = "Upland Shrub Cover Posterior Description", subtitle = "Combined Core & Noncore Winter Range") %>% 
  gt::gtsave(filename = "./output/tables/upland_cover/RICE_riceubs_stmod_cover1_posteriors.rtf")

```
### Effect Description  
#### Probability of Direction
```{r}
# Visualize the pd
#### PD - AC
p_direction(stmod_cov.rice.unc) %>% plot() +
  theme_minimal() +
  scale_fill_manual(values = colfunc3(2)) +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Upland Shrub Cover Combined Winter Range") +
  scale_fill_manual(values = colfunc3(2))

# save plots
ggsave("./output/figures_202107_stats_upland/pd_RICEcov_unc.png", dpi = 300, width = 4.75, height = 3.75)

```


#### Region of Practical Equivalence (ROPE)
```{r}
rope.RICEcov.wcwnc.gt <- rope(stmod_cov.rice.unc$stanfit, ci=0.9) %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 2,
    suffixing = TRUE
  ) 

rope.RICEcov.wcwnc.gt
# rope.RICEcov.wcwnc.gt %>%
#   gt::gtsave(filename = "./output/tables/upland_cover/rope_RICEcov_wcwnc.rtf")
```

### Contrasts

EMM summaries with type = "response", the tests and confidence intervals
are done before back-transforming. The ratios estimated are ratios of
geometric means. A model with a log response is in fact a model for
relative effects of any of its linear predictors, and this
back-transformation to ratios goes hand-in-hand with that.

```{r, include=FALSE, echo=FALSE}
# library(emmeans)
# Note: EMMs for later factor levels are subtracted from those for earlier levels; if you want the comparisons to go in the other direction, use pairs(pigs.emm.s, reverse = TRUE). Also, in multi-factor situations, you may specify by factor(s) to perform the comparisons separately at the levels of those factors.

# Results are given on the log (not the response) scale.
emmeans(stmod_cov.rice.unc, ~ time_class) %>% 
  pairs() %>% 
  plot(type="response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Combined Winter Range,RICE")

# ggsave("./output/figures_202107_stats_upland/emm_RICEcov_unc.pdf", width = 4.5, height = 3.55) # save plot
ggsave("./output/figures_202107_stats_upland/emm_RICEcov_unc.png", width = 4.5, height = 3.55) # save plot

# table of contrasts
emmeans(stmod_cov.rice.unc, ~ time_class) %>% 
  pairs() %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>%
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log scale")

# table of contrasts. Results are given on the response scale.
emmeans(stmod_cov.rice.unc, ~ time_class) %>% 
  pairs(reverse = FALSE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale.")

```

# Additional plots
```{r}
# Point estimate displayed: median 
# HPD interval probability: 0.95

pl.emm.cov.putr.ucunc <- emmeans(stmod_cov.putr, ~ time_class) %>% 
  pairs() %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Combined Range", x = "Estimate", y = "Contrast")

pl.emm.cov.putr.uc <- emmeans(stmod_cov.putr.uc, ~ time_class) %>% 
  pairs() %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Core Range", x = "Estimate", y = "Contrast")

pl.emm.cov.putr.unc <- emmeans(stmod_cov.putr.unc, ~ time_class) %>% 
  pairs() %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Noncore Range", x = "Estimate", y = "Contrast")


pl.emm.cov.putr.ucunc + pl.emm.cov.putr.uc + pl.emm.cov.putr.unc + 
  patchwork::plot_layout(ncol=1) +
  plot_annotation(title = 'Upland Shrub Cover', subtitle = expression(paste(italic("Pushia tridentata"))))
```
# here
```{r, eval=FALSE}
# Point estimate displayed: median 
# HPD interval probability: 0.95

pl.emm.cov.rice.ucunc <- emmeans(stmod_cov.rice, ~ time_class) %>% 
  pairs() %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Combined Range", x = "Estimate", y = "Contrast")

pl.emm.cov.rice.uc <- emmeans(stmod_cov.rice.uc, ~ time_class) %>% 
  pairs() %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Core Range", x = "Estimate", y = "Contrast")

pl.emm.cov.rice.unc <- emmeans(stmod_cov.rice.unc, ~ time_class) %>% 
  pairs() %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 1), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Noncore Range", x = "Estimate", y = "Contrast")


pl.emm.cov.rice.ucunc + pl.emm.cov.rice.uc + pl.emm.cov.rice.unc + 
  patchwork::plot_layout(ncol=1) +
  plot_annotation(title = 'Upland Shrub Cover', subtitle = expression(paste(italic("Pushia tridentata"))))
```

**Combined range**
```{r}
pl.emm.upl.allshr + pl.emm.cov.putr.ucunc + pl.emm.cov.rice.ucunc + 
  patchwork::plot_layout(ncol=1) +
  plot_annotation(title = 'Upland Shrub Cover', subtitle = 'Combined Range', tag_levels = "A")

ggsave("./output/figures_202107_stats_upland/uplcov_3panel_ucunc.png", dpi = 300, width = 7.75, height = 3.75)
```

**Core range**
```{r}
(pl.emm.upl.allshr.uc + labs(subtitle="")) + pl.emm.cov.putr.uc + pl.emm.cov.rice.uc + 
  patchwork::plot_layout(ncol=1) +
  plot_annotation(title = 'Upland Shrub Cover', subtitle = 'Core Range',  tag_levels = "A")

ggsave("./output/figures_202107_stats_upland/uplcov_3panel_uc.png", dpi = 300, width = 7.75, height = 3.75)
```

**Noncore range**
```{r}
pl.emm.upl.allshr.ucunc + pl.emm.cov.putr.ucunc + pl.emm.cov.rice.ucunc + 
  patchwork::plot_layout(ncol=1) +
  plot_annotation(title = 'Upland Shrub Cover', subtitle = 'Noncore Range',  tag_levels = "A")

ggsave("./output/figures_202107_stats_upland/uplcov_3panel_unc.png", dpi = 300, width = 7.75, height = 3.75)
```

# Session info

R version 4.0.3 (2020-10-10)  
Platform: x86_64-w64-mingw32/x64 (64-bit)  
Running under: Windows 10 x64 (build 19041)  
Last updated: `r format(Sys.time(), '%Y %B %d')` 

```{r, echo=FALSE}
subset(data.frame(sessioninfo::package_info()), attached==TRUE, c(package, loadedversion)) %>% 
  gt() %>% 
  tab_header(title = "Attached packages and versions")

```

```{r, echo=FALSE, eval=FALSE}
citation(package = "emmeans")
citation(package = "rstanarm")
citation(package = "bayestestR")
citation(package = "Tidyverse")
citation(package = "Bayesplot")
```


