---
title: "Appendix E. EVMP Willow Statistical Analyses"
author: ""
date: ""
output: 
  html_document: 
    theme: journal
    toc: yes
    toc_depth: 2
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

**Willow Macroplot Statistical Analyses**

*Repository:*\
<https://github.com/egage/EVMP>

```{r setup, include=FALSE}
# Rmd setup
library(knitr)
opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = 'center')
opts_knit$set(root.dir=normalizePath('../')) # required if Rmd is nested below the project directory
opts_chunk$set(fig.path = "../output/figures/") 
# set seed
set.seed(1234)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=FALSE, eval=TRUE}
# Read in libraries
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(janitor))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(lubridate))
library(emmeans)
library(rstanarm)
# rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
library(bayestestR)
library(gt)
library(ggridges)
library(bayesplot)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=FALSE, eval=TRUE}
## Functions
### Table functions
## function to describe model out
fun.desc.post1 <- function(model){
  describe_posterior(
  model,
  effects = "fixed",
  component = "all",
  ci_method = "hdi",
  test = c("p_direction", "p_significance"),
  centrality = "all"
) %>% 
    as_tibble() %>% 
    DT::datatable(caption = "Posterior description")
}

## Functions for summarizing posterior distributions, creating tables with the gt
fun.sexit.gt <- function(model){
  bayestestR::sexit(model, ci=0.9) %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric), round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Effect Existence and Significance Testing") }


# describe posterior
fun.desc.post1.gt.rope <- function(model){
  describe_posterior(
  model,
  effects = "fixed",
  component = "all",
  ci_method = "hdi",
  ci=0.9,
  test = c("p_direction", "p_significance","rope"),
  #rope_range(c(-0.1,0.1)),
  rope_range("default"),
  rope_ci = 0.9,
  centrality = "median"
) %>% 
    as_tibble() %>% 
    gt() %>% 
    tab_header(title = "Posterior description") %>% 
    fmt_number(
    columns = 2:14,
    decimals = 2,
    suffixing = TRUE
  )
}

```

```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=FALSE, eval=TRUE}

#### Plotting functions
fun.contrast.plot <- function(df){
  df %>% 
  ggplot(aes(x = contrast)) +
  geom_linerange(aes(ymin = lower.HPD, ymax = upper.HPD), color = "lightblue", size = 2) +
  geom_point(aes(y = estimate), size = 2.5) +
  theme_minimal() +
  labs(caption = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Estimate") +
  coord_flip()
}

## ppcheck.plot
ppc.plot <- function(model){
  pp_check(model) +
  labs(x = "Value", y = "Density") +
  theme_minimal()
}

# session info: verbose
package_versions <- function(packages=NULL, base=FALSE, sort=FALSE) {
    #' Determine package versions.
    #' 
    #' @param packages Packages to query; if not specified (or \code{NULL}),
    #' all currently attached packages are queried.
    #' @param base If \code{TRUE}, include the base package (R version).
    #' @param sort If \code{TRUE}, sort package names alphabetically. 
    #' @return Data frame with package names and corresponding versions.
    if (is.null(packages))
        packages <- names(sessionInfo()$otherPkgs)
    if (base)
        packages <- c("base", packages)
    if (sort)
        packages <- sort(packages)
    if (!is.null(packages)) {
        tmp <- data.frame(sapply(sapply(packages, packageVersion, simplify=FALSE), as.character))
        colnames(tmp) <- "version"
        data.frame(package=rownames(tmp), version=tmp[1], row.names=NULL)
    }
}

```

# Introduction

This document contains code for running statistical analyses of of Elk
Vegetation Management Plan (EVMP) data collected through the 2018
sampling season. Analyses are limited to macroplot data from plots
established in riparian areas in areas of the elk winter range and
Kawuneeche Valley (i.e., "willow" plots, see Zeigenfuss et al. 2011). Analyses are provided as a supplement to the NPS NRR Report "Monitoring of
Vegetation Response to Elk Population and Habitat Management in Rocky
Mountain National Park - Analysis of Elk Vegetation Management Plan
monitoring Data: 2008--2018". For information on sampling and the
broader analysis and interpretation of the results, refer to the project
report, past analyses [@zeigenfuss2015], and the original EVMP
monitoring plan [@zeigenfuss2011].

# Methods

The code and results presented here build on derived
data produced in a separate code document focused on ingesting,
compiling, and cleaning raw data provided by RMNP staff (Appendix B). Extensive
"munging" of raw files was required, so the decision was made to
separate code used to pre-process raw data from code used for the
analyses presented here. Likewise, separate code documents were
developed for other distinct elements of the larger EVMP (e.g., aspen
and upland plots). In general, packages included in the "Tidyverse"
ecosystem of R packages were used for data transformation and
visualization [@wickham2019], although specialized packages particular
to specific tasks (e.g., the "bayesplot" plot for visualization of
Bayesian posterior distributions) were also used [@gabry2019].

Bayesian repeated measures analyses were fit separately for combined
core and non-core winter range, core winter range (WC) and non-core
winter range plots (WNC). For willow height, models were fit using
weekly informative priors and a gamma distribution with the "stan_lmer"
function in the "rstanarm" package [@goodrich2020][@brilleman2018].

For continuous proportions such as plant cover, a common approach to modeling is to transform data
then use ordinary linear models, but this creates problems for
interpretation and inference [@douma2019]. Techniques relying on
transformations make estimates on the transformed scale, requiring
back‐transformation for reporting and interpretation. However, the
relationship between the original and transformed proportions is often
non‐linear, creating issues for interpretation [@douma2019]. Alternative
approaches using beta regression [@cribari-neto2009][@ferrari2004] have
been developed and are relied on here for modeling continuous
proportions (e.g., willow cover). After fitting appropriate models using
"rstanarm" functions, comparisons of estimated marginal means were made
between time classes and management classes (e.g., fencing) using
functions in the "emm" R package [@lenth2020].

The Probability of Direction (PD) was used to represent certainty associated with the most probable direction (positive or negative) of the effect (e.g., time class, fencing). The PD is correlated with the frequentist p-value, with a two-sided p-value of respectively .1, .05, and .01 approximated by a PD of 95%, 97.5%, and 99.5% [@makowski2019]. the “region of practical equivalence” (ROPE) to evaluate the probability of a parameter being outside a range that can be considered as “practically no effect”, i.e., a region enclosing values that are equivalent to the null value for practical purposes [@kruschke2018], with the proportion of the 95% HDI credible interval falling within the ROPE was used as an index for an analog to frequentist "null-hypothesis" testing.

```{r}
#### Data import and munging
# read in willow. These data are the output of extensive munging accomplished using code in "EVMP_munging_20210131.Rmd"

willow <- read_csv("./output/exported_data/willow_mcro.csv")

willow <- willow %>% 
  select(-c(contains("UTM"))) 

## create timeClass var
willow.ht <- willow %>% 
  filter(!is.na(PLANT_HT_CM)) %>% 
  mutate(timeClass = case_when(yr == 2008 ~ "BL",
                              yr == 2009 ~ "BL", 
                              TRUE ~ as.character(yr))) 

## filter to 5 yr sampling interval, chr to factor conversions
willow.ht <- willow.ht %>% 
  filter(timeClass == "BL" | timeClass == "2013" | timeClass == "2018") %>%
  mutate(timeClass = as.factor(timeClass)) %>% 
  mutate(timeClass = fct_relevel(timeClass, "2018", "2013", "BL")) %>% 
  mutate(FENCED = as.factor(FENCED))

## clean names
willow.ht <- willow.ht %>% 
  janitor::clean_names()

## filter out just the willow species
willow.ht <- willow.ht %>%
  filter(str_detect(species_code, pattern = "SA"))

## filter out 0
willow.ht <- willow.ht %>% 
  filter(plant_ht_cm > 0)

## reorder the time class
willow.ht <- willow.ht %>% 
  mutate(time_class = fct_rev(time_class))

## factor reverse: fenced
willow.ht <- willow.ht %>%
  mutate(fenced = fct_rev(fenced))

```

```{r}
## create separate df for WC, WNC, combined WC & WNC, and KV
willow.ht.wc <- willow.ht %>% 
  filter(site_type == "WC") 

# non core
willow.ht.nc <- willow.ht %>% 
  # distinct(site_type)
  filter(site_type == "WNC") 

# combined WC and WNC
willow.ht.wc.wnc <- willow.ht %>% 
  # distinct(site_type)
  filter(site_type == "WNC" | site_type == "WC") 

# kw (note: different munging workflow needed b/c of different sampling yrs)
willow.ht.wk <- willow %>%
  clean_names() %>%
  filter(str_detect(species_code, pattern = "SA")) %>% 
  filter(site_type == "WK") %>% 
  mutate(fenced = as.factor(fenced)) %>% 
  filter(plant_ht_cm > 0) %>% 
  mutate(time_class = as.factor(time_class)) %>% 
  mutate(fenced = fct_rev(fenced))

```

# Results

## Willow Height - Combined Core and Noncore Winter Range Macroplot Data

### Histograms of Willow Height

```{r}
## histogram
willow.ht.wc.wnc %>% 
  ggplot() +
  geom_histogram(aes(plant_ht_cm, fill = fenced), color = "black") +
  viridis::scale_fill_viridis(discrete=TRUE) +
  labs(x="Plant height (cm)", y = "Count", fill = "", caption = "Core and noncore winter range plots. Macroplot heights") +
  theme_minimal() +
  facet_wrap(~time_class) +
  coord_flip()

# ggsave("./output/figures_exported/willow_height/hist_willowht_TC_wcwnc.png", dpi = 300, width = 6.75, height = 3.75)
# ggsave("./output/figures_exported/willow_height/hist_willowht_TC_wcwnc.pdf", width = 6.75, height = 3.75)

```

### Modeling and Posterior Description
```{r, cache=TRUE}
## run with gamma
wc.wnc.stmod_TCxF <- stan_glmer(plant_ht_cm ~ time_class*fenced +  (1 | site_id), data = willow.ht.wc.wnc,
                      family=Gamma(link="log"),
                      iter = 8000,
                      )
```

**Prior summary**: plant_ht_cm ~ time_class*fenced +  (1 | site_id)
```{r, echo=TRUE}
# Get the prior summary
prior_summary(wc.wnc.stmod_TCxF)
```

```{r}
#### PP check plot
ppc.plot(wc.wnc.stmod_TCxF) +
  labs(caption = "WC+WNC: willow_ht~TCxF")
# ggsave("./output/figures_exported/willow_height/ppc_willowht_wcwnc.png", dpi = 300, width = 4.75, height = 3.75)
# ggsave("./output/figures_exported/willow_height/ppc_willowht_wcwnc.pdf", width = 4.75, height = 3.75)
```

```{r}
# Extract posteriors
posteriors.wc.wnc.ht <- insight::get_parameters(wc.wnc.stmod_TCxF)

## tidy
posteriors.wc.wnc.ht.tidy <- posteriors.wc.wnc.ht %>%
  pivot_longer(cols = starts_with("time"), names_to = "parameter", values_to = "values")

posteriors.wc.wnc.ht.tidy %>%
  ggplot() +
  geom_density(aes(fill = parameter, x = values),color="grey", alpha = .25) +
  scale_fill_viridis(discrete = TRUE) +
  theme_minimal()
```

### Effect Existence and Significance Testing
```{r}
# use function defined above
fun.sexit.gt(wc.wnc.stmod_TCxF)

## save as RTF
fun.sexit.gt(wc.wnc.stmod_TCxF) %>%
 tab_header(title = "Combined WC and WNC", subtitle = "Willow Height Posterior Description") %>%
 gt::gtsave(filename = "./output/tables/willow_height/postdesc_willowht_TCxF_wcwnc.rtf")

```

```{r, echo=FALSE, eval=FALSE}
#### Describe posterior distributions
fun.desc.post1.gt.rope(wc.wnc.stmod_TCxF) %>% 
  tab_header(title = "Combined WC and WNC", subtitle = "Willow Height Posterior Description")

## save as RTF
fun.desc.post1.gt.rope(wc.wnc.stmod_TCxF) %>%
 tab_header(title = "Combined WC and WNC", subtitle = "Willow Height Posterior Description") %>%
  gt::gtsave(filename = "./output/tables/willow_height/postdesc_willowht_TCxF_wcwnc.rtf")
```

#### Probability of Direction
```{r}
pd.wc.wnc <- p_direction(wc.wnc.stmod_TCxF)
# Visualize the pd
plot(pd.wc.wnc) +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Willow height, Combined Core and Non-core Winter Range")

# ggsave("./output/figures_exported/willow_height/pd_willowht_wcwnc.png", width = 6.75, height =3.75, dpi = 300)
# ggsave("./output/figures_exported/willow_height/pd_willowht_wcwnc.png", width = 6.75, height =3.75)

```

#### Region of Practical Equivalence (ROPE)

```{r}
rope.wcwnc <- rope(wc.wnc.stmod_TCxF, ci=0.9)

rope.wcwnc.gt <- rope.wcwnc %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 2,
    suffixing = TRUE
  ) 
rope.wcwnc.gt
# rope.wcwnc.gt %>%
#   gt::gtsave(filename = "./output/tables/willow_height/rope_willowht_wcwnc.rtf")
```

### Contrasts

EMM summaries with type = "response", the tests and confidence intervals
are done before back-transforming. The ratios estimated are ratios of
geometric means. A model with a log response is in fact a model for
relative effects of any of its linear predictors, and this
back-transformation to ratios goes hand-in-hand with that.

```{r, include=FALSE, echo=FALSE}
# library(emmeans)
# Note: EMMs for later factor levels are subtracted from those for earlier levels; if you want the comparisons to go in the other direction, use pairs(pigs.emm.s, reverse = TRUE). Also, in multi-factor situations, you may specify by factor(s) to perform the comparisons separately at the levels of those factors.

# Results are given on the log (not the response) scale.
# main effects
## main effect: time class
wc.wnc.emm.tc <- emmeans(wc.wnc.stmod_TCxF, ~ time_class)
wc.wnc.pairs.timeClass <- pairs(wc.wnc.emm.tc)

plot(wc.wnc.pairs.timeClass) +
  theme_minimal() +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Combined Core and Noncore Winter Range, Willow Height")

## main effect: fencing
wc.wnc.emm.fenced <- emmeans(wc.wnc.stmod_TCxF, ~ fenced)
wc.wnc.pairs.fenced <- pairs(wc.wnc.emm.fenced)

plot(wc.wnc.pairs.fenced) +
  theme_minimal() +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Combined Core and Noncore Winter Range, Willow Height")

## the above is not used b/c sig of interactions
```

```{r}
## Effect: TC x fencing
emmeans(wc.wnc.stmod_TCxF, ~ time_class | fenced) %>% 
  pairs(reverse=TRUE) 

emmeans(wc.wnc.stmod_TCxF, ~ time_class | fenced) %>% 
  pairs(reverse=TRUE) %>% 
  plot(type = "response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Combined Core and Noncore Winter Range, Willow Height")

ggsave("./output/figures_exported/willow_height/emm_willowht_TCxF_wcwnc.pdf", width = 4.5, height = 3.55) # save plot
ggsave("./output/figures_exported/willow_height/emm_willowht_TCxF_wcwnc.png", width = 4.5, height = 3.55) # save plot

# table of contrasts
# table of contrasts
emmeans(wc.wnc.stmod_TCxF, ~ time_class | fenced) %>% 
  pairs(reverse = TRUE) %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>%
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log scale")

# table of contrasts. Results are given on the response scale.
emmeans(wc.wnc.stmod_TCxF, ~ time_class | fenced) %>% 
  pairs(reverse = TRUE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale.")

```

### Central Posterior Interval Estimates

```{r}

bayesplot::mcmc_areas(posteriors.wc.wnc.ht,
           pars = c("time_class2013", "time_class2018", "time_class2013:fencedFenced","time_class2018:fencedFenced"),
           prob = 0.9) +
  theme_minimal() +
  labs(title = "Posterior distributions", subtitle = "with medians and 90% intervals", caption = "Willow height, WC&WNC")

# ggsave("./output/figures_exported/willow_height/mcmcarea_willowht_TCxF_wcwnc.png", dpi = 300, width = 6.75, height = 3.75)
# ggsave("./output/figures_exported/willow_height/mcmcarea_willowht_TCxF_wcwnc.pdf", width = 6.75, height = 3.75)

```

## Willow Height - Core Winter Range Macroplot Data

### Histograms of Willow Height

```{r}
willow.ht.wc %>% 
  mutate(time_class = fct_rev(time_class)) %>% 
  ggplot() +
  # geom_density(aes(plant_ht_cm, fill = fenced), color = "black") +
  geom_histogram(aes(plant_ht_cm, fill = fenced), color = "black") +
  viridis::scale_fill_viridis(discrete=TRUE) +
  labs(x="Plant height (cm)", y = "Count", fill = "", caption = "Core winter range plots. Macroplot heights") +
  theme_minimal() +
  facet_wrap(~time_class) +
  coord_flip()

# ggsave("./output/figures_exported/willow_height/hist_willowht_TCxF_wc.png", dpi = 300, width = 6.75, height = 3.75)
# ggsave("./output/figures_exported/willow_height/hist_willowht_TCxF_wc.pdf", width = 6.75, height = 3.75)

```

### Modeling and Posterior Description

```{r, eval = FALSE, cache=TRUE}
#### null model
# Not currently run; run in past as model comparison  
# Previously run for comparisons but not otherwise reported
stmod_ht.0 <- stan_glmer(plant_ht_cm ~ 1 + (1 | site_id), data = willow.ht.wc,
                      family=Gamma(link="log"),
                      iter = 8000,
                      seed = 1234
                      )

prior_summary(stmod_ht.0)
pp_check(stmod_ht.0)

#### main factors only
## Run, but not used for reporting after model comparisons
stmod_ht1 <- stan_glmer(plant_ht_cm ~ time_class + fenced + (1 | site_id), data = willow.ht.wc,
                      family=Gamma(link="log"),
                      iter = 8000,
                      seed = 1234
                      )

prior_summary(stmod_ht1)
pp_check(stmod_ht1)

```

```{r, cache=TRUE}
#### main factors + interactions
# After model comparisons, this is the selected model reported
stmod_ht2 <- stan_glmer(plant_ht_cm ~ time_class*fenced + (1 | site_id), data = willow.ht.wc,
                      family=Gamma(link="log"),
                      iter = 8000,
                      seed = 1234
                      )

## extract posterior
posteriors.wc.ht <- insight::get_parameters(stmod_ht2)
```

**Prior summary**: plant_ht_cm ~ time_class*fenced + (1 | site_id)  
```{r}
# summary(stmod_ht2)
prior_summary(stmod_ht2)
```

```{r, echo=TRUE, eval=FALSE}
## model comparison.
## Run originally and code retained. stmod_ht2 selected
loo1 <- loo(stmod_ht1,
            k_threshold = 0.7) 
loo2 <- loo(stmod_ht2,
            k_threshold = 0.7) 
comp <- loo_compare(loo2, loo1)
print(comp, simplify = FALSE, digits = 2)
```

```{r}
### model checks
ppc.plot(stmod_ht2) +
  labs(x = "Value", y = "Density") +
  theme_minimal()
# ggsave("./output/figures_exported/willow_height/ppc_willowht_wc.png", width = 4.75, height = 3.75, dpi = 300)
# ggsave("./output/figures_exported/willow_height/ppc_willowht_wc.pdf", width = 4.75, height = 3.75)

```

```{r}
# full posterior description
fun.desc.post1.gt.rope(stmod_ht2) %>% 
  tab_header(title = "Core Winter Range", subtitle = "Willow Height Posterior Description")

# Posterior: point estimates
# point_estimate(stmod_ht2) %>%
#   gt() %>% 
#   tab_header(title = "Indices of centrality",subtitle = "plant_ht_cm ~ time_class*fenced + (1 | site_id)")# Get indices of centrality
```

```{r}
#### PP check plot
ppc.plot(stmod_ht2) +
  labs(caption = "WC: willow_ht~TCxF")
# ggsave("./output/figures_exported/willow_height/ppc_willowht_wc.png", dpi = 300, width = 4.75, height = 3.75)
# ggsave("./output/figures_exported/willow_height/ppc_willowht_wc.pdf", width = 4.75, height = 3.75)
```

### Effect Existence and Significance Testing
```{r}
fun.sexit.gt(stmod_ht2)

## save as RTF
fun.sexit.gt(stmod_ht2) %>% 
  tab_header(title = "Core Winter Range", subtitle = "Willow Height Posterior Description") %>% 
  gt::gtsave(filename = "./output/tables/willow_height/xpostdesc_willowht_TCxF_wc.rtf")
  
```

```{r, echo=FALSE, eval=FALSE}
### Model posterior parameters to table
fun.desc.post1.gt.rope(stmod_ht2) %>% 
  tab_header(title = "Core Winter Range", subtitle = "Willow Height Posterior Description") %>% 
  gt::gtsave(filename = "./output/tables/willow_height/postdesc_willowht_TCxF_wc.rtf")

```

#### Probability of Direction
```{r}
pd.wc <- p_direction(stmod_ht2)

# Visualize the pd
plot(pd.wc) +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Willow height, core winter range")

# ggsave("./output/figures_exported/willow_height/pd_willowht_wc.png", width = 6.75, height =3.75, dpi = 300)
# ggsave("./output/figures_exported/willow_height/pd_willowht_wc.pdf", width = 6.75, height =3.75)

```

#### Region of Practical Equivalence (ROPE)
```{r}
# Region Of Practical Equivalence (ROPE)
rope.wc <- rope(stmod_ht2, ci=0.9)

rope.wc.gt <- rope.wc %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 1,
    suffixing = TRUE
  ) 
rope.wc.gt
# rope.wc.gt %>%
#   gt::gtsave(filename = "./output/tables/willow_height/rope_willowht_wc.rtf")
```

### Contrasts

Results are given on the response scale.

```{r, echo=FALSE, eval=FALSE}
# main effects. Run, but interactions used below
### Contrasts - WC time class
wc.emm.tc <- emmeans(stmod_ht2, ~ time_class)
pairs.wc.tc <- pairs(emmeans(stmod_ht2, ~ time_class))

plot(pairs.wc.tc) +
  theme_minimal() +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Core Winter Range, Willow Height")

## main effect: fencing
wc.emm.fenced <- emmeans(stmod_ht2, ~ fenced)
wc.pairs.fenced <- pairs(wc.emm.fenced)

plot(wc.pairs.fenced) +
  theme_minimal() +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Core Winter Range, Willow Height")

```

```{r}
## Effect: TC x fencing
wc.emm.TCxF <- emmeans(stmod_ht2, ~ time_class | fenced)
wc.pairs.TCxF <- pairs(wc.emm.TCxF, reverse = TRUE)

emmeans(stmod_ht2, ~ time_class | fenced) %>% 
  pairs(reverse = TRUE) %>% 
  plot(type="response") + 
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Core Winter Range, Willow Height")

ggsave("./output/figures_exported/willow_height/emm_willowht_TCxF_wc.pdf", width = 4.5, height = 3.55) # save plot
ggsave("./output/figures_exported/willow_height/emm_willowht_TCxF_wc.png", width = 4.5, height = 3.55) # save plot

# table of contrasts
emmeans(stmod_ht2, ~ time_class | fenced) %>% 
  pairs(reverse = TRUE) %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>%
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log scale")

# table of contrasts. Results are given on the response scale.
emmeans(stmod_ht2, ~ time_class | fenced) %>% 
  pairs(reverse = TRUE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale.")

```

### Central Posterior Interval Estimates

```{r}

bayesplot::mcmc_areas(posteriors.wc.ht,
           pars = c("time_class2013", "time_class2018", "time_class2013:fencedFenced","time_class2018:fencedFenced"),
           prob = 0.9) +
  theme_minimal() +
  labs(title = "Posterior distributions", subtitle = "with medians and 90% intervals", caption = "Willow height, WC")

ggsave("./output/figures_exported/willow_height/mcmcarea_willowht_TCxF_wc.png", dpi = 300, width = 6.75, height = 3.75)
ggsave("./output/figures_exported/willow_height/mcmcarea_willowht_TCxF_wc.pdf", width = 6.75, height = 3.75)

```

## Willow Height - Noncore Winter Range Macroplot Data

### Histograms of Willow Height

```{r}
## histogram
willow.ht.nc %>% 
  ggplot() +
  geom_histogram(aes(plant_ht_cm, fill = fenced), color = "black") +
  viridis::scale_fill_viridis(discrete=TRUE) +
  labs(x="Plant height (cm)", y = "Count", fill = "", caption = "Core winter range plots. Macroplot heights") +
  theme_minimal() +
  facet_wrap(~time_class) +
  coord_flip()

# ggsave("./output/figures_exported/willow_height/hist_willowht_TC_wnc.png", dpi = 300, width = 6.75, height = 3.75)
# ggsave("./output/figures_exported/willow_height/hist_willowht_TC_wnc.pdf", width = 6.75, height = 3.75)

```

### Modeling and Posterior Description  
```{r, cache=TRUE}
## model
#### main factors only
## run with gamma
nc.stmod_ht1 <- stan_glmer(plant_ht_cm ~ time_class +  (1 | site_id), data = willow.ht.nc,
                       family=Gamma(link="log"),
                      iter = 8000,
                      )
posteriors.nc.ht <- insight::get_parameters(nc.stmod_ht1)
```



**Prior summary**: plant_ht_cm ~ time_class +  (1 | site_id)    
```{r}
prior_summary(nc.stmod_ht1)
```

```{r}
#### PP check plot
ppc.plot(nc.stmod_ht1) +
  labs(caption = "WNC: willow_ht~TC")
# ggsave("./output/figures_exported/willow_height/ppc_willowht_wnc.png", dpi = 300, width = 4.75, height = 3.75)
# ggsave("./output/figures_exported/willow_height/ppc_willowht_wnc.pdf", width = 4.75, height = 3.75)

```

### Effect Existence and Significance Testing  
```{r}
# use function defined above
fun.sexit.gt(nc.stmod_ht1)

## save as RTF
fun.sexit.gt(nc.stmod_ht1) %>% 
  tab_header(title = "Noncore Winter Range", subtitle = "Willow Height Posterior Description") %>% 
  gt::gtsave(filename = "./output/tables/willow_height/xposterior_willowht_wnc.rtf")
```

```{r, echo=FALSE, eval = FALSE}
# describe posterior
fun.desc.post1.gt.rope(nc.stmod_ht1)%>% 
  tab_header(title = "Noncore Winter Range", subtitle = "Willow Height Posterior Description")

```

```{r, echo=FALSE, eval=FALSE}
# write to rtf
fun.desc.post1.gt.rope(nc.stmod_ht1) %>% 
  tab_header(title = "Noncore Winter Range", subtitle = "Willow Height Posterior Description") %>% 
  gt::gtsave(filename = "./output/tables/willow_height/posterior_willowht_wnc.rtf")
```

```{r}
# tidy the posteriors
posteriors.nc.ht.tidy <- posteriors.nc.ht %>%
  pivot_longer(cols = starts_with("time"), names_to = "parameter", values_to = "values")

```

#### Probability of Direction 
```{r}
# Visualize the pd
p_direction(nc.stmod_ht1) %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Willow height, Noncore winter range")

ggsave("./output/figures_exported/willow_height/pd_willowht_wnc.png", width = 6.75, height =3.75, dpi = 300)
ggsave("./output/figures_exported/willow_height/pd_willowht_wnc.pdf", width = 6.75, height =3.75)

```

#### Region of Practical Equivalence (ROPE)

```{r}
rope.wnc <- rope(nc.stmod_ht1, ci=0.9)

rope.wnc.gt <- rope.wnc %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 1,
    suffixing = TRUE
  ) 

rope.wnc.gt

rope.wnc.gt %>%
  gt::gtsave(filename = "./output/tables/willow_height/rope_willowht_wnc.rtf")

```

### Contrasts

Results are given on the the response scale.

```{r}
### Contrasts - WC time class
nc.emm <- emmeans(nc.stmod_ht1, ~ time_class)

nc.emm %>% 
  pairs(reverse = TRUE) %>% 
  plot(type="response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Non-core Winter Range, Willow Height")

ggsave("./output/figures_exported/willow_height/emm_willowht_TC_wnc.pdf", width = 4.5, height = 3.55) # save plot
ggsave("./output/figures_exported/willow_height/emm_willowht_TC_wnc.png", width = 4.5, height = 3.55) # save plot

# table of contrasts
emmeans(nc.stmod_ht1, ~ time_class) %>% 
  pairs(reverse = TRUE) %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>%
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log scale")

# table of contrasts. Results are given on the response scale.
emmeans(nc.stmod_ht1, ~ time_class) %>% 
  pairs(reverse = TRUE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale.")

```

### Central Posterior Interval Estimates

```{r}
bayesplot::mcmc_areas(posteriors.nc.ht,
           pars = c("time_class2013","time_class2018"),
           prob = 0.9) +
  theme_minimal() +
  labs(title = "Posterior distributions", subtitle = "with medians and 90% intervals", caption = "Willow height, WNC")

ggsave("./output/figures_exported/willow_height/mcmcarea_willowht_TC_wnc.png", dpi = 300, width = 6.75, height = 3.75)
ggsave("./output/figures_exported/willow_height/mcmcarea_willowht_TC_wnc.pdf", width = 6.75, height = 3.75)

```


```{r}
# Revised plot 2/15/21
# willow height combined plot per revisions
## main effect: fencing
wc.emm.fenced <- emmeans(stmod_ht2, ~ fenced)
wc.pairs.fenced <- pairs(wc.emm.fenced)

plot(wc.pairs.fenced) +
  theme_minimal() +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Core Winter Range, Willow Height")


nc.emm %>% 
  pairs(reverse = TRUE) %>% 
  plot(type="response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Non-core Winter Range, Willow Height")

ggsave("./output/figures_exported/willow_height/emm_willowht_TC_wnc.pdf", width = 4.5, height = 3.55) # save plot
ggsave("./output/figures_exported/willow_height/emm_willowht_TC_wnc.png", width = 4.5, height = 3.55) # save plot


```


## Willow Height - Kawuneeche Valley Macroplot Data

### Histograms of Willow Height

```{r}
willow.ht.wk %>% 
  ggplot() +
  geom_histogram(aes(plant_ht_cm, fill = fenced), color = "black") +
  viridis::scale_fill_viridis(discrete=TRUE) +
  labs(x="Plant height (cm)", y = "Count", fill = "", caption = "KV. Macroplot heights") +
  theme_minimal() +
  facet_wrap(~time_class) +
  coord_flip() +
  theme(legend.position = "none")

ggsave("./output/figures_exported/willow_height/hist_willowht_TC_WK.png", dpi = 300, width = 6.75, height = 3.75)
```

### Modeling and Posterior Description

```{r, cache=TRUE}
#### TC only (no fencing or burned)
stmod_ht_wk_TC <- stan_glmer(plant_ht_cm ~ time_class + (1 | site_id), data = willow.ht.wk,
                      family=Gamma(link="log"),
                      iter = 8000,
                      seed = 1234
                      )

## extract posterior
posteriors.wk.ht <- insight::get_parameters(stmod_ht_wk_TC)
```

**Prior summary**: plant_ht_cm ~ time_class + (1 | site_id)  
```{r}
## prior summary
prior_summary(stmod_ht_wk_TC)
```

```{r, echo = TRUE, eval=TRUE}
#### PP check plot
ppc.plot(stmod_ht_wk_TC) +
  labs(caption = "WK: willow_ht~TC")

ggsave("./output/figures_exported/willow_height/ppc_willowht_wk.png", dpi = 300, width = 4.75, height = 3.75)
ggsave("./output/figures_exported/willow_height/ppc_willowht_wk.pdf", width = 4.75, height = 3.75)
```

### Effect Existence and Significance Testing
```{r}
# use function defined above
fun.sexit.gt(stmod_ht_wk_TC) 

fun.sexit.gt(stmod_ht_wk_TC) %>% 
  tab_header(title = "Kawuneeche Valley", subtitle = "Willow Height Posterior Description") %>% 
  gt::gtsave(filename = "./output/tables/willow_height/xpostdesc_willowht_TC_wk.rtf") 

```

```{r, echo=FALSE, eval=FALSE}
# full posterior description
fun.desc.post1.gt.rope(stmod_ht_wk_TC) %>% 
  tab_header(title = "Kawuneeche Valley", subtitle = "Willow Height Posterior Description")

```

```{r, echo=FALSE, eval=FALSE}
### Model posterior parameters to table
fun.desc.post1.gt.rope(stmod_ht_wk_TC) %>% 
  tab_header(title = "Kawuneeche Valley", subtitle = "Willow Height Posterior Description") %>% 
  gt::gtsave(filename = "./output/tables/willow_height/postdesc_willowht_TC_wk.rtf")
```

#### Probability of Direction
```{r}
pdkv <- p_direction(stmod_ht_wk_TC)
# Visualise the pd
plot(pdkv) +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Willow height, KV")

ggsave("./output/figures_exported/willow_height/pd_willowht_wk.png", width = 6.75, height =3.75, dpi = 300)
ggsave("./output/figures_exported/willow_height/pd_willowht_wk.pdf", width = 6.75, height =3.75)

```

#### Region of Practical Equivalence (ROPE)

```{r}
p.in.rope.kv <- rope(stmod_ht_wk_TC, ci=0.9)

pd.tab.kv <- p.in.rope.kv %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 1,
    suffixing = TRUE
  ) 
pd.tab.kv

pd.tab.kv %>%
  gt::gtsave(filename = "./output/tables/willow_height/rope_willowht_wk.rtf")

```

### Contrasts

```{r}
## contrasts
emmeans(stmod_ht_wk_TC, ~ time_class) %>% 
  pairs() %>% 
  # plot() + # on the lm predict scale
  plot(type = "response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "KV, Willow Height")

ggsave("./output/figures_exported/willow_height/emm_willowht_TC_wk.pdf", width = 4.5, height = 3.55) # save plot
ggsave("./output/figures_exported/willow_height/emm_willowht_TC_wk.png", width = 4.5, height = 3.55) # save plot

# table of contrasts. Results are given on the log (not the response) scale.
emmeans(stmod_ht_wk_TC, ~ time_class) %>% 
  pairs(reverse = TRUE) %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log (not the response) scale.")

# table of contrasts. Results are given on the response scale.
emmeans(stmod_ht_wk_TC, ~ time_class) %>% 
  pairs(reverse = TRUE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale.")

# EMM summaries with type = "response", the tests and confidence intervals are done before back-transforming. The ratios estimated here are actually ratios of geometric means. In general, a model with a log response is in fact a model for relative effects of any of its linear predictors, and this back-transformation to ratios goes hand-in-hand with that.
```

## Willow Cover - Combined Core and Noncore Winter Range Macroplot Data
```{r}
# read in data
# canopy.all.willow.sum <- read_csv("./output/exported_data/willow_cover_20200725.csv") %>% 
#   janitor::clean_names()

canopy.all.willow.sum <- read_csv("./output/exported_data/willow_macroplot_cover_BLto2018.csv") %>% 
  janitor::clean_names()

```

```{r, echo=FALSE}
# munge
canopy.all.willow.sum <- canopy.all.willow.sum %>% 
  mutate(burned = case_when(burned == "Y" ~ "Burned",
                            burned == "N" ~ "Unburned",
                            is.na(burned) ~ "Unburned",
                            TRUE ~ burned)) %>% 
  mutate(time_class = as_factor(time_class)) %>%
  mutate(fenced = as_factor(fenced)) %>%
  mutate(burned = as_factor(burned)) %>% 
  mutate(time_class = fct_rev(time_class)) %>% 
  mutate(fenced = fct_rev(fenced))

```

```{r}
### create subsets of data for modeling
# pull out WK, note different years of data
cov.wk <- canopy.all.willow.sum %>% 
  filter(site_type == "WK") %>%
  mutate(time_class = as.character(time_class)) %>% 
  filter(time_class == "2016" | time_class == "2017" | time_class == "2018") %>% 
  mutate(time_class = as.factor(time_class))

# pull out the core plus noncore winter range plots
cov.wc.wnc <- canopy.all.willow.sum %>% 
  filter(time_class == "BL" | time_class == "2013" | time_class == "2018") %>%
  filter(site_type == "WNC" | site_type == "WC") %>% 
  mutate(cover_allwillow = cover_allwillow/100) %>% 
  filter(!is.na(cover_allwillow))

# pull out the core winter range plots
cov.wc <- canopy.all.willow.sum %>% 
  filter(site_type == "WC") %>%
  filter(time_class == "BL" | time_class == "2013" | time_class == "2018") %>%
  mutate(cover_allwillow = cover_allwillow/100) %>%
  filter(!is.na(cover_allwillow))

# pull out the noncore winter range plots
cov.wnc <- canopy.all.willow.sum %>% 
  filter(site_type == "WNC") %>% 
  filter(time_class == "BL" | time_class == "2013" | time_class == "2018") %>%
  mutate(cover_allwillow = cover_allwillow/100) %>% 
  filter(!is.na(cover_allwillow))

```

```{r, echo=FALSE, eval=FALSE, fig.width=7, fig.height=7}
## histograms
canopy.all.willow.sum %>% 
  ggplot(aes(cover_allwillow)) +
  geom_histogram() +
  facet_grid(time_class~site_type) +
  theme_minimal()

```

### Summary Tables

```{r}
#### summary tables: time class and burned
cov.wc.wnc %>% 
  group_by(time_class, fenced) %>% 
  summarytools::descr(cover_allwillow,stats = "common") %>% 
  summarytools::tb() %>%
  mutate_if(.predicate = is.numeric,.funs = round, digits = 2) %>% 
  gt() %>% 
  tab_header(title = "Combined WC and WNC macroplot data: willow cover - fenced + time_class")

## time class, fenced and burned
cov.wc.wnc %>% 
  group_by(time_class, fenced, burned) %>% 
  summarytools::descr(cover_allwillow, stats = "common") %>% 
  summarytools::tb() %>%
  mutate_if(.predicate = is.numeric,.funs = round, digits = 2) %>% 
  gt() %>% 
  tab_header(title = "Combined WC and WNC macroplot data: willow cover ~ burned + fenced + time_class")

```

### Histograms of Willow Cover

```{r, fig.width=7, fig.height=5}
cov.wc.wnc %>% 
  ggplot(aes(cover_allwillow)) +
  geom_histogram() +
  facet_grid(time_class~site_type) +
  theme_minimal()

ggsave("./output/figures_exported/willow_cover/hist_willowht_TC_wcwnc.png", dpi = 300, width = 6.75, height = 3.75)
```

### Modeling and Posterior Description


```{r, cache=TRUE}
# cover_allwillow \~ time_class \* fenced + (1 \| site_id)
## prep for modeling
## subtract from max to conform
#### Model
cov.wc.wnc <- cov.wc.wnc %>% 
  mutate(cover_allwillow = case_when(cover_allwillow == 1 ~ 0.999999,
                                     TRUE ~ cover_allwillow))

## model
stmod_cov.wc.wnc <- stan_glmer(cover_allwillow ~ time_class * fenced + (1 | site_id), 
                      data = cov.wc.wnc,
                      family = mgcv::betar,
                      prior_aux = exponential(1),
                      iter = 8000,
                      seed = 1234)

```

**Prior summary**: cover_allwillow ~ time_class * fenced + (1 | site_id)  
```{r}
prior_summary(stmod_cov.wc.wnc)
```


```{r, echo=FALSE, eval=FALSE}
# alt model with burned as predictor
# stmod_cov.wc.wnc.b <- stan_glmer(cover_allwillow ~ time_class * fenced * burned + (1 | site_id), 
#                       data = cov.wc.wnc,
#                       family = mgcv::betar,
#                       #prior_aux = exponential(2),
#                       iter = 10000,
#                       seed = 12345)

# there are limits to the beta, but it still seems like the most sensible distribution. Unless want to go with NLR

# pp_check(stmod_cov.wc.wnc.b)

## loo - model comparison  
# cov.wc.wnc.loo1 <- loo(stmod_cov.wc.wnc)
# cov.wc.wnc.loo2 <- loo(stmod_cov.wc.wnc.b)
# loo_compare(cov.wc.wnc.loo1, cov.wc.wnc.loo2)
#                    elpd_diff se_diff
# stmod_cov.wc.wnc    0.0       0.0   
# stmod_cov.wc.wnc.b -2.2       2.5 
# original model better.
```

```{r}
#### PP check plot
# prior_summary(stmod_cov.wc.wnc)

pp_check(stmod_cov.wc.wnc) +
  labs(x = "Value", y = "Density") +
  theme_minimal() +
  labs(caption = "WC & WNC: cover_allwillow ~ time_class * fenced + (1 | site_id)")

ggsave("./output/figures_exported/willow_cover/ppc_willowcov_wcwnc.png", dpi = 300, width = 4.75, height = 3.75)
ggsave("./output/figures_exported/willow_cover/ppc_willowcov_wcwnc.pdf", width = 4.75, height = 3.75)

```

### Effect Existence and Significance Testing
```{r}
fun.sexit.gt(stmod_cov.wc.wnc$stanfit)

fun.sexit.gt(stmod_cov.wc.wnc$stanfit) %>%
  tab_header(title = "Combined Core and Noncore Winter Range", subtitle = "Willow Cover Posterior Description") %>% 
  gt::gtsave(filename = "./output/tables/willow_cover/xposterior_willowht_wcwnc.rtf")
```

```{r, echo=FALSE, eval=FALSE}
#### Model posterior parameters to table
fun.desc.post1.gt.rope(stmod_cov.wc.wnc$stanfit)
```

```{r, echo=FALSE, eval=FALSE}
## save rtf
fun.desc.post1.gt.rope(stmod_cov.wc.wnc$stanfit) %>%
  tab_header(title = "Combined Core and Noncore Winter Range", subtitle = "Willow Cover Posterior Description") %>% 
  gt::gtsave(filename = "./output/tables/willow_cover/posterior_willowht_wcwnc.rtf")
```


#### Probability of Direction  
```{r}
# Visualize the pd
p_direction(stmod_cov.wc.wnc) %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Willow cover, Core and noncore winter range")

ggsave("./output/figures_exported/willow_cover/pd_willowht_wcwnc.png", width = 6.75, height =3.75, dpi = 300)
ggsave("./output/figures_exported/willow_cover/pd_willowht_wcwnc.pdf", width = 6.75, height =3.75)

```

#### Region of Practical Equivalence (ROPE)  
```{r}
rope.cov.wcwnc.gt <- rope(stmod_cov.wc.wnc$stanfit, ci=0.9) %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 2,
    suffixing = TRUE
  ) 

rope.cov.wcwnc.gt %>%
  gt::gtsave(filename = "./output/tables/willow_cover/rope_willowcov_wcwnc.rtf")
```

### Contrasts  
```{r}
# contrast - time class and fencing
emmeans(stmod_cov.wc.wnc, ~ time_class | fenced) %>% 
  pairs(reverse=TRUE) %>% 
  # plot() + # Results are given on the logit (not the response) scale.
  plot(type = "response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Combined Core and Noncore Winter Range, Willow Cover")

ggsave("./output/figures_exported/willow_cover/emm_willowht_TCxF_wcwnc.png", width = 6.75, height =3.75, dpi = 300)
ggsave("./output/figures_exported/willow_cover/emm_willowht_TCxF_wcwnc.pdf", width = 6.75, height =3.75)


# table of contrasts
emmeans(stmod_cov.wc.wnc, ~ time_class | fenced) %>% 
  pairs(reverse = TRUE) %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log odds ratio, not the response scale")

emmeans(stmod_cov.wc.wnc, ~ time_class | fenced) %>% 
  pairs(reverse = TRUE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale")

```

### Central Posterior Interval Estimates  
```{r}
## extract posterior
posteriors.cov.wcwnc <- insight::get_parameters(stmod_cov.wc.wnc)
# posteriors.cov.wcwnc %>% names()

bayesplot::mcmc_areas(posteriors.cov.wcwnc,
           pars = c("time_class2013", "time_class2018", "time_class2013:fencedFenced","time_class2018:fencedFenced"),
           prob = 0.9) +
  theme_minimal() +
  labs(title = "Posterior distributions", subtitle = "with medians and 90% intervals", caption = "Willow height, WC & WNC")

ggsave("./output/figures_exported/willow_cover/mcmcarea_willowcov_TCxF_wcwnc.png", dpi = 300, width = 6.75, height = 3.75)
ggsave("./output/figures_exported/willow_cover/mcmcarea_willowcov_TCxF_wcwnc.pdf", width = 6.75, height = 3.75)

```

### Burning

```{r, cache=TRUE, eval=TRUE}
#### Fenced and burned

stmod_cov.wc.wnc.TCxFxB <- stan_glmer(cover_allwillow ~ time_class * burned * fenced + (1 | site_id), 
                      data = cov.wc.wnc,
                      family = mgcv::betar,
                      prior_aux = exponential(1),
                      iter = 10000,
                      seed = 1234)

```

**Prior summary**: plant_ht_cm ~ time_class + (1 | site_id)  
```{r}
prior_summary(stmod_cov.wc.wnc.TCxFxB)
```

```{r}
pp_check(stmod_cov.wc.wnc.TCxFxB) +
  labs(x = "Value", y = "Density") +
  theme_minimal()
# 
ggsave("./output/figures_exported/wc_wnc_willowcover_ppc_TCxFxB.png", dpi = 300, width = 4.75, height = 3.75)
ggsave("./output/figures_exported/wc_wnc_willowcover_ppc_TCxFxB.pdf", width = 4.75, height = 3.75)

```

### Effect Existence and Significance Testing
```{r}

fun.sexit.gt(stmod_cov.wc.wnc.TCxFxB$stanfit)
```


## Willow Cover - Core Winter Range Macroplot Data

### Summary Tables

```{r}
#### summary tables: time class and burned
cov.wc %>% 
  group_by(time_class) %>% 
  summarytools::descr(cover_allwillow,stats = "common") %>% 
  summarytools::tb() %>%
  mutate_if(.predicate = is.numeric,.funs = round, digits = 2) %>% 
  gt() %>% 
  tab_header(title = "WC macroplot data: willow cover ~ WC: all willow cover - time_class")

## time class, fenced and burned
cov.wc %>% 
  group_by(time_class, burned) %>% 
  summarytools::descr(cover_allwillow, stats = "common") %>% 
  summarytools::tb() %>%
  mutate_if(.predicate = is.numeric,.funs = round, digits = 2) %>% 
  gt() %>% 
  tab_header(title = "WC macroplot data: All willow cover - burned + time_class")

```

### Modeling and Posterior Description

```{r, cache=TRUE}
#### Model
## add to min and subtract from max to conform
cov.wc <- cov.wc %>% 
  mutate(cover_allwillow = case_when(cover_allwillow == 1 ~ 0.999999,
                                     TRUE ~ cover_allwillow))


stmod_cov.wc <- stan_glmer(cover_allwillow ~ time_class * fenced + (1 | site_id), 
                      data = cov.wc,
                      family = mgcv::betar,
                      prior_aux = exponential(2),
                      iter = 10000,
                      seed = 1234)

```

**Prior summary**: cover_allwillow ~ time_class * fenced + (1 | site_id)    
```{r}
prior_summary(stmod_cov.wc)
```

```{r}
#### PP check plot
pp_check(stmod_cov.wc) +
  labs(x = "Value", y = "Density") +
  theme_minimal() +
  labs(caption = "WC: cover_allwillow ~ time_class * fenced + (1 | site_id)")

ggsave("./output/figures_exported/willow_cover/ppc_willowcov_wc.png", dpi = 300, width = 4.75, height = 3.75)
ggsave("./output/figures_exported/willow_cover/ppc_willowcov_wc.pdf", width = 4.75, height = 3.75)

```
### Effect Existence and Significance Testing
```{r}

fun.sexit.gt(stmod_cov.wc$stanfit)

fun.sexit.gt(stmod_cov.wc$stanfit) %>%
  tab_header(title = "Core Winter Range", subtitle = "Willow Cover Posterior Description") %>% 
  gt::gtsave(filename = "./output/tables/willow_cover/posterior_willowht_wc.rtf")
```

```{r, echo=FALSE, eval=FALSE}
#### Model posterior parameters to table
fun.desc.post1.gt.rope(stmod_cov.wc$stanfit)
```

```{r, echo=FALSE, eval=FALSE}
## save rtf
fun.desc.post1.gt.rope(stmod_cov.wc$stanfit) %>%
  tab_header(title = "Core Winter Range", subtitle = "Willow Cover Posterior Description") %>% 
  gt::gtsave(filename = "./output/tables/willow_cover/posterior_willowht_wc.rtf")

```

#### Probability of Direction
```{r}
# Visualize the pd
p_direction(stmod_cov.wc) %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Willow cover, Core and noncore winter range")

ggsave("./output/figures_exported/willow_cover/pd_willowht_wc.png", width = 6.75, height =3.75, dpi = 300)
ggsave("./output/figures_exported/willow_cover/pd_willowht_wc.pdf", width = 6.75, height =3.75)

```

#### Region of Practical Equivalence (ROPE)
```{r}
rope.cov.wc.gt <- rope(stmod_cov.wc$stanfit, ci=0.9) %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 2,
    suffixing = TRUE
  ) 

rope.cov.wcwnc.gt %>%
  gt::gtsave(filename = "./output/tables/willow_cover/rope_willowcov_wc.rtf")
```

### Contrasts

Results are given on the the response scale.

```{r}
# library(emmeans)
wc.cov.emm.TCxF <- emmeans(stmod_cov.wc, ~ time_class | fenced)

pairs(wc.cov.emm.TCxF, reverse = TRUE, type="response") %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Core Winter Range, Willow Cover")

ggsave("./output/figures_exported/willow_cover/emm_willowht_TCxF_wc.png", width = 6.75, height =3.75, dpi = 300)
ggsave("./output/figures_exported/willow_cover/emm_willowht_TCxF_wc.pdf", width = 6.75, height =3.75)

# table of contrasts
emmeans(stmod_cov.wc, ~ time_class | fenced) %>% 
  pairs(reverse = TRUE) %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log odds ratio, not the response scale")

emmeans(stmod_cov.wc, ~ time_class | fenced) %>% 
  pairs(reverse = TRUE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale")

```

## Willow Cover - Noncore Winter Range Macroplot Data

### Summary Tables

```{r}
#### summary tables: time class and burned
cov.wnc %>% 
  group_by(time_class, fenced) %>% 
  summarytools::descr(cover_allwillow,stats = "common") %>% 
  summarytools::tb() %>%
  mutate_if(.predicate = is.numeric,.funs = round, digits = 2) %>% 
  gt() %>% 
  tab_header(title = "WNC macroplot data: willow cover")

```

### Modeling and Posterior Description

```{r, cache=TRUE}
#### Model
## add to min and subtract from max to conform
cov.wnc <- cov.wnc %>% 
  mutate(cover_allwillow = case_when(cover_allwillow == 1 ~ 0.999999,
                                     TRUE ~ cover_allwillow))

stmod_cov.wnc <- stan_glmer(cover_allwillow ~ time_class + (1 | site_id), 
                      data = cov.wnc,
                      family = mgcv::betar,
                      # prior_aux = exponential(2),
                      iter = 15000,
                      seed = 1234)



```

```{r, echo=FALSE, eval=FALSE}
shinystan::launch_shinystan(stmod_cov.wnc)
```

**Prior summary**: cover_allwillow ~ time_class + (1 | site_id)      
```{r}
prior_summary(stmod_cov.wnc)
```

```{r, echo=FALSE, eval=FALSE}
#### PP check plot
# summary(stmod_cov.wc)
# prior_summary(stmod_cov.wnc)

pp_check(stmod_cov.wnc) +
  labs(x = "Value", y = "Density") +
  theme_minimal() +
  labs(caption = "WNC: cover_allwillow ~ time_class * fenced + (1 | site_id)")

ggsave("./output/figures_exported/willow_cover/ppc_willowcov_wc.png", dpi = 300, width = 4.75, height = 3.75)
ggsave("./output/figures_exported/willow_cover/ppc_willowcov_wc.pdf", width = 4.75, height = 3.75)

# cov.wnc %>% 
#   ggplot(aes(cover_allwillow)) +
#   geom_density()
```

### Effect Existence and Significance Testing
```{r}
fun.sexit.gt(stmod_cov.wnc$stanfit)

fun.sexit.gt(stmod_cov.wnc$stanfit) %>% 
  tab_header(title = "Noncore Winter Range", subtitle = "Willow Cover Posterior Description") %>% 
  gt::gtsave(filename = "./output/tables/willow_cover/xposterior_willowht_wnc.rtf")
```

```{r, echo=FALSE, eval=FALSE}
#### Model posterior parameters to table
fun.desc.post1.gt.rope(stmod_cov.wnc$stanfit)
```

```{r, echo=FALSE, eval=FALSE}
## save rtf
fun.desc.post1.gt.rope(stmod_cov.wnc$stanfit) %>%
  tab_header(title = "Noncore Winter Range", subtitle = "Willow Cover Posterior Description") %>% 
  gt::gtsave(filename = "./output/tables/willow_cover/posterior_willowht_wnc.rtf")

```

#### Probability of Direction  
```{r}
# Visualize the pd
p_direction(stmod_cov.wnc) %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Willow cover, Nnoncore winter range")

ggsave("./output/figures_exported/willow_cover/pd_willowht_wnc.png", width = 6.75, height =3.75, dpi = 300)
ggsave("./output/figures_exported/willow_cover/pd_willowht_wnc.pdf", width = 6.75, height =3.75)

```

#### Region of Practical Equivalence (ROPE) 
```{r}
rope.cov.wnc.gt <- rope(stmod_cov.wnc$stanfit, ci=0.9) %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 2,
    suffixing = TRUE
  ) 

rope.cov.wnc.gt %>%
  gt::gtsave(filename = "./output/tables/willow_cover/rope_willowcov_wnc.rtf")
```

### Contrasts
```{r}
# library(emmeans)
wnc.cov.emm.TC <- emmeans(stmod_cov.wnc, ~ time_class)

pairs(wnc.cov.emm.TC, reverse = TRUE, type="response") %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Noncore Winter Range, Willow Cover. Results are given on the response scale.")

# ggsave("./output/figures_exported/willow_cover/emm_willowht_TCxF_wnc.png", width = 6.75, height =3.75, dpi = 300)
# ggsave("./output/figures_exported/willow_cover/emm_willowht_TCxF_wnc.pdf", width = 6.75, height =3.75)

# table of contrasts
emmeans(stmod_cov.wnc, ~ time_class) %>% 
  pairs(reverse = TRUE) %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log odds ratio, not the response scale")

emmeans(stmod_cov.wnc, ~ time_class) %>% 
  pairs(reverse = TRUE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale")

```

```{r, wbasin_ht, echo=FALSE, eval=FALSE}
## Wild basin
# Wild Basin willow data survey 2018.csv

wbasin <- read_csv("./data/wild_basin/Wild Basin willow data survey 2018.csv")

wbasin <- wbasin %>%
  janitor::clean_names()
  
wbasin.loc <- wbasin %>% 
  filter(!is.na(utm_e_nad83_end_point))

```

# Session info

R version 4.0.3 (2020-10-10)  
Platform: x86_64-w64-mingw32/x64 (64-bit)  
Running under: Windows 10 x64 (build 19041)  
Run date: `r format(Sys.time(), '%Y %B %d')`  

```{r, echo=FALSE}
subset(data.frame(sessioninfo::package_info()), attached==TRUE, c(package, loadedversion)) %>% 
  gt() %>% 
  tab_header(title = "Attached packages and versions")

```

# References
