---
title: "Appendix C. EVMP Aspen Statistical Analyses"
author: ""
date: ""
output: 
  html_document: 
    theme: journal
    toc: yes
    toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

_Code author:_ 
E. Gage  
https://github.com/egage/EVMP    

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = 'center')
opts_knit$set(root.dir=normalizePath('../')) # this is required if Rmd is nested below the project directory
opts_chunk$set(fig.path = "../output/figures/") # corrected path and added dev. Needed to specify a subdirectory for figs

set.seed(1234)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=FALSE, eval=TRUE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(janitor))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(lubridate))
library(emmeans)
library(rstanarm)
options(mc.cores = parallel::detectCores())
library(bayestestR)
library(gt)
library(ggridges)
library(shinystan)
library(patchwork)
```

```{r, functions, echo=FALSE, warning=FALSE, message=FALSE, comment=FALSE, eval=TRUE}
### Table functions

fun.sexit.gt <- function(model){
  bayestestR::sexit(model, ci=0.9) %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric), round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Effect Existence and Significance Testing") 
  }

## model summary to DT
fun.model.summary <- function(model){
  summary(model,
          probs = c(0.1, 0.5, 0.9),
          digits = 1) %>%
  as_tibble(rownames = "Parameter") %>%
  mutate(across(where(is.numeric),round,2)) %>%
  datatable(caption = "Model summary")
}

## function to describe model output
fun.desc.post1 <- function(model){
  describe_posterior(
  model,
  effects = "fixed",
  component = "all",
  ci_method = "hdi",
  ci=0.9,
  test = c("p_direction", "p_significance"),
  centrality = "all"
) %>% 
    as_tibble() %>% 
    DT::datatable(caption = "Posterior description")
}

## gt
# describe posterior
fun.desc.post1.gt.rope <- function(model){
  describe_posterior(
  model,
  effects = "fixed",
  component = "all",
  ci_method = "hdi",
  ci=0.9,
  test = c("p_direction", "p_significance","rope"),
  #rope_range(c(-0.1,0.1)),
  rope_range("default"),
  rope_ci = 0.95,
  centrality = "median"
) %>% 
    as_tibble() %>% 
    gt() %>% 
    tab_header(title = "Posterior description") %>% 
    fmt_number(
    columns = 2:14,
    decimals = 2,
    suffixing = TRUE
  )
}

#### Plotting functions
fun.contrast.plot <- function(df){
  df %>% 
  ggplot(aes(x = contrast)) +
  geom_linerange(aes(ymin = lower.HPD, ymax = upper.HPD), color = "lightblue", size = 2) +
  geom_point(aes(y = estimate), size = 2.5) +
  theme_minimal() +
  labs(caption = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Estimate") +
  coord_flip()
}

## ppcheck.plot
ppc.plot <- function(model){
  pp_check(model) +
  labs(x = "Value", y = "Density") +
  theme_minimal()
}

## describe posterior to DT
desc.post1 <- function(model){
  describe_posterior(
  model,
  effects = "all",
  component = "all",
  test = c("p_direction", "p_significance"),
  centrality = "all"
) %>% 
    as_tibble() %>% 
    DT::datatable(caption = "Posterior description")
}

```

# Introduction

This document contains code for running statistical analyses of of Elk
Vegetation Management Plan (EVMP) data collected through the 2018
sampling season. Analyses are limited to data from plots established in
aspen communities in areas of the elk winter range (core and noncore
areas) and the Kawuneeche Valley (i.e., "aspen" plots, see Zeigenfuss et
al. 2011) and is provided as a supplement to the NPS NRR Report
"Monitoring of Vegetation Response to Elk Population and Habitat
Management in Rocky Mountain National Park - Analysis of Elk Vegetation
Management Plan monitoring Data: 2008--2018". For information on
sampling and the broader analysis and interpretation of the results,
refer to this report, past analyses [@zeigenfuss2015], and the original
EVMP monitoring plan [@zeigenfuss2011].

# Methods

The code and results presented here start from and build upon derived
data produced in a separate code document focused on ingesting,
compiling, and cleaning raw data provided by RMNP staff. Extensive
"munging" of raw files was required, so the decision was made to
separate code used to pre-process raw data from code used for the
analyses presented here. Likewise, separate code documents were
developed for other distinct elements of the larger EVMP (e.g., willow
and upland plots). In general, packages included in the "Tidyverse"
ecosystem of R packages were used for data transformation and
visualization [@wickham2019], although specialized packages particular
to specific tasks (e.g., the "bayesplot" plot for visualization of
Bayesian posterior distributions) were also used [@gabry2019].

Bayesian repeated measures analyses were fit separately for combined
core and non-core winter range, core winter range (WC) and non-core
winter range plots (WNC) with plots treated as random factors using the
'stan_glmer' function in the rstanrm package
[@goodrich2020][@brilleman2018]. Bayesian estimation was performed via
MCMC adding independent weakly informed priors specific to the data type
being modeled for regression coefficients Lemoine (2019) for difference
response variable types (e.g., aspen stem count). Continuous response
variables (e.g., willow height) were fit with a Gamma distribution,
which restricts possible responses to positive real numbers. Count data
like aspen stem counts were modeled as poisson processes, while
proportion data (e.g., cover) were modeled using a beta distribution
(Ferrari and Cribari-Neto 2004) that constrains values from 0 to 1. For
willow height, models were fit using weekly informative priors and a
gamma distribution with the "stan_lmer" function in the "rstanarm"
package [@goodrich2020][@brilleman2018].

The Probability of Direction (PD) was used to represent certainty
associated with the most probable direction (positive or negative) of
the effect (e.g., time class, fencing). The PD is correlated with the
frequentist p-value, with a two-sided p-value of respectively .1, .05,
and .01 approximated by a PD of 95%, 97.5%, and 99.5% [@makowski2019].
the "region of practical equivalence" (ROPE) to evaluate the probability
of a parameter being outside a range that can be considered as
"practically no effect", i.e., a region enclosing values that are
equivalent to the null value for practical purposes [@kruschke2018],
with the proportion of the 95% HDI credible interval falling within the
ROPE was used as an index for an analog to frequentist "null-hypothesis"
testing.

```{r}
## read in data
# Aspen ht
ht_perc1 <- read_csv("./output/exported_data/asp_ht_perc1_20200309.csv")

# Data munging

## filter out in-between years
ht_perc1 <- ht_perc1 %>%
  filter(timeClass %in% c("BL","2013", "2018")) 

### factor relevel
ht_perc1 <- ht_perc1 %>% 
  mutate(timeClass = as.factor(timeClass)) %>% 
  mutate(timeClass = fct_relevel(timeClass, "BL", "2013", "2018")) %>% 
  mutate(RANGE_TYPE = as.factor(RANGE_TYPE)) %>% 
  mutate(RANGE_TYPE = fct_relevel(RANGE_TYPE, "core winter range","non-core winter range","Kawuneeche Valley"))   

```

```{r}
ht_perc1 <- ht_perc1 %>% 
  clean_names()

## set factor levels for burned and fenced
ht_perc1 <- ht_perc1 %>%
  mutate(fenced.long = case_when(fenced == "N" ~ "Unfenced",
                                 fenced == "Y" ~ "Fenced",
                                 TRUE ~ fenced)) %>% 
  mutate(burned = case_when(burned == "Not burned" ~ "Unburned",
                            TRUE ~ burned)) %>%
  mutate(fenced.long = as_factor(fenced.long)) %>%
  mutate(burned = as_factor(burned))

# ht_perc1 %>% 
#   distinct(site_type)
  
asp.ac <- ht_perc1 %>% 
  filter(site_type == "AC") 

# non core
asp.anc <- ht_perc1 %>% 
  filter(site_type == "ANC")

asp.ak <- ht_perc1 %>% 
  filter(site_type == "AK")

```

# Results

```{r, eval=FALSE}
### Traditional AOV
aov_model_asp_htperc <- aov(stemDen.ha ~ RANGE_TYPE*timeClass + Error(SITE_ID/(RANGE_TYPE*timeClass)), data = ht_perc1)

aov_model_asp_htperc <- aov(stemDen.ha ~ RANGE_TYPE*timeClass + Error(1/SITE_ID), data = ht_perc1)

```

## Aspen Stem Count - Core Winter Range (AC)

### Modeling and Posterior Description

```{r, eval=FALSE}
#### Main Factors Only
## Not run. Used in model comparisons, but model w/ interaction selected
asp.ac <- asp.ac %>%
  mutate(stem_den_ha = stem_den_ha + 0.0000001)

## poisson
stmod_stally1 <- stan_glmer(stem_tally ~ time_class + fenced + (1 | site_id), data = asp.ac,
                      family=poisson,
                      iter = 8000,
                      seed = 1234
                      )
prior_summary(stmod_stally1)
pp_check(stmod_stally1)
```

```{r}
#### main factors + interactions:
stmod_stally2 <- stan_glmer(stem_tally ~ time_class * fenced + (1 | site_id), data = asp.ac,
                      family= poisson,
                      iter = 8000,
                      seed = 1234
                      )

```

**Prior summary**: stem_tally ~ time_class * fenced + (1 | site_id)  
```{r}
prior_summary(stmod_stally2)
```

```{r}
## model summary
fun.model.summary(stmod_stally2)
```

```{r}
# extract posteriors
posteriors.ac.stcnt <- insight::get_parameters(stmod_stally2)

# tidy
posteriors.ac.stcnt.tidy <- posteriors.ac.stcnt %>%
  pivot_longer(cols = starts_with("time"), names_to = "parameter", values_to = "values")

```

```{r}
### ppc plot
ppc.plot(stmod_stally2) +
  labs(caption = "AC: stem_cntt~TCxF")

# save plots
ggsave("./output/figures_exported/aspen_stem_count/ppc_stemcnt_ac.png", dpi = 300, width = 4.75, height = 3.75)
ggsave("./output/figures_exported/aspen_stem_count/ppc_stemcnt_ac.pdf", width = 4.75, height = 3.75)
```

```{r, eval = FALSE}
## model comparison. Run previously for model comparison;
# not currently set to run to minimize computational overhead

loo1 <- loo(stmod_stally1,
            k_threshold = 0.7) 
loo2 <- loo(stmod_stally2,
            k_threshold = 0.7) 
comp <- loo_compare(loo2, loo1)

## create table of comparisons
modcomp_ac <- print(comp, simplify = TRUE, digits = 2)

modcomp_ac %>% 
  gt()
# citation(package = "loo")

# show more details with simplify=FALSE
print(modcomp_ac, simplify = FALSE, digits = 3)

# model with interactions selected
```

### Effect Existence and Significance Testing
```{r}
fun.sexit.gt(stmod_stally2)

fun.sexit.gt(stmod_stally2) %>%
 tab_header(title = "Aspen Count Posterior Description", subtitle = "Core Winter Range") %>%
  gt::gtsave(filename = "./output/tables/aspen_stem_count/xpostdesc_stemcnt_TCxF_ac.rtf")

```


```{r, echo=FALSE, eval=FALSE}
### Model posterior parameters to table
fun.desc.post1.gt.rope(stmod_stally2) %>% 
  tab_header(title = "Aspen Count Posterior Description", subtitle = "Core Winter Range")

## save as RTF
fun.desc.post1.gt.rope(stmod_stally2) %>%
 tab_header(title = "Aspen Count Posterior Description", subtitle = "Core Winter Range") %>%
  gt::gtsave(filename = "./output/tables/aspen_stem_count/postdesc_stemcnt_TCxF_ac.rtf")

```

```{r}
# library(ggridges)
posteriors.ac.stcnt.tidy %>% 
  ggplot() +
  geom_density(aes(fill = parameter, x = values, color = parameter), alpha = .12) +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  theme_minimal()

# save plots
ggsave("./output/figures_exported/aspen_stem_count/postden_stemcnt_ac.png", dpi = 300, width = 4.75, height = 3.75)
ggsave("./output/figures_exported/aspen_stem_count/postden_stemcnt_ac.pdf", width = 4.75, height = 3.75)
```

#### Probability of Direction
```{r}
# Visualize the pd
#### PD - AC
p_direction(stmod_stally2) %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Aspen stem count, core winter range")

# save plots
ggsave("./output/figures_exported/aspen_stem_count/pd_stemcnt_ac.png", dpi = 300, width = 4.75, height = 3.75)
ggsave("./output/figures_exported/aspen_stem_count/pd_stemcnt_ac.pdf", width = 4.75, height = 3.75)

```

#### Region of Practical Equivalence (ROPE)
```{r}
rope.stemcnt.ac <- rope(stmod_stally2, ci=0.9) %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 1,
    suffixing = TRUE
  ) 

rope.stemcnt.ac %>% 
  gt::gtsave(filename = "./output/tables/aspen_stem_count/rope_stemcnt_ac.rtf")

```

### Contrasts
```{r}
#### Contrasts
# EMMs for later factor levels are subtracted from those for earlier levels; if you want the comparisons to go in the other direction, use pairs(pigs.emm.s, reverse = TRUE).

# library(emmeans)
emmeans(stmod_stally2, ~ time_class | fenced) %>% 
  pairs(reverse = TRUE) %>% 
  # plot(type = "response") +
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  # labs(title= "Core winter range", caption = "Point estimate displayed: median 
# HPD interval probability: 0.95, Results on log scale", x = "Estimate", y = "Contrast")
  labs(subtitle = "Point estimate displayed: median HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Core Winter Range, Aspen Stem Count, Results are given on the log scale")


# ggsave("./output/figures_exported/aspen_stem_count/emm_stemcnt_TCxF_ac.pdf", width = 4.5, height = 3.55) # save plot
# ggsave("./output/figures_exported/aspen_stem_count/emm_stemcnt_TCxF_ac.png", width = 4.5, height = 3.55) # save plot

# table of contrasts. Results are given on the log (not the response) scale.
emmeans(stmod_stally2, ~ time_class | fenced) %>% 
  pairs(reverse = TRUE) %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log (not the response) scale.")

# table of contrasts. Results are given on the response scale.
emmeans(stmod_stally2, ~ time_class | fenced) %>% 
  pairs(reverse = TRUE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale.")

# EMM summaries with type = "response", the tests and confidence intervals are done before back-transforming. The ratios estimated here are actually ratios of geometric means. In general, a model with a log response is in fact a model for relative effects of any of its linear predictors, and this back-transformation to ratios goes hand-in-hand with that.

```

### Effects of Burning
```{r}
ht_perc1 %>% 
  filter(range_type == "core winter range") %>% 
  filter(time_class == "BL" | time_class == "2013" | time_class == "2018") %>%
  group_by(time_class, range_type, burned, fenced) %>% 
  summarytools::descr(stats = "fivenum") %>%
  summarytools::tb() %>% 
  gt() %>% 
  tab_header(title="Summary statistics")

```

```{r, cache=TRUE}
### STAN model
stmod_stally_TCxFxB_ac <- stan_glmer(stem_tally ~ fenced * time_class * burned + (1 | site_id), data = asp.ac,
                      family=poisson,
                      iter = 8000,
                      seed = 1234
                      )

```

```{r}
fun.sexit.gt(stmod_stally_TCxFxB_ac)

fun.sexit.gt(stmod_stally_TCxFxB_ac) %>%
 tab_header(title = "Aspen Count Posterior Description", subtitle = "Core Winter Range") %>%
  gt::gtsave(filename = "./output/tables/aspen_stem_count/xpostdesc_stemcnt_TCxFxB_ac.rtf")

```


```{r, eval=FALSE}
prior_summary(stmod_stally_TCxFxB_ac)
pp_check(stmod_stally_TCxFxB_ac)
```

```{r}
# Visualize the pd
p_direction(stmod_stally_TCxFxB_ac) %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Aspen stem count, core winter range, TCxFxB")

# save plots
ggsave("./output/figures_exported/aspen_stem_count/pd_stemcnt_TCxFxB_ac.png", dpi = 300, width = 4.75, height = 3.75)
ggsave("./output/figures_exported/aspen_stem_count/pd_stemcnt_TCxFxB_ac.pdf", width = 4.75, height = 3.75)

```

```{r}
#### Contrasts
# linear predictors
# emmeans(stmod_stally_TCxFxB_ac, ~ time_class | burned | fenced) %>% 
#   pairs(reverse = TRUE) %>% 
#   plot() +
#   theme_minimal() +
#   geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
#   labs(subtitle = "Point estimate displayed: median 
# HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Core Winter Range, Aspen Stem Count, TCxFxB")

## response scale
emmeans(stmod_stally_TCxFxB_ac, ~ time_class | burned | fenced) %>% 
  pairs(reverse = TRUE) %>% 
  plot(type = "response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Core Winter Range, Aspen Stem Count, TCxFxB")


ggsave("./output/figures_exported/aspen_stem_count/emm_stemcnt_TCxFxB_ac.pdf", width = 4.5, height = 3.55) # save plot
ggsave("./output/figures_exported/aspen_stem_count/emm_stemcnt_TCxFxB_ac.png", width = 4.5, height = 3.55) # save plot

# table of contrasts
emmeans(stmod_stally_TCxFxB_ac, ~ time_class | burned | fenced) %>% 
  pairs(reverse = TRUE) %>%  
  as_tibble() %>% 
  mutate(across(where(is.numeric),exp)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale")

```

## Aspen Stem Count - Noncore Winter Range (ANC)

### Modeling and Posterior Description

```{r}
stmod_anc_stally1 <- stan_glmer(stem_tally ~ time_class + (1 | site_id), data = asp.anc,
                      family=poisson,
                      iter = 8000,
                      seed = 1234
                      )

```

```{r}
# get posteriors
posteriors.anc.stmcnt <- insight::get_parameters(stmod_anc_stally1)
# tidy
posteriors.anc.stcnt.tidy <- posteriors.anc.stmcnt %>%
  pivot_longer(cols = starts_with("time"), names_to = "parameter", values_to = "values")
```

**Prior summary**: stem_tally ~ time_class + (1 | site_id)  
```{r}
prior_summary(stmod_anc_stally1)
```

```{r}
## model summary
fun.model.summary(stmod_anc_stally1)
```

```{r, echo=FALSE, eval = FALSE}
## shiny eval
launch_shinystan(stmod_anc_stally1)
```

```{r}
#### PP check plot
ppc.plot(stmod_anc_stally1) +
  labs(caption = "ANC: stem_tally~TC")
ggsave("./output/figures_exported/aspen_stem_count/ppc_willowht_anc.png", dpi = 300, width = 4.75, height = 3.75)
ggsave("./output/figures_exported/aspen_stem_count/ppc_willowht_anc.pdf", width = 4.75, height = 3.75)
```

```{r}
# library(ggridges)

posteriors.anc.stcnt.tidy %>% 
  ggplot() +
  geom_density(aes(fill = parameter, x = values, color = parameter), alpha = .12) +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  theme_minimal()

# save plots
ggsave("./output/figures_exported/aspen_stem_count/postden_stemcnt_anc.png", dpi = 300, width = 4.75, height = 3.75)
ggsave("./output/figures_exported/aspen_stem_count/postden_stemcnt_anc.pdf", width = 4.75, height = 3.75)
```

### Effect Existence and Significance Testing
```{r}
fun.sexit.gt(stmod_anc_stally1)

fun.sexit.gt(stmod_anc_stally1) %>%
 tab_header(title = "Aspen Count Posterior Description", subtitle = "Noncore Winter Range") %>%
  gt::gtsave(filename = "./output/tables/aspen_stem_count/xpostdesc_stemcnt_TCxF_anc.rtf")

```


```{r, echo=FALSE, eval=FALSE}
### Model posterior parameters to table
fun.desc.post1.gt.rope(stmod_anc_stally1) %>% 
  tab_header(title = "Aspen Count Posterior Description", subtitle = "Noncore Winter Range")

## save as RTF
fun.desc.post1.gt.rope(stmod_anc_stally1) %>%
 tab_header(title = "Aspen Count Posterior Description", subtitle = "Noncore Winter Range") %>%
  gt::gtsave(filename = "./output/tables/aspen_stem_count/postdesc_stemcnt_TCxF_anc.rtf")

```

#### Probability of Direction

```{r}
# Visualize the pd
#### PD - ANC
p_direction(stmod_anc_stally1) %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Aspen stem count, Noncore winter range")

# save plots
ggsave("./output/figures_exported/aspen_stem_count/pd_stemcnt_anc.png", dpi = 300, width = 4.75, height = 3.75)
ggsave("./output/figures_exported/aspen_stem_count/pd_stemcnt_anc.pdf", width = 4.75, height = 3.75)

```

#### Region of Practical Equivalence (ROPE)

```{r}
rope.stemcnt.ac <- rope(stmod_anc_stally1, ci=0.9) %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 1,
    suffixing = TRUE
  ) 

rope.stemcnt.ac %>% 
  gt::gtsave(filename = "./output/tables/aspen_stem_count/rope_stemcnt_anc.rtf")

```

### Contrasts

```{r}
#### Contrasts
# EMMs for later factor levels are subtracted from those for earlier levels; if you want the comparisons to go in the other direction, use pairs(pigs.emm.s, reverse = TRUE).

emmeans(stmod_anc_stally1, ~ time_class) %>% 
  pairs(reverse = TRUE) %>% 
  plot(type = "response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Noncore Winter Range, Aspen Stem Count")

ggsave("./output/figures_exported/aspen_stem_count/emm_stemcnt_TCxF_anc.pdf", width = 4.5, height = 3.55) # save plot
ggsave("./output/figures_exported/aspen_stem_count/emm_stemcnt_TCxF_anc.png", width = 4.5, height = 3.55) # save plot

# table of contrasts. Results are given on the log (not the response) scale.
emmeans(stmod_anc_stally1, ~ time_class) %>% 
  pairs(reverse = TRUE) %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log (not the response) scale.")

# table of contrasts. Results are given on the response scale.
emmeans(stmod_anc_stally1, ~ time_class) %>% 
  pairs(reverse = TRUE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale.")

# For EMM summaries with type = "response", the tests and confidence intervals are done before back-transforming. The ratios estimated here are actually ratios of geometric means. In general, a model with a log response is in fact a model for relative effects of any of its linear predictors, and this back-transformation to ratios goes hand-in-hand with that.

```


```{r, echo=TRUE, eval=TRUE}
# comment 2.2 "provide analyses on single graphic...
## Plot combinations
contr.plot.ac <- emmeans(stmod_stally2, ~ time_class | fenced) %>% 
  pairs(reverse = TRUE) %>% 
  # plot(type = "response") +
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(title= "Core winter range", caption = "", x = "Estimate", y = "Contrast")

contr.plot.anc <- emmeans(stmod_anc_stally1, ~ time_class) %>% 
  pairs(reverse = TRUE) %>% 
  plot(type = "response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(title= "Noncore winter range", caption = "Point estimate displayed: median 
HPD interval probability: 0.95, Results on log scale", x = "Estimate", y = "")

# patchwork
contr.plot.ac + contr.plot.anc 

ggsave("./output/figures_exported/revised_figs_aspen/emm_stemcnt_AC_ANC.png", width = 8, height = 3.75, dpi=300)
ggsave("./output/figures_exported/revised_figs_aspen/emm_stemcnt_AC_ANC.pdf", width = 8, height = 3.75)

```

## Aspen Stem Count - Kawuneeche Valley (AK)

### Modeling and Posterior Description

```{r}
stmod_ak_stally1 <- stan_glmer(stem_tally ~ time_class + (1 | site_id), data = asp.ak,
                       family=poisson,
                      iter = 8000,
                      seed = 1234
                      )
```

**Prior summary**: stem_tally ~ time_class + (1 | site_id)  
```{r}
prior_summary(stmod_ak_stally1)
```

```{r}
## model summary
fun.model.summary(stmod_ak_stally1)
```


```{r}
# extract posteriors
posteriors.ak.stcnt <- insight::get_parameters(stmod_ak_stally1)

# tidy
posteriors.ak.stcnt.tidy <- posteriors.ak.stcnt %>%
  pivot_longer(cols = starts_with("time"), names_to = "parameter", values_to = "values")
```

```{r}
prior_summary(stmod_ak_stally1)
```

```{r}
#### PP check plot
ppc.plot(stmod_ak_stally1) +
  labs(caption = "AK: stem_tally~TC")
ggsave("./output/figures_exported/aspen_stem_count/ppc_willowht_ak.png", dpi = 300, width = 4.75, height = 3.75)
ggsave("./output/figures_exported/aspen_stem_count/ppc_willowht_ak.pdf", width = 4.75, height = 3.75)
```

```{r}
# library(ggridges)

posteriors.ak.stcnt.tidy %>% 
  ggplot() +
  geom_density(aes(fill = parameter, x = values, color = parameter), alpha = .12) +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  theme_minimal()

# save plots
ggsave("./output/figures_exported/aspen_stem_count/postden_stemcnt_ak.png", dpi = 300, width = 4.75, height = 3.75)
ggsave("./output/figures_exported/aspen_stem_count/postden_stemcnt_ak.pdf", width = 4.75, height = 3.75)
```

### Effect Existence and Significance Testing  
```{r}
fun.sexit.gt(stmod_ak_stally1)

fun.sexit.gt(stmod_ak_stally1) %>%
  tab_header(title = "Aspen Count Posterior Description", subtitle = "Kawuneeche Valley") %>%
  gt::gtsave(filename = "./output/tables/aspen_stem_count/xpostdesc_stemcnt_TCxF_anc.rtf")

```

```{r, ech=FALSE, eval=FALSE}
### Model posterior parameters to table
fun.desc.post1.gt.rope(stmod_ak_stally1) %>% 
  tab_header(title = "Aspen Count Posterior Description", subtitle = "Kawuneeche Valley")

## save as RTF
fun.desc.post1.gt.rope(stmod_anc_stally1) %>% 
  tab_header(title = "Aspen Count Posterior Description", subtitle = "Kawuneeche Valley") %>%
  gt::gtsave(filename = "./output/tables/aspen_stem_count/postdesc_stemcnt_TCxF_anc.rtf")

```

#### Probability of Direction  
```{r}
# Visualize the pd
#### PD - ANC
p_direction(stmod_ak_stally1) %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Aspen stem count, Noncore winter range")

# save plots
ggsave("./output/figures_exported/aspen_stem_count/pd_stemcnt_ak.png", dpi = 300, width = 4.75, height = 3.75)
ggsave("./output/figures_exported/aspen_stem_count/pd_stemcnt_ak.pdf", width = 4.75, height = 3.75)

```

#### Region of Practical Equivalence (ROPE)

```{r}
rope.stemcnt.ak <- rope(stmod_ak_stally1, ci=0.9) %>% 
  gt() %>%
  tab_header(title = "Percent in ROPE") %>% 
  fmt_number(
    columns = 5,
    decimals = 1,
    suffixing = TRUE
  ) 

rope.stemcnt.ak %>% 
  gt::gtsave(filename = "./output/tables/aspen_stem_count/rope_stemcnt_ak.rtf")

```

### Contrasts

```{r}
#### Contrasts
# EMMs for later factor levels are subtracted from those for earlier levels; if you want the comparisons to go in the other direction, use pairs(pigs.emm.s, reverse = TRUE).

emmeans(stmod_ak_stally1, ~ time_class) %>% 
  pairs(reverse = TRUE) %>% 
  plot(type = "response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Noncore Winter Range, Aspen Stem Count")

ggsave("./output/figures_exported/aspen_stem_count/emm_stemcnt_TC_anc.pdf", width = 4.5, height = 3.55) # save plot
ggsave("./output/figures_exported/aspen_stem_count/emm_stemcnt_TC_anc.png", width = 4.5, height = 3.55) # save plot

# table of contrasts. Results are given on the log (not the response) scale.
emmeans(stmod_ak_stally1, ~ time_class) %>% 
  pairs(reverse = TRUE) %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log (not the response) scale.")

# table of contrasts. Results are given on the response scale.
emmeans(stmod_ak_stally1, ~ time_class) %>% 
  pairs(reverse = TRUE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale.")

# For EMM summaries with type = "response", the tests and confidence intervals are done before back-transforming. The ratios estimated here are actually ratios of geometric means. In general, a model with a log response is in fact a model for relative effects of any of its linear predictors, and this back-transformation to ratios goes hand-in-hand with that.

```

```{r}
## version for plot combination
# ANC 
emmeans(stmod_ak_stally1, ~ time_class) %>% 
  pairs(reverse = TRUE) %>% 
  plot(type = "response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "Noncore Winter Range, Aspen Stem Count")
```


# Session info

R version 4.0.3 (2020-10-10)\
Platform: x86_64-w64-mingw32/x64 (64-bit)\
Running under: Windows 10 x64 (build 19041)  
Run date: `r format(Sys.time(), '%Y %B %d')`  

```{r, echo=FALSE}
subset(data.frame(sessioninfo::package_info()), attached==TRUE, c(package, loadedversion)) %>% 
  gt() %>% 
  tab_header(title = "Attached packages and versions")

```

# References