---
title: "EVMP covariate analysis"
author: ""
date: ""
output: 
  html_document: 
    theme: journal
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(knitr)
opts_knit$set(root.dir=normalizePath('../')) # this is required if Rmd is nested below the project directory
opts_chunk$set(fig.path = "../output/figures/") # corrected path and added dev. Needed to specify a subdirectory for figs
```

_Code author:_ 
E. Gage  
mycologica@gmail.com  
https://github.com/egage/EVMP    

```{r, echo=FALSE}
# package install
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(fs))
suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(janitor))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(mapview))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(ggstance))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(skimr)) 
suppressPackageStartupMessages(library(dataMaid))
library(lubridate)
library(mapview)
library(visdat)
library(gt)
```

# Introduction  
This document contains code for running ancillary analyses of Elk
Vegetation Management Plan (EVMP) data collected through the 2018
sampling season. Analyses are limited to macroplot data from plots
established in riparian areas in areas of the elk winter range and
Kawuneeche Valley (i.e., "willow" plots, see Zeigenfuss et al. 2011) and
is provided as a supplement to the NPS NRR Report "Monitoring of
Vegetation Response to Elk Population and Habitat Management in Rocky
Mountain National Park - Analysis of Elk Vegetation Management Plan
monitoring Data: 2008--2018". For information on sampling and the
broader analysis and interpretation of the results, refer to this
report, past analyses [@zeigenfuss2015], and the original EVMP
monitoring plan [@zeigenfuss2011].

# Methods

Weather and ungulate population data were obtained and explored for use as covariates in analyses of EVMP data. Weather data were obtained from were obtained from the National Climate Data Center for Estes Park (Estes Park 3 SSE, CO US, #52761). Snow depth data were acquired from the Bear Lake Snotel Site (Site #322) located southwest of the core elk winter range. Climate data were analyzed for reference to EVMP data and used to examine the relationship of weather and elk population to EVMP data. Elk population estimates produced by RMNP and outside collaborators (T. Hobbs and H. Abouelezz, unpublished data), produced using a combination of aerial and ground surveys, were used to create annual estimates of the winter population size of elk. A linear model estimated using OLS was used to predict median winter elk as a function of annual precipitation data. 

# Results  

### Climate

#### Estes Park

Station: GHCND:USC00052761

```{r}
# Dataset	Daily Summaries
# Order Start Date	1998-08-01 00:00
# Order End Date	2020-08-22 23:59
# Output Format	Custom GHCN-Daily CSV
# Data Types	MDPR, DAPR, PRCP, SNWD, SNOW, TMAX, TMIN, TOBS
# Custom Flag(s)	Station Name
# Units	Metric
# Stations/Locations	ESTES PARK 3 SSE, CO US (Station ID: GHCND:USC00052761)
```


```{r}

## import
ep.raw <- read_csv("./data/covariates/climate/NCDC_EstesPark/2257130.csv") ## METRIC UNITS

# The five core values are:
# PRCP = Precipitation (mm or inches as per user preference, inches to hundredths on Daily Form pdf file)
# SNOW = Snowfall (mm or inches as per user preference, inches to tenths on Daily Form pdf file)
# SNWD = Snow depth (mm or inches as per user preference, inches on Daily Form pdf file)
# TMAX = Maximum temperature (Fahrenheit or Celsius as per user preference, Fahrenheit to tenths on
# Daily Form pdf file
# TMIN = Minimum temperature (Fahrenheit or Celsius as per user preference, Fahrenheit to tenths on
# Daily Form pdf file

# Processing Completed	2020-08-24
# Stations	
# GHCND:USC00052761
# Begin Date	1998-08-01 00:00
# End Date	2020-08-22 23:59
# Data Types	
# MDPRDAPRPRCPSNWDSNOWTMAXTMINTOBS
# Units	
# Metric
# Custom Flag(s)	
# Station Name
# Eligible for Certification	No

```

```{r}
ep.cln <- ep.raw %>% 
  clean_names() %>% 
  mutate(month = month(date, label = TRUE, abbr = TRUE)) %>% 
  mutate(doy = lubridate::yday(date)) %>%
  mutate(year = lubridate::year(date))

```

** Annual Precipitation**
```{r}

ep.annual.sum.prcp <- ep.cln %>% 
  group_by(year) %>% 
  summarise(ppt_mm_ann = sum(prcp)) %>% 
  ungroup()

ep.plt1 <- ep.annual.sum.prcp %>%
  filter(year > 2007 & year < 2019) %>% 
  mutate(year = as.integer(year)) %>% 
  ggplot(aes(year,ppt_mm_ann)) +
  # geom_line() +
  # geom_point() +
  geom_col(fill = 'grey40') +
  theme_minimal() +
  scale_x_continuous(breaks = seq(from = 2008, to = 2018, by = 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Year", y = "Annual precipitation (mm)", caption = "Estes Park Station GHCND:USC00052761")

ep.plt1
# ggsave("./output/figures_exported/EstesPark_AnnualPrcp.png", width = 3.85, height = 3.85)

```

**Daily Maximum Temperature**
```{r}
# temperature
ep.annual.mean.tmax <- ep.cln %>% 
  group_by(year) %>% 
  summarise(temp_mean_ann = mean(tmax, na.rm=TRUE)) %>% 
  ungroup()

ep.cln %>%
  filter(year > 2007 & year < 2019) %>% 
  mutate(year = as.character(year)) %>% 
  ggplot(aes(year,tmax)) +
  # geom_line() +
  # geom_point() +
  geom_boxplot() +
  # geom_col(fill = 'grey40') +
  theme_minimal() +
  # scale_x_continuous(breaks = seq(from = 2008, to = 2018, by = 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Year", y = "Maxiumum daily temp (C)", caption = "Estes Park Station GHCND:USC00052761")

```


#### Bear Lake Snotel
```{r}
# data\covariates\climate\BearToothSnotel
snotel <- read_csv("./data/covariates/climate/BearLakeSnotel/snotel_2read.csv") %>% 
  janitor::clean_names()
```

```{r}
snotel <- snotel %>% 
  rename(swe = snow_water_equivalent_in_start_of_day_values) %>% 
  rename(ppt_accum_in = precipitation_accumulation_in_start_of_day_values) %>% 
  rename(temp_max_f = air_temperature_maximum_deg_f) %>%
  rename(temp_min_f = air_temperature_minimum_deg_f) %>%
  rename(temp_avg_f = air_temperature_average_deg_f) %>%
  rename(ppt_incr_in = precipitation_increment_in) 

snotel <- snotel %>%
  mutate(date = mdy(date)) %>% 
  mutate(month = month(date,label = TRUE),
         doy = yday(date),
         year = year(date))

```

```{r}

snotel <- snotel %>% 
  mutate(ppt_accum_mm = 25.4*ppt_accum_in) %>% 
  mutate(ppt_incr_mm = 25.4*ppt_incr_in)

# snotel %>%
#   visdat::vis_dat()
```

```{r}
# What day had the latest peak swe?
top_ppt_accum <- snotel %>% 
  group_by(year) %>% 
  summarise(ppt_accum_mm_max = max(ppt_accum_mm)) %>% 
  ungroup()

ppt.pl <- top_ppt_accum %>% 
  filter(year >1980 & year < 2019) %>% 
  # mutate(year = as.character(year)) %>%
  ggplot(aes(year,ppt_accum_mm_max)) +
  geom_col(fill = "ivory4") +
  labs(x="", y = "Accumulated precipitation (mm)", caption = "Bear Lake Snotel Data") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_continuous(breaks = seq(from = 1980, to = 2018, by = 5))
  
# ppt.pl

```

```{r}
top_swe <- snotel %>% 
  mutate(year = as.character(year)) %>%
  group_by(year) %>% 
  summarise(sum_swe = sum(swe)) %>% 
  ungroup()

swe.pl <- top_swe %>%
  filter(year>1980 & year <2019) %>% 
  # mutate(year = as.character(year)) %>%
  mutate(year = as.integer(year)) %>%
  ggplot(aes(year,sum_swe)) +
  geom_col(fill = "grey60") +
  # geom_jitter() +
  labs(x="", y = "SWE") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_continuous(breaks = seq(from = 1980, to = 2018, by = 5))

# swe.pl

# ggsave("./output/figures_exported/bearLk_SWE.png", width = 3.85, height = 3.85)

```
**Annual SWE and Accumulated Precipitation**
```{r}
cowplot::plot_grid(swe.pl,ppt.pl, labels = "AUTO")
# ggsave("./output/figures_exported/bear_lk_swe_ppt.png", width = 6.5, height = 3.8)
```


```{r}
top_swe %>% 
  select(year, sum_swe) %>% 
  datatable(rownames = FALSE, filter = "top")
```



```{r, echo=FALSE, eval=FALSE}

### Physiographic
# distance from stream



```

### Elk
```{r}
elk.raw <- read_csv("./data/covariates/elk/RMNP_Population_summary.csv") %>% 
  janitor::clean_names()

```

```{r}
elk.raw <- elk.raw %>% 
  mutate(year = as.integer(year))

## write to disk
# write_csv(elk.raw, "./data/covariates/elk/RMNP_elk_pop_summary_cln.csv")

elk.raw %>% 
  ggplot(aes(x = year, y = median)) +
  geom_point(color = "blue") +
  geom_line(color = "blue") +
  geom_line(aes(y = p025), color = "red", lty = "dashed") +
  geom_line(aes(y = p975), color = "red", lty = "dashed") +
  theme_minimal() +
  labs(x = "Year", y = "Median elk populatiop estimate", caption = "Modeled elk pop: RMNP_Population_summary.csv ")

# ggsave("./output/figures_exported/elk_population.png", width = 5, height = 4, dpi = 300)
```

```{r}
elk.raw %>% 
  gt::gt() %>% 
  gt::gtsave("./output/tables/elkpop.rtf")

elk.raw %>% 
  summarytools::descr(median) 

elk.raw %>% 
  arrange(desc(median))

elk.raw %>% 
  arrange((median))
```


Modeled elk population estimates (T. Hobbs unpublished data) are negatively correlated with annual precipitation at both the Estes Park and Bear Lake climate stations. The model had moderate explanatory power (R2' median = 0.24, 909% CI [1.07e-05, 0.45], adj. R2 = 0.07). The model's intercept, corresponding to elk = 0 and precipitation = 0, is at 1291.28 (90% CI [788.17, 1818.43], 0.04% in ROPE). Precipitation in the model had a probability of 97.56% of being negative and was medium in size (median = -0.79, 100% in ROPE, std. median = -0.51). The algorithm successfully converged (R^ = 1.001) and estimates were stable (ESS = 1846). A similar trend but weaker correlation occurred with Bear Lake precipitation data. 



```{r}
### elk vs climate
## Hobbs modeled elk vs Estes Park prcp
# elk.raw %>% arrange(desc(year))
elk.ep_prcp <- left_join(ep.annual.sum.prcp,elk.raw, by = "year")

# elk.ep_prcp %>% 
#   ggplot(aes(x = year)) +
#   geom_col(aes(y = ppt_mm_ann)) +
#   geom_point(aes(y = median), color = "red") +
#   labs(x = "", y = "")

pl.cov1 <- elk.ep_prcp %>% 
  ggplot() +
  geom_point(aes(x = ppt_mm_ann, y = median)) +
  # geom_smooth(aes(x = ppt_mm_ann, y = median)) +
  geom_smooth(aes(x = ppt_mm_ann, y = median), method = "glm") +
  theme_minimal() +
  labs(x= "Annual precipitation (mm)", y = "Number of elk", caption = "Hobbs modeled elk vs Estes Park Precip") 

pl.cov1
ggsave("./output/figures_exported/covariates/elk_v_prcp.png", width = 3.75, height = 4)


```

```{r, eval=FALSE, echo=FALSE}
top_swe <- top_swe %>% 
  mutate(year = as.integer(year))


```


```{r}

elk.ep_EPprcp_BLkppt <- left_join(elk.ep_prcp, top_ppt_accum, by = "year")

pl.cov2 <- elk.ep_EPprcp_BLkppt %>% 
  ggplot() +
  geom_point(aes(x = ppt_accum_mm_max , y = median)) +
  # geom_smooth(aes(x = ppt_mm_ann, y = median)) +
  geom_smooth(aes(x = ppt_accum_mm_max , y = median), method = "glm") +
  theme_minimal() +
  labs(x= "Annual precipitation (mm)", y = "Number of elk", caption = "Hobbs modeled elk vs Bear Lake SWE")
pl.cov2

```

A linear model (estimated using OLS) predicting median elk with annual precipitation (formula: median ~ ppt_mm_ann) explains a significant and substantial proportion of variance (R2 = 0.39, F(1, 13) = 8.16, p = 0.013, adj. R2 = 0.34). The model's intercept, corresponding to ppt_mm_ann = 0, is at 1145.02 (95% CI [741.20, 1548.84], t(13) = 6.13, p < .001). Within the model, the effect of ppt_mm_ann is significantly negative (beta = -1.19, 95% CI [-2.09, -0.29], t(13) = -2.86, p < .05; Std. beta = -0.62, 95% CI [-1.09, -0.15]). Standardized parameters were obtained by fitting the model on a standardized version of the data.

```{r}

# lm(median ~ ppt_mm_ann, data = elk.ep_EPprcp_BLkppt)
library(moderndive)
mod.elkXepPrcp <- lm(median ~ ppt_mm_ann, data = elk.ep_EPprcp_BLkppt) 

mod.elkXepPrcp %>% 
  moderndive::get_regression_table() %>% 
  gt() %>% 
  tab_header(title = "lm: elk media ~ Estes Park annual prcp")


## elk ~ Bear lake
mod.elkXblkPrcp <- lm(median ~ ppt_accum_mm_max, data = elk.ep_EPprcp_BLkppt) 

## broom::
broom::tidy(mod.elkXblkPrcp)

library(report)
# mod.elkXblkPrcp
# report::report(mod.elkXepPrcp)
# report::report(mod.elkXblkPrcp)

broom::tidy(mod.elkXblkPrcp) %>% 
  gt() %>% 
  tab_header(title = "lm: elk media ~ Bear Lk annual prcp")

```

Bayesian generalized linear model of median winter elk numbers (Hobbs unpublished) and accumulated annual precipitation at the Estes Park climate station. 
```{r}
# bayesian lm
library(rstanarm)

stanlm.elkXep <- stan_glm(median ~ ppt_accum_mm_max, data = elk.ep_EPprcp_BLkppt,
                         chains = 5, iter = 1000,
                         seed = 12345,
                         refresh = 0)

```

A Bayesian linear model estimated using MCMC sampling predict median elk with accumulated annual precipitation had moderate explanatory power (R2 = 0.24, 89% CI [1.07e-05, 0.45], adj. R2 = 0.07). The model's intercept, corresponding to precipitation = 0, is at 1291 (95% CI [619.77, 1912.85]). Within this model the effect of precipitation (Median = -0.79, 0.95% CI [-1.54, -0.01]) had a 97.56% probability of being negative (< 0), but a low probability of being significant or large. The estimation successfully converged (Rhat = 1.001) and the indices are reliable (ESS = 1846).


```{r, echo=FALSE}
plot(stanlm.elkXep, "hist", pars = c("ppt_accum_mm_max", "sigma"), 
     transformations = list(sigma = "log"))

# report::report(stanlm.elkXep )

```


```{r}
## plot grid
cowplot::plot_grid(pl.cov1, pl.cov2, labels = "AUTO")
# ggsave("./output/figures_exported/covariates/prcp_x_elk_2panel.png", width = 6.5, height = 3.75, dpi = 300)
```

```{r, echo=FALSE, eval=FALSE}

### elk vs offtake upland


### climate vs elk

### Moose


```



# Session info

R version 4.0.3 (2020-10-10)  
Platform: x86_64-w64-mingw32/x64 (64-bit)  
Running under: Windows 10 x64 (build 19041)  
Run date: `r format(Sys.time(), '%Y %B %d')`  

```{r, echo=FALSE}
subset(data.frame(sessioninfo::package_info()), attached==TRUE, c(package, loadedversion)) %>% 
  gt() %>% 
  tab_header(title = "Attached packages and versions")

```

# References

