---
title: "Appendix C - Willow Line Intercept Data"
author: ''
date: ''
output:
  html_document:
    theme: journal
    toc: yes
    toc_float: no
    toc_depth: 3
    fig_caption: yes
  word_document:
    toc: yes
    toc_depth: '3'
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = 'center', cache = FALSE)
opts_knit$set(root.dir=normalizePath('../')) # this is required if Rmd is nested below the project directory
opts_chunk$set(fig.path = "../output/figures/") 
set.seed(1234)
```

**Willow plots: Line intercept data**

*Code author:* E. Gage
[mycologica\@gmail.com](mailto:mycologica@gmail.com){.email}\
<https://github.com/egage>\
Updated: `r format(Sys.time(), '%Y %B %d')`

```{r, echo=FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(fs))
suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(janitor))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(mapview))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(ggstance))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(skimr)) 
suppressPackageStartupMessages(library(dataMaid))
library(lubridate)
library(mapview)
library(bayestestR)
library(gt)
library(labelled)
library(summarytools)
library(emmeans)
library(rstanarm)
options(mc.cores = parallel::detectCores())

```

```{r, echo=FALSE, warning=FALSE,message=FALSE, comment=FALSE, eval=TRUE}
# function definitions
## themes
theme_smFacet <- theme_minimal() + theme(strip.text.x = element_text(size = 7))
theme_smFacet2 <- theme_minimal() + theme(strip.text.x = element_text(size = 6))
## data import
read_sheets <- function(file){
  xlsx_file <- file
  xlsx_file %>%
    excel_sheets() %>%
    set_names() %>%
    map_df(read_excel, path = xlsx_file, .id = 'sheet_name', trim_ws = TRUE, skip = 1, col_types = "text", range = "A2:DW700") %>% 
    mutate(file_name = file) %>% 
    select(file_name, sheet_name, everything())
}

## read xlsx then csv cache
read_then_csv <- function(sheet, path) {
  pathbase <- path %>%
    basename() %>%
    tools::file_path_sans_ext()
  path %>%
    read_excel(sheet = sheet) %>% 
    write_csv(paste0(pathbase, "-", sheet, ".csv"))
}

## plotting
ggTile_yr_season_site2 <- function(df){
  df %>%
  group_by(yr, season, site2) %>%
  summarise(n= n()) %>%
  ggplot(aes(yr, site2)) +
  geom_tile(aes(fill = n), color = 'white') +
  viridis::scale_fill_viridis(option = "B") +
  facet_wrap(~season) +
  theme(axis.text.x = element_text(angle = 55, hjust = 1, size = 7)) +
  theme(axis.text.y = element_text(size = 7)) +
  facet_wrap(~season)}

ggTile_yr_season_site <- function(df){
  df %>%
  group_by(yr, season, site) %>%
  summarise(n= n()) %>%
  ggplot(aes(yr, site)) +
  geom_tile(aes(fill = n), color = 'white') +
  viridis::scale_fill_viridis(option = "D") +
  theme(axis.text.x = element_text(angle = 55, hjust = 1, size = 7)) +
  theme(axis.text.y = element_text(size = 7)) +
  facet_wrap(~season)} 

## ppcheck.plot
ppc.plot <- function(model){
  pp_check(model) +
  labs(x = "Value", y = "Density") +
  theme_minimal()
}

### Table functions
## gt field names
names.gt <- function(df){
  df %>% 
  names() %>% enframe %>% gt::gt() %>% 
    tab_header(title = "field names")
}

names.dt <- function(df){
  df %>% 
  names() %>% enframe %>% 
    DT::datatable(caption = "field names")
}

# describe posterior
fun.desc.post1.gt.rope <- function(model){
  describe_posterior(
  model,
  effects = "fixed",
  component = "all",
  ci_method = "hdi",
  ci=0.9,
  test = c("p_direction", "p_significance","rope"),
  #rope_range(c(-0.1,0.1)),
  rope_range("default"),
  rope_ci = 0.9,
  centrality = "median"
) %>% 
    as_tibble() %>% 
    gt() %>% 
    tab_header(title = "Posterior description") %>% 
    fmt_number(
    columns = 2:14,
    decimals = 2,
    suffixing = TRUE
  )
}

## summary funcs
mean_r1 <- function(x){
  round(mean(x, na.rm = TRUE),1)
}

sd_r1 <- function(x){
  round(sd(x, na.rm = TRUE),1)
}

range_r1 <- function(x){
  round(range(x, na.rm = TRUE),1)
}


```

# Introduction

This document is an appendix to NRR report "Monitoring of Vegetation
Response to Elk Population and Habitat Management in Rocky Mountain
National Park. Analysis of Elk Vegetation Management Plan monitoring
data: 2008--2018" presenting additional analyses of line intercept data
not presented in the main report. The code and analyses aim to
characterize basic data structure, clean and reorganize as needed for
plotting and modeling. Analyses are limited to line intercept data from
plots established in areas of the elk winter range and Kawuneeche Valley
(i.e., see Zeigenfuss et al. 2011). For information on sampling and the
broader analysis and interpretation of the results, refer to this
report, past analyses [@zeigenfuss2015], and the original EVMP
monitoring plan [@zeigenfuss2011].

# Methods

The code and results presented here build on derived data produced in a
separate code document focused on ingesting, compiling, and cleaning raw
data provided by RMNP staff (Appendix B). Extensive "munging" of raw
files was required. Separate code documents were developed for other
distinct elements of the larger EVMP (e.g., aspen and upland plots).\
In general, packages included in the "Tidyverse" ecosystem of R packages
were used for data transformation and visualization [@wickham2019],
although specialized packages particular to specific tasks (e.g., the
"bayesplot" plot for visualization of Bayesian posterior distributions)
were also used [@gabry2019].

Bayesian repeated measures analyses were fit separately for combined
core and non-core winter range, core winter range (WC) and non-core
winter range plots (WNC). For willow height, models were fit using
weekly informative priors and a gamma distribution with the "stan_lmer"
function in the "rstanarm" package [@goodrich2020][@brilleman2018].

For continuous proportions such as plant cover (i.e, those not derived
from count data), a common approach to modeling is to transform data
then use ordinary linear models, however this creates problems for
interpretation and inference [@douma2019]. Techniques relying on
transformations make estimates on the transformed scale, requiring
back‐transformation for reporting and interpretation. However, the
relationship between the original and transformed proportions is often
non‐linear, creating issues for interpretation [@douma2019]. Alternative
approaches using beta regression [@cribari-neto2009][@ferrari2004] have
been developed and are relied on here for modeling continuous
proportions (e.g., willow cover). After fitting appropriate models using
"rstanarm" functions, comparisons of estimated marginal means were made
between time classes and management classes (e.g., fencing) using
functions in the "emm" R package [@lenth2020].

```{r}
## read in data from the derived folder. See appx_
csv.all.lc.li.df <- read_csv("./data/EVMP_derived/line_intercept_willow_cleaned.csv")

```

```{r, eval = FALSE, echo=FALSE}
# > From discussions with HA in the field, the orientation of line-intercept transects may not have had a consistent orientation from sampling interval to interval.
```

```{r}
### Counts of plots in fenced and unfenced contexts
## address missing attributes for pType
csv.all.lc.li.df <- csv.all.lc.li.df %>%
  mutate(pType = case_when(is.na(pType) ~ "willow",
                            TRUE ~ pType))

## address missing attributes for RANGE_TYPE
csv.all.lc.li.df <- csv.all.lc.li.df %>%
  mutate(RANGE_TYPE = case_when(is.na(RANGE_TYPE)  & SITE_TYPE == "WC" ~ "core winter range",
                            TRUE ~ RANGE_TYPE))

```

```{r}
## Make NA unburned; presumes NA are "unburned"
## standardize encoding of BURNED
csv.all.lc.li.df <- csv.all.lc.li.df %>%
  mutate(BURNED = case_when(is.na(BURNED) ~ "Unburned",
                            BURNED == "Not burned" ~ "Unburned",
                            TRUE ~ BURNED))

## reverse factor levels
csv.all.lc.li.df <- csv.all.lc.li.df %>%
  mutate(timeClass = forcats::fct_rev(timeClass))

## Reorder levels
csv.all.lc.li.df <- csv.all.lc.li.df %>% 
  mutate(SITE_TYPE = as_factor(SITE_TYPE)) %>% 
  mutate(SITE_TYPE = forcats::fct_relevel(SITE_TYPE, "WK", after = 2))

csv.all.lc.li.df <- csv.all.lc.li.df %>% 
  mutate(timeClass = fct_relevel(timeClass, "BL", after=Inf)) %>% 
  mutate(timeClass = fct_rev(timeClass)) %>% 
  mutate(FENCED = fct_rev(FENCED))

```

```{r}
## Address missing RANGE_TYPE attribution
csv.all.lc.li.df <- csv.all.lc.li.df %>%
  mutate(RANGE_TYPE = case_when(SITE_TYPE == "WC" ~ "core winter range",
                                TRUE ~ RANGE_TYPE)) %>% 
  distinct()
```

# Results

## Maximum Height

### Descriptive Statistics and plots

#### All Shrubs Species

```{r}
## Boxplot
csv.all.lc.li.df %>% 
  filter(pType == "willow") %>%
  mutate(yr = as.character(yr)) %>%
  filter(yr == 2008 | yr == 2013 | yr == 2018) %>%
  ggplot(aes(timeClass, MAX_HEIGHT_CM)) +
  geom_boxplot(aes(fill = FENCED), outlier.shape = NA) +
  geom_hline(aes(yintercept = 110), color = "red", lty = "dashed", size = 1) +
  ylim(0,300) +
  labs(x="", y= "Height (cm)", title = "Maximum shrub height", caption = "WC, All shrub species line intercept plots") +
  scale_fill_manual(values = c("grey90","grey50")) +
  theme_minimal() + 
  # theme(axis.text.x=element_text(angle = 45, hjust = 1)) #+
  facet_wrap(~SITE_TYPE)

# ggsave("./output/figures_exported/WCWNCWK_LI_shrubHt_boxplot.png", width = 6.5, height = 4.875)
```

```{r}
#### Table of mean max height
csv.all.lc.li.df %>% 
  filter(timeClass == "BL" | timeClass == "2013" | timeClass == "2018") %>% 
  group_by(timeClass, SITE_TYPE) %>%
  descr(MAX_HEIGHT_CM, stats = "common") %>% 
  summarytools::tb() %>% 
  mutate(across(c('mean','sd','pct.valid'), ~round(.,digits = 1))) %>% 
  gt() %>% 
  tab_header(title = "Max height", subtitle = "All shrub species combined, all site and mgt group types")

```

```{r}
## all shrub species
csv.all.lc.li.df %>%
  filter(pType == "willow") %>% 
  filter(yr == 2008 | yr == 2013 | yr == 2018) %>%
  filter(!is.na(MAX_HEIGHT_CM)) %>%
  filter(SITE_TYPE == "WC") %>% 
  ggplot(aes(timeClass, MAX_HEIGHT_CM)) +
  geom_boxplot(aes(fill=FENCED), color = 'black', outlier.shape = NA) +
  geom_hline(aes(yintercept = 110), color = "red", lty = "dashed", size = 1) +
  scale_fill_viridis(discrete = TRUE) +
  theme_minimal() +
  ylim(0,320) +
  scale_fill_manual(values = c("grey90","grey50")) +
  labs(x = "Year", y = "Max height (cm)", fill = "", title = "Core Winter Range Plots: All Shrub Species", caption = "Line intercept data")

```

```{r}
### All shrubs, site type, fenced/unfenced
#### Table of mean max height
csv.all.lc.li.df %>% 
  filter(timeClass == "BL" | timeClass == "2013" | timeClass == "2018") %>% 
  group_by(timeClass, SITE_TYPE, FENCED) %>%
  descr(MAX_HEIGHT_CM, stats = "common") %>% 
  summarytools::tb() %>% 
  mutate(across(c('mean','sd','pct.valid'), ~round(.,digits = 1))) %>% 
  gt() %>% 
  tab_header(title = "LI, Mean max height", subtitle = "All shrubs, all site types, fenced/unfenced")

```

```{r}
### All shrubs, site type, fenced/unfenced, burned/unburned
## Boxplot
csv.all.lc.li.df %>% 
  filter(pType == "willow") %>%
  mutate(yr = as.character(yr)) %>%
  filter(yr == 2008 | yr == 2013 | yr == 2018) %>%
  ggplot(aes(timeClass, MAX_HEIGHT_CM)) +
  geom_boxplot(aes(fill = FENCED), outlier.shape = NA) +
  geom_hline(aes(yintercept = 110), color = "red", lty = "dashed", size = 1) +
  ylim(0,300) +
  labs(x="", y= "Height (cm)", title = "Maximum shrub height", caption = "WC, All shrub species line intercept plots") +
  scale_fill_manual(values = c("grey90","grey50")) +
  theme_minimal() + 
  # theme(axis.text.x=element_text(angle = 45, hjust = 1)) #+
  facet_wrap(BURNED~SITE_TYPE)

# ggsave("./output/figures_exported/WCWNCWK_LI_shrubHt_FenBur_boxplot.png", width = 6.5, height = 4.875)
```

#### Willow Species
```{r}
csv.wil.lc.li.df <- csv.all.lc.li.df %>% 
  filter(str_detect(SPECIES_CODE, "^SA")) %>% 
  mutate(FENCED = fct_rev(FENCED))

## Boxplot
csv.wil.lc.li.df %>% 
  filter(pType == "willow") %>%
  mutate(yr = as.character(yr)) %>%
  filter(yr == 2008 | yr == 2013 | yr == 2018) %>%
  ggplot(aes(timeClass, MAX_HEIGHT_CM)) +
  geom_boxplot(aes(fill = FENCED), outlier.shape = NA) +
  geom_hline(aes(yintercept = 110), color = "red", lty = "dashed", size = 1) +
  ylim(0,600) +
  labs(x="", y= "Height (cm)", title = "Maximum willow height", caption = "WC, All willow species line intercept plots") +
  scale_fill_manual(values = c("grey90","grey50")) +
  theme_minimal() + 
  facet_wrap(~SITE_TYPE)

```

```{r}
### All willow species, site type, fenced/unfenced
#### Table of mean max height
csv.wil.lc.li.df %>% 
  filter(timeClass == "BL" | timeClass == "2013" | timeClass == "2018") %>% 
  group_by(timeClass, SITE_TYPE, FENCED) %>%
  descr(MAX_HEIGHT_CM, stats = "common") %>% 
  summarytools::tb() %>% 
  mutate(across(c('mean','sd','pct.valid'), ~round(.,digits = 1))) %>% 
  gt() %>% 
  tab_header(title = "Maximum height", subtitle = "All willows combined, all site types, fenced/unfenced")

```

```{r}
### All Willow Species, site type, fenced/unfenced, burned/unburned
## Boxplot
csv.wil.lc.li.df %>% 
  filter(pType == "willow") %>%
  mutate(yr = as.character(yr)) %>%
  filter(yr == 2008 | yr == 2013 | yr == 2018) %>%
  ggplot(aes(timeClass, MAX_HEIGHT_CM)) +
  geom_boxplot(aes(fill = FENCED), outlier.shape = NA) +
  geom_hline(aes(yintercept = 110), color = "red", lty = "dashed", size = 1) +
  ylim(0,600) +
  labs(x="", y= "Height (cm)", title = "Maximum shrub height", caption = "WC, All willow species line intercept plots") +
  scale_fill_manual(values = c("grey90","grey50")) +
  theme_minimal() + 
  facet_wrap(BURNED~SITE_TYPE)

```

```{r}
#### Table of mean max height
csv.wil.lc.li.df %>% 
  filter(timeClass == "BL" | timeClass == "2013" | timeClass == "2018") %>% 
  group_by(timeClass, SITE_TYPE, FENCED, BURNED) %>%
  descr(MAX_HEIGHT_CM, stats = "common") %>% 
  summarytools::tb() %>% 
  mutate(across(c('mean','sd','pct.valid'), ~round(.,digits = 1))) %>% 
  gt() %>% 
  tab_header(title = "Maximum height", subtitle = "All willow, all site types, fenced/unfenced, burned/unburned")

```

```{r}
csv.all.lc.li.df %>%
  filter(yr !=2009 & yr != 2015 & yr !=2017) %>% 
  filter(SITE_TYPE != "WK") %>%
  filter(SPECIES_CODE == "SAMO" | SPECIES_CODE == "SAGE" | SPECIES_CODE == "SAPL" ) %>% 
  filter(!is.na(MAX_HEIGHT_CM)) %>% 
  group_by(SPECIES_CODE, timeClass, FENCED) %>% 
  descr(MAX_HEIGHT_CM, stats = "common") %>% 
  tb() %>% 
  select(-variable) %>% 
  mutate(across(where(is.numeric), round, 1)) %>% 
  gt() %>% 
  tab_header(title = "Maximum height: common willow species comparisons", subtitle = "Combined WC and WNC; SAPL, SAGE, and SAMO")

```

```{r, eval=TRUE}
## salix monitcola
csv.all.lc.li.df %>%
  filter(pType == "willow") %>% 
  filter(yr == 2008 | yr == 2013 | yr == 2018) %>%
  filter(!is.na(MAX_HEIGHT_CM)) %>%
  filter(SPECIES_CODE =="SAMO") %>% 
  filter(SITE_TYPE == "WC") %>% 
  ggplot(aes(timeClass, MAX_HEIGHT_CM)) +
  geom_boxplot(aes(fill=FENCED), color = 'black', outlier.shape = NA) +
  geom_hline(aes(yintercept = 110), color = "red", lty = "dashed", size = 1) +
  theme_minimal() +
  ylim(0,320) +
  scale_fill_manual(values = c("grey90","grey50")) +
  labs(x = "Year", y = "Max height (cm)", fill = "", title = "Core Winter Range Plots: SAMO only", caption = "Line intercept data")

```

```{r, eval=TRUE}
## salix geyeriana
csv.all.lc.li.df %>%
  filter(pType == "willow") %>% 
  filter(yr == 2008 | yr == 2013 | yr == 2018) %>%
  filter(!is.na(MAX_HEIGHT_CM)) %>%
  filter(SPECIES_CODE =="SAGE") %>% 
  filter(SITE_TYPE == "WC") %>% 
  ggplot(aes(timeClass, MAX_HEIGHT_CM)) +
  geom_boxplot(aes(fill=FENCED), color = 'black', outlier.shape = NA) +
  geom_hline(aes(yintercept = 110), color = "red", lty = "dashed", size = 1) +
  theme_minimal() +
  ylim(0,320) +
  scale_fill_manual(values = c("grey90","grey50")) +
  labs(x = "Year", y = "Max height (cm)", fill = "", title = "Core Winter Range Plots: SAGE only", caption = "Line intercept data")

```

```{r, eval=TRUE}
## Salix planifolia
csv.all.lc.li.df %>%
  filter(pType == "willow") %>% 
  filter(yr == 2008 | yr == 2013 | yr == 2018) %>%
  filter(!is.na(MAX_HEIGHT_CM)) %>%
  filter(SPECIES_CODE =="SAPL") %>% 
  filter(SITE_TYPE == "WC") %>% 
  ggplot(aes(timeClass, MAX_HEIGHT_CM)) +
  geom_boxplot(aes(fill=FENCED), color = 'black', outlier.shape = NA) +
  geom_hline(aes(yintercept = 110), color = "red", lty = "dashed", size = 1) +
  theme_minimal() +
  ylim(0,320) +
  scale_fill_manual(values = c("grey90","grey50")) +
  labs(x = "Year", y = "Max height (cm)", fill = "", title = "Core Winter Range Plots: SAPL only", caption = "Line intercept data")

```

```{r}
csv.all.lc.li.df %>%
  filter(pType == "willow") %>% 
  filter(SPECIES_CODE != "SAXX") %>% 
  filter(yr == 2008 | yr == 2013 | yr == 2018) %>%
  filter(!is.na(MAX_HEIGHT_CM)) %>%
  filter(FENCED == "Unfenced" | FENCED == "Fenced") %>% 
  group_by(yr, SPECIES_CODE, FENCED) %>% 
  summarise(mean.ht = mean(MAX_HEIGHT_CM, na.rm = TRUE)) %>% 
  pivot_wider(names_from = yr, values_from = c(mean.ht)) %>% 
  gt::gt() %>% 
  fmt_number(
    columns = vars('2008','2013','2018'),
    decimals = 1,
    use_seps = FALSE
  ) %>% 
  tab_header("Mean height by shrub species")

```

```{r, echo=FALSE, eval=FALSE}
csv.all.lc.li.df %>%
  filter(pType == "willow") %>% 
  filter(SPECIES_CODE != "SAXX") %>% 
  filter(yr == 2008 | yr == 2013 | yr == 2018) %>%
  filter(!is.na(MAX_HEIGHT_CM)) %>%
  filter(str_detect(SPECIES_CODE, "^SA")) %>%
  filter(FENCED == "Unfenced" | FENCED == "Fenced") %>% 
  group_by(yr, FENCED) %>% 
  summarise(mean.ht = mean(MAX_HEIGHT_CM, na.rm = TRUE), sd.ht = sd(MAX_HEIGHT_CM, na.rm = TRUE)) %>% 
  pivot_wider(names_from = yr, values_from = c(mean.ht,sd.ht)) %>% 
  select(contains("2008"),contains("2013"), contains("2018")) %>% 
  gt::gt() %>% 
  fmt_number(
    columns = vars("mean.ht_2008","mean.ht_2013","mean.ht_2018","sd.ht_2008","sd.ht_2013", "sd.ht_2018"),
    decimals = 1,
    use_seps = FALSE
  ) %>% 
  tab_header("Mean height (cm) by all willow species, line intercept")

```

### Modeling and Posterior Description - Combined Willow Species

```{r, cache=FALSE}
# clean names
csv.wil.lc.li.df <- csv.wil.lc.li.df %>% 
  clean_names() %>% 
  filter(!is.na(max_height_cm)) %>% 
  filter(time_class == "BL" | time_class == "2013" | time_class == "2018")

## run with gamma
wc.wnc.stmod_li_TCxF <- stan_glmer(max_height_cm ~ time_class*fenced +  (1 | site_id), data = csv.wil.lc.li.df,
                      family=Gamma(link="log"),
                      iter = 10000,
                      )
```

```{r, echo=FALSE}
# Get the prior summary
prior_summary(wc.wnc.stmod_li_TCxF)
```

```{r}
#### PP check plot
ppc.plot(wc.wnc.stmod_li_TCxF) +
  labs(caption = "WC+WNC: willow_ht~TCxF")

```

```{r}
#### Describe posterior distributions
fun.desc.post1.gt.rope(wc.wnc.stmod_li_TCxF) %>% 
  tab_header(title = "Combined WC and WNC", subtitle = "Willow Height Posterior Description - Line Intercept")

```

### Probability of Direction

```{r}
p_direction(wc.wnc.stmod_li_TCxF) %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(caption = "Willow height, Combined Core and Non-core Winter Range")

p_direction(wc.wnc.stmod_li_TCxF) %>% 
  as_tibble() %>% 
  gt() %>% 
  tab_header(title = "Probability of Direction", subtitle = "WC & WNC Combined Willows")

```

### Contrasts

```{r}
## contrasts
emmeans(wc.wnc.stmod_li_TCxF, ~ time_class | fenced) %>% 
  pairs(reverse = TRUE) %>% 
  # plot() + # on the lm predict scale
  plot(type = "response") +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "WC & WNC Max Height, Combined Willow, Results are given on the response scale.")

# table of contrasts. Results are given on the response scale.
emmeans(wc.wnc.stmod_li_TCxF, ~ time_class | fenced) %>% 
  pairs(reverse = TRUE, type="response") %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the response scale.")

```

```{r}
## contrasts: 
emmeans(wc.wnc.stmod_li_TCxF, ~ time_class | fenced) %>% 
  pairs(reverse = TRUE) %>% 
  plot() +
  theme_minimal() +
  geom_vline(aes(xintercept = 0), color= "black", size=1, lty="dashed") +
  labs(subtitle = "Point estimate displayed: median 
HPD interval probability: 0.95", x = "Estimate", y = "Contrast", caption = "WC & WNC Max Height, Combined Willow, Results are given on the log (not response) scale.")

# table of contrasts. Results are given on the log (not the response) scale.
emmeans(wc.wnc.stmod_li_TCxF, ~ time_class | fenced) %>% 
  pairs(reverse = TRUE) %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric),round, 2)) %>% 
  gt() %>% 
  tab_header(title = "Pairwise contrasts", subtitle = "Results are given on the log (not the response) scale.")
```

# References
