---
title: "EVMP covariate analysis"
author: ""
date: ""
output: 
  html_document: 
    theme: yeti
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(knitr)
opts_knit$set(root.dir=normalizePath('../')) # this is required if Rmd is nested below the project directory
opts_chunk$set(fig.path = "../output/figures/") # corrected path and added dev. Needed to specify a subdirectory for figs
```

```{r, echo=FALSE}
# package install
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(fs))
suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(janitor))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(mapview))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(ggstance))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(skimr)) 
suppressPackageStartupMessages(library(dataMaid))
library(lubridate)
library(mapview)
library(visdat)
library(gt)
```


## Climate

### Estes Park

Dataset	Daily Summaries
Order Start Date	1998-08-01 00:00
Order End Date	2020-08-22 23:59
Output Format	Custom GHCN-Daily CSV
Data Types	MDPR, DAPR, PRCP, SNWD, SNOW, TMAX, TMIN, TOBS
Custom Flag(s)	Station Name
Units	Metric
Stations/Locations	ESTES PARK 3 SSE, CO US (Station ID: GHCND:USC00052761)

```{r}

## import
ep.raw <- read_csv("./data/covariates/climate/NCDC_EstesPark/2257130.csv") ## METRIC UNITS

# The five core values are:
# PRCP = Precipitation (mm or inches as per user preference, inches to hundredths on Daily Form pdf file)
# SNOW = Snowfall (mm or inches as per user preference, inches to tenths on Daily Form pdf file)
# SNWD = Snow depth (mm or inches as per user preference, inches on Daily Form pdf file)
# TMAX = Maximum temperature (Fahrenheit or Celsius as per user preference, Fahrenheit to tenths on
# Daily Form pdf file
# TMIN = Minimum temperature (Fahrenheit or Celsius as per user preference, Fahrenheit to tenths on
# Daily Form pdf file

# Processing Completed	2020-08-24
# Stations	
# GHCND:USC00052761
# Begin Date	1998-08-01 00:00
# End Date	2020-08-22 23:59
# Data Types	
# MDPRDAPRPRCPSNWDSNOWTMAXTMINTOBS
# Units	
# Metric
# Custom Flag(s)	
# Station Name
# Eligible for Certification	No

```

```{r}
ep.cln <- ep.raw %>% 
  clean_names() %>% 
  mutate(month = month(date, label = TRUE, abbr = TRUE)) %>% 
  mutate(doy = lubridate::yday(date)) %>%
  mutate(year = lubridate::year(date))

```

```{r}

ep.annual.sum.prcp <- ep.cln %>% 
  group_by(year) %>% 
  summarise(ppt_mm_ann = sum(prcp)) %>% 
  ungroup()

ep.plt1 <- ep.annual.sum.prcp %>%
  filter(year > 2007 & year < 2019) %>% 
  mutate(year = as.integer(year)) %>% 
  ggplot(aes(year,ppt_mm_ann)) +
  # geom_line() +
  # geom_point() +
  geom_col(fill = 'grey40') +
  theme_minimal() +
  scale_x_continuous(breaks = seq(from = 2008, to = 2018, by = 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Year", y = "Annual precipitation (mm)", caption = "Estes Park Station GHCND:USC00052761")

ep.plt1
# ggsave("./output/figures_exported/EstesPark_AnnualPrcp.png", width = 3.85, height = 3.85)

```


### Bear Lake Snotel
```{r}
# data\covariates\climate\BearToothSnotel
snotel <- read_csv("./data/covariates/climate/BearLakeSnotel/snotel_2read.csv") %>% 
  janitor::clean_names()
```

```{r}
snotel <- snotel %>% 
  rename(swe = snow_water_equivalent_in_start_of_day_values) %>% 
  rename(ppt_accum_in = precipitation_accumulation_in_start_of_day_values) %>% 
  rename(temp_max_f = air_temperature_maximum_deg_f) %>%
  rename(temp_min_f = air_temperature_minimum_deg_f) %>%
  rename(temp_avg_f = air_temperature_average_deg_f) %>%
  rename(ppt_incr_in = precipitation_increment_in) 

snotel <- snotel %>%
  mutate(date = mdy(date)) %>% 
  mutate(month = month(date,label = TRUE),
         doy = yday(date),
         year = year(date))

```

```{r}

snotel <- snotel %>% 
  mutate(ppt_accum_mm = 25.4*ppt_accum_in) %>% 
  mutate(ppt_incr_mm = 25.4*ppt_incr_in)

snotel %>%
  visdat::vis_dat()
```

What day had the latest peak swe?


```{r}
top_ppt_accum <- snotel %>% 
  group_by(year) %>% 
  summarise(ppt_accum_mm_max = max(ppt_accum_mm)) %>% 
  ungroup()

ppt.pl <- top_ppt_accum %>% 
  filter(year >1980 & year < 2019) %>% 
  # mutate(year = as.character(year)) %>%
  ggplot(aes(year,ppt_accum_mm_max)) +
  geom_col(fill = "ivory4") +
  labs(x="", y = "Accumulated precipitation (mm)", caption = "Bear Lake Snotel Data") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_continuous(breaks = seq(from = 1980, to = 2018, by = 5))
  
# ppt.pl

```

```{r}
top_swe <- snotel %>% 
  mutate(year = as.character(year)) %>%
  group_by(year) %>% 
  summarise(sum_swe = sum(swe)) %>% 
  ungroup()


swe.pl <- top_swe %>%
  filter(year>1980 & year <2019) %>% 
  # mutate(year = as.character(year)) %>%
  mutate(year = as.integer(year)) %>%
  ggplot(aes(year,sum_swe)) +
  geom_col(fill = "grey60") +
  # geom_jitter() +
  labs(x="", y = "SWE") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_continuous(breaks = seq(from = 1980, to = 2018, by = 5))

swe.pl

# ggsave("./output/figures_exported/bearLk_SWE.png", width = 3.85, height = 3.85)

```

```{r}
cowplot::plot_grid(swe.pl,ppt.pl, labels = "AUTO")
# ggsave("./output/figures_exported/bear_lk_swe_ppt.png", width = 6.5, height = 3.8)
```


```{r}
top_swe %>% 
  select(year, sum_swe) %>% 
  datatable(rownames = FALSE, filter = "top")
```

## Physiographic
```{r}
# distince from stream



```

## Elk
```{r}

elk.raw <- read_csv("./data/covariates/elk/RMNP_Population_summary.csv") %>% 
  janitor::clean_names()

```

```{r}
elk.raw <- elk.raw %>% 
  mutate(year = as.integer(year))

## write to disk
# write_csv(elk.raw, "./data/covariates/elk/RMNP_elk_pop_summary_cln.csv")

elk.raw %>% 
  ggplot(aes(x = year, y = median)) +
  geom_point(color = "blue") +
  geom_line(color = "blue") +
  geom_line(aes(y = p025), color = "red", lty = "dashed") +
  geom_line(aes(y = p975), color = "red", lty = "dashed") +
  theme_minimal() +
  labs(x = "Year", y = "Median elk populatiop estimate", caption = "Modeled elk pop: RMNP_Population_summary.csv ")

# ggsave("./output/figures_exported/elk_population.png", width = 5, height = 4, dpi = 300)
```

```{r}
elk.raw %>% 
  gt::gt() %>% 
  gt::gtsave("./output/tables/elkpop.rtf")

elk.raw %>% 
  summarytools::descr(median) 

elk.raw %>% 
  arrange(desc(median))

elk.raw %>% 
  arrange((median))
```


## elk vs climate
```{r}
## Hobbs modeled elk vs Estes Park prcp

elk.raw %>% arrange(desc(year))
elk.ep_prcp <- left_join(ep.annual.sum.prcp,elk.raw, by = "year")

elk.ep_prcp %>% 
  ggplot(aes(x = year)) +
  geom_col(aes(y = ppt_mm_ann)) +
  geom_col(aes(y = median), color = "red")

pl.cov1 <- elk.ep_prcp %>% 
  ggplot() +
  geom_point(aes(x = ppt_mm_ann, y = median)) +
  # geom_smooth(aes(x = ppt_mm_ann, y = median)) +
  geom_smooth(aes(x = ppt_mm_ann, y = median), method = "glm") +
  theme_minimal() +
  labs(x= "Annual precipitation (mm)", y = "Number of elk", caption = "Hobbs modeled elk vs Estes Park prcp") 

pl.cov1
# ggsave("./output/figures_exported/covariates/elk_v_prcp.png", width = 3.75, height = 4)  

top_swe <- top_swe %>% 
  mutate(year = as.integer(year))

elk.ep_EPprcp_BLkppt <- left_join(elk.ep_prcp, top_ppt_accum, by = "year")
```

```{r}
pl.cov2 <- elk.ep_EPprcp_BLkppt %>% 
  ggplot() +
  geom_point(aes(x = ppt_accum_mm_max , y = median)) +
  # geom_smooth(aes(x = ppt_mm_ann, y = median)) +
  geom_smooth(aes(x = ppt_accum_mm_max , y = median), method = "glm") +
  theme_minimal() +
  labs(x= "Annual precipitation (mm)", y = "Number of elk", caption = "Hobbs modeled elk vs Bear Lake SWE")
pl.cov2

lm(median ~ ppt_mm_ann, data = elk.ep_EPprcp_BLkppt)

```


```{r}
library(moderndive)
mod.elkXepPrcp <- lm(median ~ ppt_mm_ann, data = elk.ep_EPprcp_BLkppt) 

mod.elkXepPrcp %>% 
  moderndive::get_regression_table() %>% 
  gt() %>% 
  tab_header(title = "lm: elk media ~ Estes Park annual prcp")


mod.elkXblkPrcp <- lm(median ~ ppt_accum_mm_max, data = elk.ep_EPprcp_BLkppt) 

## broom::
broom::tidy(mod.elkXblkPrcp)

## report::
# devtools::install_github("neuropsychology/report")

library(report)
mod.elkXblkPrcp
# report::report(mod.elkXepPrcp)
# report::report(mod.elkXblkPrcp)

broom::tidy(mod.elkXblkPrcp) %>% 
  gt() %>% 
  tab_header(title = "lm: elk media ~ Bear Lk annual prcp")

```

```{r}
# bayesian lm
library(rstanarm)

stanlm.elkXep <- stan_glm(median ~ ppt_accum_mm_max, data = elk.ep_EPprcp_BLkppt,
                         chains = 5, iter = 1000,
                         seed = 12345,
                         refresh = 0)

plot(stanlm.elkXep, "hist", pars = c("ppt_accum_mm_max", "sigma"), 
     transformations = list(sigma = "log"))
report::report(stanlm.elkXep )

```

```{r}
## plot grid
cowplot::plot_grid(pl.cov1, pl.cov2, labels = "AUTO")
# ggsave("./output/figures_exported/covariates/prcp_x_elk_2panel.png", width = 6.5, height = 3.75, dpi = 300)
```


## elk vs offtake upland


## climate vs elk

## Moose

