---
title: "EVMP_01"
author: "Ed Gage"
date: "October 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE)
opts_knit$set(root.dir=normalizePath('../')) # this is required if my Rmd is nested below the project directory
opts_chunk$set(fig.path = "../output/figures/") # corrected path and added dev .. if I want to specify a subdirectory for figs

# see discussion here: https://stackoverflow.com/questions/24585254/working-with-knitr-using-subdirectories

```


**Updated: `r format(Sys.time(), '%Y %B %d')`**

## Introduction

This document provides a basic exploration of EVMP files provided by RMNP in September, 2018.

Analyses aim to characterize basic data structure and identify potential issues such as:

* Inconsistently named/typed factors    
* Missing values
* Data values outside of expected range or showing unusual patterns

In addition, data are cleaned in preparation for further analyses.

```{r,echo=FALSE}
# library(here)
# here()
# install.packages("bindrcpp")
suppressPackageStartupMessages(library(tidyverse))
library(fs)
suppressPackageStartupMessages(library(sf))
# library(raster)
suppressPackageStartupMessages(library(janitor))
suppressPackageStartupMessages(library(readxl))
# library(glue)
suppressPackageStartupMessages(library(mapview))
# library(ggmap)
# library(ggrepel)
suppressPackageStartupMessages(library(viridis))
library(ggExtra)
library(DT)
library(kableExtra)
suppressPackageStartupMessages(library(lubridate))
library(anytime)
suppressPackageStartupMessages(library(compare))
suppressPackageStartupMessages(library(skimr)) ## some useful functions
suppressPackageStartupMessages(library(dataMaid))

```

##  