---
title: "EVMP working analysis"
output: revealjs::revealjs_presentation
---


```{r,echo=FALSE}
# library(here)
# here()
# install.packages("bindrcpp")

knitr::opts_knit$set(root.dir=normalizePath('../')) # this is required if Rmd is nested below the project directory


suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(fs))
suppressPackageStartupMessages(library(sf))
# library(raster)
suppressPackageStartupMessages(library(janitor))
suppressPackageStartupMessages(library(readxl))
# library(glue)
suppressPackageStartupMessages(library(mapview))
# library(ggmap)
# library(ggrepel)
suppressPackageStartupMessages(library(viridis))
# library(ggExtra)
suppressPackageStartupMessages(library(ggstance))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(DT))
# library(kableExtra)
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(skimr)) 
suppressPackageStartupMessages(library(dataMaid))
# library(ggmap)
library(lubridate)
library(mapview)
library(gt)

```

# Introduction

```{r, eval=FALSE}
# This document provides a basic exploration of EVMP files provided by RMNP in October of 2018.  The code and analyses aim to characterize basic data structure, identify problems, and clean and reorganize as needed for plotting and modeling. Specific objectives include identifying issues such as:
# 
# * Inconsistently named/typed factors    
# * Missing values
# * Data values outside of expected range or showing unusual patterns

# Cleaned and archivable data sets will be produced to facilitate future data analysis and aid RMNP in their management goals.

```



# Aspen data

- Vegetation Collection
- Data from the first 10 years of monitoring aspart of the EVMP and will be used to assess progress towards program goals

## Methods: Data collected

Vegetation shrub offtake by herbivores
- assessed through annual subsampling of sites  
2. Willow height and cover
- resampled every 5 years  
3. Aspen stem density and stem size distribution
- assessed every 5 years through a full resampling of sites. 


## Baseline data collection
 - Baseline measurements collected between 2006–2009 
 - Subsample of plots visited annually beginning in 2007 for upland plots and 2009 willow plots and conducted through 2018
  - Willow offtake measured in May and June (weather depedent)
  - Other variables were assessed the year of plot establishment and during June through September in 2013 and 2018.
## Upland line interrcept

Upland line intercept transect data were collected in core upland (UC) and upland noncore areas (UNC) along a 30m transect line. Observations included the presence of tracks, scat, burrows, plant damage, or browse. In addition, observation of animal sign were made for elk, deer, moose, ground squirrells, tent catippilars, big horn sheep, and raptors.

[add more...]

```{r}
# xls_example <- readxl_example("datasets.xls")
# excel_sheets(xls_example)

# willow.off <- read_xlsx("data/EVMP_data/provisional_data_20180920/Willow Offtake Data 2009-2018.xlsx")
# read_excel(willow.off, sheet = "chickwts")
# excel_sheets(willow.off)
# 
```


# Site info

```{r}
site.info.clean <- read_csv("./data/EVMP_derived/site_info_clean.csv")
```


```{r, eval=FALSE}
## create some tables

gt.siteinfo <- site.info.clean %>%
  dplyr::select(-REMOVED) %>% 
  gt::gt() %>%
  tab_header(title = md("**Valid EVMP plots from 'site_info' worksheet**")) 

gt.siteinfo <- gt.siteinfo %>% 
  tab_style(
    style = list(
      cell_fill(color = "lightcyan"),
      cell_text(weight = "bold")
      ),
    locations = cells_data(
      columns = vars(FENCED),
      rows = FENCED == "Y")
  ) %>% 
  tab_style(
    style = list(
      cell_fill(color = "ivory")),
    locations = cells_data(
      columns = vars(FENCED),
      rows = FENCED == "N")
  ) 

## add color for pType
gt.siteinfo %>% 
    tab_style(
    style = list(
      cell_fill(color = "lightgreen"),
      cell_text(weight = "bold")
      ),
    locations = cells_data(
      columns = vars(pType),
      rows = pType == "willow")
  ) %>% 
  tab_style(
    style = list(
      cell_fill(color = "ivory2"),
      cell_text(weight = "bold")
    ),
    locations = cells_data(
      columns = vars(pType),
      rows = pType == "aspen")
  ) %>% 
  tab_style(
    style = list(
      cell_fill(color = "pink"),
      cell_text(weight = "bold")
      ),
    locations = cells_data(
      columns = vars(pType),
      rows = pType == "upland")
  )

```


```{r}
# return here
#### create SF from coordinates
site.info.clean.sf <- site.info.clean %>%
  mutate_at(2:3, as.numeric) %>% 
  st_as_sf(coords = c("UTM_E_NAD83", "UTM_N_NAD83"), crs = 26913)

```

## Map of Kawuneeche Valley EVMP sites


```{r}

site.info.clean.sf %>% 
  filter(VALLEY == "KV") %>% 
  mapview::mapview(., zcol = 'pType')

```
